---
title: "San Jacinto ForestGEO"
author: "Emily Marie Purvis"
date: "2023-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## DATA MANIPULATED IN R

```{r, echo=FALSE}

# Load libraries

library(tidyverse)
library(sf)
library(stringr)
```

Load data

```{r}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load data
trees <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0008\\data\\SJFDP_Census_4ha_full_202230427.csv")

subplots <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0008\\data\\SJFDP_Env_4ha_20230504.csv")

```

We want to any stems that aren't main stems-- spatial data was not collected on them

```{r}

# Only keep main stems

trees <- subset(trees, Main!="S")

```

Calculate plot centroid

```{r}

# x coordinate centroid is the 16s

# y coordinate centroid is the Fs

# therefore F16 is the centroid of the whole plot

```

Add OFO plot ID and subplot IDs

OH NO PROBLEMS:
- some plants were shot from quadrat corners outside the grid (e.g. the Ks and the 21s)
- some plants were measured from along the grid line with a tape measure (e.g. not shot from a quadrat corner)

```{r}

# plot ID

trees <- trees %>%
  add_column(plot_id_ofo = "0081")

# subplot ID

trees <- trees %>%
  add_column(subplot_id_ofo = "0")

# going to add subplot ID to the subplot datasheet and then merge it into the tree datasheet

# add column

subplots <- subplots %>%
  add_column(subplot_id_ofo = "0")

# okay so the subplots for this dataset are confusing! the quadrats are named by the point of the lower left corner. however, we are defining subplots based on what quadrat corner they were measured from, not what quadrat corner they are in. as a result, 1. subplots may overlap and 2. plants may not be in the "actual" subplot of the corner they were measured from.

# if there is an "RF_grid_corner" value in the tree datasheet, that's the subplot. if there is no "RF_grid_corner," the subplot is the "Quad" value

# sort in descending order

subplots <- arrange(subplots, Quad)

# remove subplots not used for our purposes here

# first the 21s

# subplots <- subset(subplots, Quad!="A21" & Quad!="B21" & Quad!="C21" & Quad!="D21" & Quad!="E21" & Quad!="F21" & Quad!="G21" & Quad!="H21" & Quad!="I21" & Quad!="J21")

# then the Ks

# subplots <- subset(subplots, Quad!="K11" & Quad!="K12" & Quad!="K13" & Quad!="K14" & Quad!="K15" & Quad!="K16" & Quad!="K17" & Quad!="K18" & Quad!="K19" & Quad!="K20" & Quad!="K21")

# add subplot ID numbers to each quadrat

for(i in 1:nrow(subplots)) {
subplots$subplot_id_ofo[i] <- i + 211
}

# change some column names and get rid of some columsn in the subplot data sheet to prepare to merge it into the tree data sheet

subplots <- subset(subplots, select = c(Quad, Lat, Long, subplot_id_ofo))

subplots <- subplots %>% rename("QuadLat" = "Lat") %>% rename("QuadLong" = "Long") 

# !!!!we're going to assign trees to subplots based on which quadrat corner they were measured from, NOT which quadrat they are in!!!!

# change the name of the "RF_Grid_Corner"	(The grid pole that the rangefinder was shot from) column to "Quad"

trees <- trees %>% 
       rename("Quad" = "RF_Grid_Corner")

```

Tree lat and tree lon

```{r}

```


Tree DBH is in cm :)

Tree species

```{r}

# putting FIA species codes in the pre-existing "Code" column, which I'm renaming to "Species Code"

trees$Code[trees$Code == 'ABICON'] <- '015'
trees$Code[trees$Code == 'CALDEC'] <- '081'
trees$Code[trees$Code == 'PINLAM'] <- '117'
trees$Code[trees$Code == 'PINPON'] <- '122'
trees$Code[trees$Code == 'PINJEF'] <- '116'
trees$Code[trees$Code == 'QUEKEL'] <- '818'
trees$Code[trees$Code == 'UNKN'] <- '000'

trees$Code[trees$Code == 'QUEWIS'] <- '839'
trees$Code[trees$Code == 'QUECHR'] <- '805'
trees$Code[trees$Code == 'PINCOU'] <- '109'
trees$Code[trees$Code == 'QUEXMO'] <- '800' # the Quercus x moreha (Q. kelloggii x wislizeni) is being labelled as Quercus spp.


trees <- trees %>% 
       rename("SpeciesCode" = "Code")
```

Status Live/Dead

```{r}

# codes to L and D

trees$Status[trees$Status == 'DS'] <- 'D'
trees$Status[trees$Status == 'LI'] <- 'L'

# remove trees that are dead and down

trees <- subset(trees, Status!="Dt")
```

Damage codes

BB	Bark beetle
CD	Crown >75% damaged
FS	Fire scar
L	Leaning > 30 degrees
MS	Mistletoe
PD	Pest damage
SBAD	Stem broken above DBH
T	Topped

```{r}

trees$Codes <- str_replace(trees$Codes, "BB", "11000")
trees$Codes <- str_replace(trees$Codes, "CD", "90000") # unclear what kind of crown damage this is referring to (biotic or abiotic, mechanical damage, flagging...it's a mystery) so I'm labelling this as "other damages"
trees$Codes <- str_replace(trees$Codes, "FS", "30000")
trees$Codes <- str_replace(trees$Codes, "L", "90006")
trees$Codes <- str_replace(trees$Codes, "MS", "23000") # general code for parisitic/epiphytic plants, there are so many mistletoes 
trees$Codes <- str_replace(trees$Codes, "PD", "10000") # general code for insect damage
trees$Codes <- str_replace(trees$Codes, "SBAD", "90001")
trees$Codes <- str_replace(trees$Codes, "T", "60001")

```

Distance to pith from wherever the tree was shot from

```{r}

# add a new column for horizontal distance to pith

trees <- trees %>%
  add_column(horiz_distance_pith = "")

# calculate

trees$horiz_distance_pith = trees$HD + (0.5 * 0.01 * trees$DBH)

# get rid of NAs-- only the trees measured with distance/azimuth from a quadrat corner will have this measurement

trees$horiz_distance_pith[is.na(trees$horiz_distance_pith)] <- ""

```






## ALTERATIONS MADE IN EXCEL

Manually separated damage codes into damage code columns 
