---
title: "vp_ctlsierra2022_v1"
author: "Emily Marie Purvis"
date: "2023-10-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ALTERATIONS MADE IN R

```{r}
# Load data

# tree data

treedata = read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\{0001}\\data\\data\\vp_ctlsierra2022_v1_trees.csv", header = TRUE)

# plot data 

plotdata = read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\{0001}\\data\\data\\vp_ctlsierra2022_v1_plots.csv", header = TRUE)
```

- Live/dead from 1/2 to L/D

```{r}
# Live/Dead to L/D/M

treedata$live_dead[treedata$live_dead == '1'] <- 'L'

treedata$live_dead[treedata$live_dead == '2'] <- 'D'
```

- Changed from imperial to metric (height, DBH, distance)

```{r}

# DBH in inches to DBH in centimeters

# add a new column for DBH in centimeters 

treedata <- treedata %>%
  add_column(DBH_cm = "")

# calculate 

treedata$DBH_cm = treedata$dbh * 2.54

# height in feet to height in meters 

# add a new column for height in meters 

treedata <- treedata %>%
  add_column(height_m = "")

# calculate 

treedata$height_m = treedata$height * 0.3048

# horizontal distance in feet to horizontal distance in meters 

# add a new column for distance in meters 

treedata <- treedata %>%
  add_column(distance_m = "")

# calculate 

treedata$distance_m = treedata$distance * 0.3048
```

- Calculated distance to pith with distance + 1/2(DBH)

```{r}
# Horizontal distance from plot center to tree surface to horizontal distance from plot center to tree pith 

# add a new column for horizontal distance to pith 

treedata <- treedata %>%
  add_column(horiz_distance_pith = "")

# calculate 

treedata$horiz_distance_pith = treedata$distance_m + (0.5 * 0.01 * treedata$DBH_cm)
```

- Calculated ohvis from canopy position

```{r}
# Calculating tree ohvis from crown position 

# add a new column for ohvis 

treedata <- treedata %>%
  add_column(ohvis = "")

# calculate ohvis from canopy position 

treedata$ohvis <- case_match(
  treedata$canopy_position,
  c(1, 2, 3) ~ "TRUE",
  c(4, 5) ~ "FALSE",
  .default = ""
)
```

- FIA forest type codes in this dataset:
  261 – White fir
  371 – California mixed conifer
  262 – Red fir
  223 – Jeffrey pine / Coulter pine / bigcone Douglas-fir [these were all Jeffery   pine]
  221 – Ponderosa pine
  941 – Tanoak

- Added a non-FIA-coded forest type, "001 – Mixed evergreen"

```{r}
# Forest types to FIA codes

plotdata$forest_type_field[plotdata$forest_type_field == 'White fir'] <- '261 – White fir'

plotdata$forest_type_field[plotdata$forest_type_field == 'Red fir'] <- '262 – Red fir'

plotdata$forest_type_field[plotdata$forest_type_field == 'California mixed conifer'] <- '371 – California mixed conifer'

plotdata$forest_type_field[plotdata$forest_type_field == 'Mixed evergreen'] <- '001 – Mixed evergreen'

plotdata$forest_type_field[plotdata$forest_type_field == 'Jeffrey pine'] <- '223 – Jeffrey pine / Coulter pine / bigcone Douglas-fir'

plotdata$forest_type_field[plotdata$forest_type_field == 'Ponderosa pine'] <- '221 – Ponderosa pine'

plotdata$forest_type_field[plotdata$forest_type_field == 'Tanoak'] <- '941 – Tanoak'
```

- Export

```{r}
# Export to csv

write.csv(treedata, "C:\\Users\\emily\\Desktop\\ALTERED_vp_ctlsierra2022_v1_trees.csv")
```

# ALTERATIONS MADE IN EXCEL

- NAs were examined and deleted manually

- FIA damage codes were all in one column in the original data set; these were separated manually due to lots of variation in how they were written

- FIA damage codes in this dataset: 
  90006 = crook/sweep
  90005 = forked below merch
  90004 = forked top
  90011 = open wound
  90002 = dead top
  90001 = broken top
  21000 = Root/Butt Diseases
  26000 = Stem Rusts

- Some column values accidentally contained notes instead of their intended values. These were manually moved to the "notes" column, leaving blanks

- Six trees were eliminated from the dataset manually because they lacked spatial data

- Some errors were corrected manually (e.g. trees listed as alive that had both a decay class and a "0" crown ratio were changed to dead; trees listed as dead that had both no decay class and a positive crown ratio were changed to live)


