---
title: "20231004_SSI_grassvalley_oneplot"
author: "Emily Marie Purvis"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ALTERATIONS MADE IN R

```{r, echo=FALSE}

# Load libraries 

library(tidyverse)
library(readxl)
library(sf)
library(pracma)
```


```{r}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load data

treedata = read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0003\\data\\FOCAL_LAB_OnePlot_SSI_Corrected.xlsx"),sheet=3,col_names = TRUE)

```

Adding OFO plot_id

```{r}
# add a new column with the right plot ID

treedata <- treedata %>%
  add_column(plot_id_ofo = "0052")

```

Imperial to metric

```{r}

treedata$distance_meters = treedata$DISTANCE_FT*0.3048

treedata$dbh_cm = treedata$DIAMETER*2.54

treedata$height_meters = treedata$HT*0.3048

```

One column of azimuths (right now there are two)

```{r}

# some azimuth measurements are in one column and some are in another, making a new column that has all the correct azimuth measurements

treedata$Azimuth4Real <- 0

for(i in 1:nrow(treedata)) {
  if(is.na(treedata$AZM_corrected[i])) {
    treedata[i,]$Azimuth4Real = (treedata[i,]$AZM)
  }
  else {
    treedata[i,]$Azimuth4Real = treedata[i,]$AZM_corrected
  }
}
```

Add plot center coordinates to a dataframe

```{r}

# make the plot center coordinates into a spatial dataframe

plotcenter <- matrix(c('1', -120.905293, 39.199336), ncol=3, byrow=TRUE)

colnames(plotcenter) <- c('PLOT_ID','PlotCenterEasting','PlotCenterNorthing')

plotcenter <- as.data.frame.matrix(plotcenter)

plotcenter_sp <- st_as_sf(plotcenter, coords = c("PlotCenterEasting", "PlotCenterNorthing"), crs = 4326, remove=F)

# convert plot center coordinates to UTM10N so calculations in the next step work

plotcenter_UTM10N <- st_transform (plotcenter_sp, 32610)

# extract the geometry features into columns of lat and long, then merge back into the spatial data frame

UTM10N_plotcoords <- data.frame(plotcenter_UTM10N$'PLOT_ID', st_coordinates(plotcenter_UTM10N[,1], st_coordinates(plots_UTM10N[,2]))) %>% rename ('PLOT_ID'=plotcenter_UTM10N.PLOT_ID, PlotEastingUTM10N=X, PlotNorthingUTM10N=Y)

plotcenter_UTM10N <- full_join(plotcenter_UTM10N, UTM10N_plotcoords, by="PLOT_ID")

```


Tree lat and tree lon

```{r}

# add new columns to the treedata dataframe so that each individual tree has a populated plot center easting and northing

plotcenter_UTM10N$PLOT_ID = as.numeric(as.character(plotcenter_UTM10N$PLOT_ID))

treedata = full_join(treedata,st_drop_geometry(plotcenter_UTM10N),by='PLOT_ID')

# calculate the coordinates of each tree using the plot center, horizontal distance, and azimuth 

treedata = treedata %>%
  mutate(TreeEasting = as.numeric(PlotEastingUTM10N) + sin(deg2rad(Azimuth4Real)) * distance_meters,
         TreeNorthing = as.numeric(PlotNorthingUTM10N) + cos(deg2rad(Azimuth4Real)) * distance_meters)

# Tree eastings and northings from UTM back to WGS84

treedata_sp <- st_as_sf(treedata, coords = c("TreeEasting", "TreeNorthing"), crs = 32610, remove=F) 

treedata_sp <- st_transform (treedata_sp, 4326)

# extract the geometry features into columns of lat and long, then merge back into the spatial data frame

treedata_coords <- data.frame(treedata_sp$'TREE_ID', st_coordinates(treedata_sp[,1], st_coordinates(treedata_sp[,2]))) %>% rename (tree_lon=X, tree_lat=Y, TREE_ID=treedata_sp.TREE_ID)

treedata <- full_join(treedata, treedata_coords, by="TREE_ID")

```

Species: BO is black oak (818), DF is douglas fir (202), CY is canyon live oak (805), PP is ponderosa pine (122)

```{r}

treedata$SPECIES[treedata$SPECIES == 'BO'] <- '818'

treedata$SPECIES[treedata$SPECIES == 'DF'] <- '202'

treedata$SPECIES[treedata$SPECIES == 'CY'] <- '805'

treedata$SPECIES[treedata$SPECIES == 'PP'] <- '122'

```

Live/Dead

```{r}

# if the tree has a snag decay class listed, it's dead. if not, it's alive

treedata$LiveDead <- 0

treedata$LiveDead <- as.character(treedata$LiveDead)

for(i in 1:nrow(treedata)) {
  if(is.na(treedata$`SNAG DEC`[i])) {
    treedata[i,]$LiveDead = "L"
  }
  else {
    treedata[i,]$LiveDead = "D"
  }
}

# Can also be calculated with tree history codes! I did both methods and got the same results 

#treedata$LiveDeadHist <- "a"

# for(i in 1:nrow(treedata)) {
#  if((treedata$HISTORY[i]) == 1) {
#    treedata[i,]$LiveDeadHist = "L"
#  }
#  else {
#    treedata[i,]$LiveDeadHist = "D"
#  }
#}

```

Crown position: C= codominant, I= intermediate, D= dominant, O= overtopped

```{r}

# 2_dominant, 3_codominant, 4_intermediate, 5_overtopped

treedata$POSITION[treedata$POSITION == 'D'] <- '2'

treedata$POSITION[treedata$POSITION == 'C'] <- '3'

treedata$POSITION[treedata$POSITION == 'I'] <- '4'

treedata$POSITION[treedata$POSITION == 'O'] <- '5'

treedata$POSITION[treedata$POSITION == 'NA'] <- ''

treedata$POSITION[is.na(treedata$POSITION)] <- ""

treedata$POSITION <- as.numeric(treedata$POSITION)

```


Calculated ohvis from crown position

```{r}
# Calculating tree ohvis from crown position 

# add a new column for ohvis 

treedata <- treedata %>%
  add_column(ohvis = "")

# calculate ohvis from canopy position 

treedata$ohvis <- case_match(
  treedata$POSITION,
  c(1, 2, 3) ~ "TRUE",
  c(4, 5) ~ "FALSE",
  .default = ""
)

```

Get rid of NAs in snag decay class, crown ratio, and height

```{r}

treedata$`SNAG DEC`[is.na(treedata$`SNAG DEC`)] <- ""

treedata$`CRRATIO`[is.na(treedata$`CRRATIO`)] <- ""

treedata$height_meters[is.na(treedata$height_meters)] <- ""

```

Damage/Severity codes:  Broken/missing top is code 96, code 45 indicates rot, and 98 is a crook or lean.

```{r}

# broken top

treedata$DAMAGE1[treedata$DAMAGE1 == '96'] <- '90001'
treedata$DAMAGE2[treedata$DAMAGE2 == '96'] <- '90001'
treedata$DAMAGE3[treedata$DAMAGE3 == '96'] <- '90001'

# rot-- classifying this as "general diseases" because the type of rot is not specified

treedata$DAMAGE1[treedata$DAMAGE1 == '45'] <- '19000'
treedata$DAMAGE2[treedata$DAMAGE2 == '45'] <- '19000'
treedata$DAMAGE3[treedata$DAMAGE3 == '45'] <- '19000'

# crook or lean

treedata$DAMAGE1[treedata$DAMAGE1 == '98'] <- '90006'
treedata$DAMAGE2[treedata$DAMAGE2 == '98'] <- '90006'
treedata$DAMAGE3[treedata$DAMAGE3 == '98'] <- '90006'

# get rid of NAs

treedata$DAMAGE1[is.na(treedata$DAMAGE1)] <- ""
treedata$DAMAGE2[is.na(treedata$DAMAGE2)] <- ""
treedata$DAMAGE3[is.na(treedata$DAMAGE3)] <- ""

```

Calculated distance to pith with distance + 1/2(DBH)

```{r}
# Horizontal distance from plot center to tree surface to horizontal distance from plot center to tree pith 

# add a new column for horizontal distance to pith 

treedata <- treedata %>%
  add_column(horiz_distance_pith = "")

# calculate 

treedata$horiz_distance_pith = treedata$distance_meters + (0.5 * 0.01 * treedata$dbh_cm)

```

Crown ratios are coded: 1 - 1 to 10%, 2 - 11 to 20%, 3 - 21 to 30%, 4 – 31 to 40%, 5 - 41 to 50%, 6 - 51 to 60%, 7 - 61 to 70%, 8 - 71 to 80%. Decoding into percentages using the midpoint of each range.

```{r}

# get rid of NAs

treedata$CRRATIO[is.na(treedata$CRRATIO)] <- ""

# decode using midpoints

treedata$CRRATIO[treedata$CRRATIO == '1'] <- '5'
treedata$CRRATIO[treedata$CRRATIO == '2'] <- '15'
treedata$CRRATIO[treedata$CRRATIO == '3'] <- '25'
treedata$CRRATIO[treedata$CRRATIO == '4'] <- '35'
treedata$CRRATIO[treedata$CRRATIO == '5'] <- '45'
treedata$CRRATIO[treedata$CRRATIO == '6'] <- '55'
treedata$CRRATIO[treedata$CRRATIO == '7'] <- '65'
treedata$CRRATIO[treedata$CRRATIO == '8'] <- '75'

```


Export CSV

```{r}

write.csv(treedata, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\SSI.GrassValley\\20231010_SSI_grassvalley_oneplot.csv")

```


#ALTERATIONS MADE IN EXCEL



