---
title: "Lamping"
author: "Emily Marie Purvis"
date: "2023-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## DATA MANIPULATED IN R

```{r, echo=FALSE}

# Load libraries

library(tidyverse)
library(sf)
library(stringr)
```

Load data

```{r}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load data

# First the tree data! Each of these data sets is a different plot

BS2 <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Stem_maps_Lamping_et_al_2021\\BS2_StemMap.csv")

MC1 <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Stem_maps_Lamping_et_al_2021\\MC1_StemMap.csv")

SJER <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Stem_maps_Lamping_et_al_2021\\SJER_StemMap.csv")

TO1 <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Stem_maps_Lamping_et_al_2021\\TO1_StemMap.csv")

TO2 <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Stem_maps_Lamping_et_al_2021\\TO2_StemMap.csv")

UN3 <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Stem_maps_Lamping_et_al_2021\\UN3_StemMap.csv")

# now the plot data

plots <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Lamping_UAS_StemMap_boundarys\\doc.kml")

# st_crs(plots)
# plot KML is in 4326

```

PLOT LEVEL DATA

Calculate centroid of plot polygons

```{r}

#### Calculate centroid of each polygon ####

plots$centroid <- 0

plots$centroid <- st_centroid(plots)

#### calculate area of each polygon ####

# First project to a projected (meters) coordinate system with equal x and y distances, CONUS Albers Equal Area (EPSG: 5070) 

plots = st_transform(plots, crs = 5070)

st_area (plots)

# Add a new column to the spatial data frame for area and calculate

plots$area_meters <- 0

plots$area_meters = st_area (plots)
```

TREE LEVEL DATA

Add OFO plot ID

```{r}

# plot ID

BS2 <- BS2 %>%
  add_column(plot_id_ofo = "0082")

MC1 <- MC1 %>%
  add_column(plot_id_ofo = "0083")

SJER <- SJER %>%
  add_column(plot_id_ofo = "0084")

TO1 <- TO1 %>%
  add_column(plot_id_ofo = "0085")

TO2 <- TO2 %>%
  add_column(plot_id_ofo = "0086")

UN3 <- UN3 %>%
  add_column(plot_id_ofo = "0087")

```

Convert tree lat and tree lon to the right crs


```{r}

# right now tree lat and tree lon are in UTM! MC1, TO1, and TO2 are NAD83 UTM zone 10N (EPSG:26910); BS2, UN3, and SJER are NAD83 UTM zone 11N (EPSG:26911)

#### let's combine the UTM10N data sets into one data frame ####

by <- join_by(name, SPECIES, STATUS, DBH, HEIGHT, CANPOS, DRIPLINE, COMMENTS, XUTM, YUTM, elevation, plot_id_ofo)

treesUTM10N <- full_join(MC1, TO1, by)

treesUTM10N <- full_join (treesUTM10N, TO2, by)

#### and the UTM11N data sets into one dataframe ####

treesUTM11N <- full_join(BS2, UN3, by)

treesUTM11N <- transform(treesUTM11N, name = as.character(name))

# two of the SJER trees have nonsensical heights (one is "B" and one is "-"); I need this column to be numeric so I'm forcing it into numeric format and then replacing the resulting two "NA"s with blanks

SJER <- transform(SJER, HEIGHT = as.numeric(HEIGHT))

# now I can join all the UTM11N data together

treesUTM11N <- full_join (treesUTM11N, SJER, by)

treesUTM11N$HEIGHT[is.na(treesUTM11N$HEIGHT)] <- ""

#### make the regular tree dataframes into spatial dataframes ####

# first remove trees that have been bulldozed and no longer have coordinates

treesUTM10N <- treesUTM10N %>% drop_na(XUTM)

# add ID columns to make the re-merging of WGS coordinate columns easier 

treesUTM10N$ID <- seq.int(nrow(treesUTM10N))

treesUTM11N$ID <- seq.int(nrow(treesUTM11N))

# and convert to spatial dataframes

treesUTM10N_sp <- st_as_sf(treesUTM10N, coords = c("XUTM", "YUTM"), crs = 26910, remove=F)

treesUTM11N_sp <- st_as_sf(treesUTM11N, coords = c("XUTM", "YUTM"), crs = 26911, remove=F)

# change CRS to WGS84

treesUTM10N_sp <- st_transform (treesUTM10N_sp, 4326)

treesUTM11N_sp <- st_transform (treesUTM11N_sp, 4326)

# Extract new coordinates

treesUTM10N_sp_coords <- data.frame(treesUTM10N_sp$ID, st_coordinates(treesUTM10N_sp[,1], st_coordinates(treesUTM10N_sp[,2]))) 

treesUTM10N_sp_coords <- treesUTM10N_sp_coords %>% rename (ID=treesUTM10N_sp.ID, TreeWGSX=X, TreeWGSY=Y)

treesUTM11N_sp_coords <- data.frame(treesUTM11N_sp$ID, st_coordinates(treesUTM11N_sp[,1], st_coordinates(treesUTM11N_sp[,2]))) 

treesUTM11N_sp_coords <- treesUTM11N_sp_coords %>% rename (ID=treesUTM11N_sp.ID, TreeWGSX=X, TreeWGSY=Y)

# merge back into tree dataframe

treesUTM10N <- full_join(treesUTM10N, treesUTM10N_sp_coords, by="ID")

treesUTM11N <- full_join(treesUTM11N, treesUTM11N_sp_coords, by="ID")

# and combine the new WGS84 dataframes

by <- join_by(name, SPECIES, STATUS, DBH, HEIGHT, CANPOS, DRIPLINE, COMMENTS, XUTM, YUTM, elevation, plot_id_ofo, ID, TreeWGSX, TreeWGSY)

treesUTM10N <- transform(treesUTM10N, name = as.character(name))

treesUTM10N <- transform(treesUTM10N, HEIGHT = as.character(HEIGHT))

alltrees <- full_join(treesUTM10N, treesUTM11N, by)

```

DBH is in cm and height is in m :)

```{r}

# there are two trees with "NA" heights-- getting rid of these NAs

alltrees$HEIGHT[is.na(alltrees$HEIGHT)] <- ""

```


Tree species to FIA codes
		Species Code			Species Name
			 PILA				Pinus lambertiana
			 ABCO				Abies concolor
			 PIJE				Pinus Jeffreyi
			 CADE				Calocedrus decurrens
			 ABMA				Abies magnifica
			 QUKE				Quercus kelloggii
			 ABGR				Abies grandis
			 PSME				Pseudotsuga menziesii
			 CETH				Ceanothus thyrsiflorus
			 NODE				Notholithocarpus densiflorus
			 UMCA				Umbellularia californica
			 RHMA				Rhamnus sp.		
			 ARME				Arbutus menziesii
			 ACMA				Acer macrophyllum
			 FRLA				Fraxinus latifolia
			 FRPU				Fraxinus purshiana
			 SESE 				Sequoia sempervirens
			 SASP				Salix sp.
			 CONU 				Cornus nuttallii
			 COSP				Cornus nttallii
			 ALRU				Alnus rubra
			 QUWI				Quercus wislizenii
			 PISA 				Pinus sabiniana
			 RHCA				Rhamnus californica
			 ARBU 				Arbutus sp.
			 AECA				Aesculus californica
			 QUDU				Quercus douglassii
			 RHPU				Rhamnus purshiana
			 ACSP				Acer sp.
			 TSHE 				Tsuga heterophylla
			 PISP				Pinus sp.

```{r}

alltrees$SPECIES[alltrees$SPECIES == 'PILA'] <- '117'
alltrees$SPECIES[alltrees$SPECIES == 'ABCO'] <- '015'
alltrees$SPECIES[alltrees$SPECIES == 'PIJE'] <- '116'
alltrees$SPECIES[alltrees$SPECIES == 'CADE'] <- '081'
alltrees$SPECIES[alltrees$SPECIES == 'ABMA'] <- '020'
alltrees$SPECIES[alltrees$SPECIES == 'QUKE'] <- '818'
alltrees$SPECIES[alltrees$SPECIES == 'ABGR'] <- '017'
alltrees$SPECIES[alltrees$SPECIES == 'PSME'] <- '202'
alltrees$SPECIES[alltrees$SPECIES == 'NODE'] <- '631'
alltrees$SPECIES[alltrees$SPECIES == 'LIDE'] <- '631'
alltrees$SPECIES[alltrees$SPECIES == 'UMCA'] <- '981'
alltrees$SPECIES[alltrees$SPECIES == 'ARME'] <- '361'
alltrees$SPECIES[alltrees$SPECIES == 'ASME'] <- '361'
alltrees$SPECIES[alltrees$SPECIES == 'ACMA'] <- '312'
alltrees$SPECIES[alltrees$SPECIES == 'FRLA'] <- '542'
alltrees$SPECIES[alltrees$SPECIES == 'SESE'] <- '211'
alltrees$SPECIES[alltrees$SPECIES == 'CONU'] <- '492'
alltrees$SPECIES[alltrees$SPECIES == 'COSP'] <- '492'
alltrees$SPECIES[alltrees$SPECIES == 'ALRU'] <- '351'
alltrees$SPECIES[alltrees$SPECIES == 'QUWI'] <- '839'
alltrees$SPECIES[alltrees$SPECIES == 'PISA'] <- '127'
alltrees$SPECIES[alltrees$SPECIES == 'AECA'] <- '333'
alltrees$SPECIES[alltrees$SPECIES == 'QUDU'] <- '807'
alltrees$SPECIES[alltrees$SPECIES == 'ACSP'] <- '310'
alltrees$SPECIES[alltrees$SPECIES == 'TSHE'] <- '263'
alltrees$SPECIES[alltrees$SPECIES == 'PISP'] <- '100'
alltrees$SPECIES[alltrees$SPECIES == 'SASP'] <- '920'
alltrees$SPECIES[alltrees$SPECIES == 'UNKN'] <- '000'
alltrees$SPECIES[alltrees$SPECIES == 'ARBU'] <- '361'
alltrees$SPECIES[alltrees$SPECIES == 'RHCA'] <- '7200' # Rhamnus californica/Frangula californica
alltrees$SPECIES[alltrees$SPECIES == 'FRPU'] <- 'FRPU7' # Frangula/Rhamnus purshiana
alltrees$SPECIES[alltrees$SPECIES == 'RHPU'] <- 'FRPU7' # Frangula/Rhamnus purshiana
alltrees$SPECIES[alltrees$SPECIES == 'RMCA'] <- '7200 or 361' #Frangula californica; sometimes madrones were accidentally labelled with this 

# unsuccessfully trying to label RMCAs above 11 meters tall as madrones

# for(i in 1:nrow(alltrees)) {
#   if(alltrees$SPECIES [i] == 'RMCA') {
#     if (alltrees$HEIGHT[i] > 11) {
#       alltrees$SPECIES [i] == 361      # RMCA trees over 11m tall are pacific madrones
#     }
#   }
# }


# for(i in 1:nrow(alltrees)) {
#   if((alltrees$SPECIES [i] == 'RMCA') & (alltrees$HEIGHT[i] > 11)) {
#       alltrees$SPECIES [i] == 361      # RMCA trees over 11m tall are pacific madrones
#   }
# }


# alltrees$SPECIES[alltrees$SPECIES == 'RMCA' && alltrees$HEIGHT > 11] <- '361'

```

Growth form (tree, shrub, intermediate)

```{r}

alltrees <- alltrees %>%
  add_column(growth_form = "")

alltrees$growth_form[alltrees$SPECIES == '117'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '015'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '116'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '081'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '020'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '818'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '017'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '202'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '631'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '981'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '361'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '312'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '542'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '211'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '492'] <- 'intermediate'
alltrees$growth_form[alltrees$SPECIES == '351'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '839'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '127'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '333'] <- 'intermediate'
alltrees$growth_form[alltrees$SPECIES == '807'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '310'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '263'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '100'] <- 'tree'
alltrees$growth_form[alltrees$SPECIES == '920'] <- 'intermediate'
alltrees$growth_form[alltrees$SPECIES == '000'] <- 'intermediate' # unknowns could be either tree or shrub
alltrees$growth_form[alltrees$SPECIES == '7200 or 361'] <- 'intermediate'
alltrees$growth_form[alltrees$SPECIES == '7200'] <- 'intermediate'
alltrees$growth_form[alltrees$SPECIES == 'FRPU7'] <- 'intermediate'

```


Status 1/2 to live/dead

```{r}

# 1 is live, 2 is dead

# codes to L and D

alltrees$STATUS[alltrees$STATUS == '1'] <- 'L'
alltrees$STATUS[alltrees$STATUS == '2'] <- 'D'

```

Canopy position
D = Dominant 	C = Co-dominant	I = Intermediate S = Subcanopy	O = Open Growth/Isolated

```{r}

# 1 (open grown or isolated), 2 (dominant), 3 (codominant), 4 (intermediate), 5 (overtopped)

alltrees$CANPOS[alltrees$CANPOS == 'D'] <- '2'
alltrees$CANPOS[alltrees$CANPOS == 'C'] <- '3'
alltrees$CANPOS[alltrees$CANPOS == 'I'] <- '4'
alltrees$CANPOS[alltrees$CANPOS == 'S'] <- '5'
alltrees$CANPOS[alltrees$CANPOS == 'O'] <- '1'

# a bunch are labelled A mistakenly. what are they supposed to be? I'm not sure. because a bunch are blank anyways I'm going to get rid of these As

alltrees$CANPOS[alltrees$CANPOS == 'A'] <- ''

```

Calculate overhead visibility from canopy position

```{r}
# add a new column for ohvis

alltrees <- alltrees %>%
  add_column(ohvis = "")

alltrees$CANPOS <- as.numeric(alltrees$CANPOS)

# calculate ohvis from canopy position

alltrees$ohvis <- case_match(
  alltrees$CANPOS,
  c(1, 2, 3) ~ "TRUE",
  c(4, 5) ~ "FALSE",
  .default = ""
)

# introduced NAs into CANPOS column when I switched it to numeric, get rid of those now

alltrees$CANPOS[is.na(alltrees$CANPOS)] <- ""

```

Notes to damage codes

```{r}

alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "crushed tops from fallen tree", "50019")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "windsnap", "50013")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "topped", "60001")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "TOPPED", "60001")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "SWEEP", "90006")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "sweep", "90006")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "snapped", "90001")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "snap top", "90001")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "top kill", "90002")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "stump", "DELETE FROM DATA")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "heavy lean", "90006")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "lean", "90006")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "fallen", "DELETE FROM DATA")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "forks", "90004")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "FORKS", "90004")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "fork", "90004")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "FORK", "90004")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "SPLITS", "90004")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "DEAD TOP", "90002")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "dead top", "90002")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "bulldozed", "70008")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "broken top", "90001")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "arching", "90006")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "arches", "90006")
alltrees$COMMENTS <- str_replace(alltrees$COMMENTS, "arched", "90006")

```


Export CSV

```{r}

write.csv(alltrees, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Lamping\\20231113_Lamping.csv")

```


## ALTERATIONS MADE IN EXCEL

Manually assigned "madrone" to individuals in the madrone/coffeeberry confusion group above 5 m tall

Manually moved damage codes into new column

Manually deleted fallen trees