---
title: "Weeks Emerald Point"
author: "Emily Marie Purvis"
date: "2023-10-11"
output: html_document
---

```{r, echo=FALSE}

# Load libraries

library(tidyverse)
library(sf)
```

Load data

```{r, echo=FALSE}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load data
trees <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0005\\data\\ept_trees_01_rectified_inclSmall.geojson")

subplots <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0005\\data\\ept_plots_05.geojson")
  
plotpolygon <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0005\\data\\ground_map_mask_precise.geojson")
```


Calculate the area and centroid of the main plot polygon

```{r}
# First project to a projected (meters) coordinate system with equal x and y distances, CONUS Albers Equal Area (EPSG: 5070) 

plotpolygon = st_transform(plotpolygon, crs = 5070)

st_area (plotpolygon)

# Add a new column to the spatial data frame for area and calculate

plotpolygon$area_meters <- 0

plotpolygon$area_meters = st_area (plotpolygon)

# Convert back to 4326

plotpolygon = st_transform(plotpolygon, crs = 4326)

#### Calculate centroid of each polygon ####

plotpolygon$centroid <- 0

plotpolygon$centroid <- st_centroid(plotpolygon)
```
Export subplot names and geometry to a csv

```{r}
# extract the geometry features into columns of lat and long, then merge back into the spatial data frame

subplotscoords <- data.frame(subplots$'Plot_Loc', st_coordinates(subplots[,1], st_coordinates(subplots[,2]))) 

subplotscoords <- subplotscoords %>% rename ('Plot_Loc'= subplots.Plot_Loc, PlotEasting=X, PlotNorthing=Y)

#Export

write.csv(subplotscoords, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Weeks.EmeraldPoint\\20231012_WeeksEmeraldPoint_subplots.csv")
```

Add OFO plot ID and subplot IDs

```{r}

# plot ID

trees <- trees %>%
  add_column(plot_id_ofo = "0068")

# subplot IDs-- first add to subplot data, then to treedata

subplots <- subplots %>%
  add_column(subplot_id_ofo = "") 

subplots$subplot_id_ofo <- (1:nrow(subplots) + 74)

subplots_merge <- subset(subplots, select = (c(subplot_id_ofo, Plot_Loc)))

trees <- trees %>%
  add_column(subplot_id_ofo = "") 

trees = full_join(trees, st_drop_geometry(subplots_merge), by='Plot_Loc')

```

Add tree lat and lon

```{r}

# add a tree id column 

trees <- trees %>%
  add_column(tree_id_ofo = "") 

trees$tree_id_ofo <- 1:nrow(trees)

# extract the geometry features into columns of lat and long, then merge back into the spatial data frame

treescoords <- data.frame(trees$'tree_id_ofo', st_coordinates(trees[,1], st_coordinates(trees[,2]))) 

treescoords <- treescoords %>% rename ('tree_id_ofo'= trees.tree_id_ofo, TreeEasting=X, TreeNorthing=Y)

trees = full_join(trees, st_drop_geometry(treescoords), by='tree_id_ofo')
```

Units of height, dbh, height to crown, crown width (x2)

```{r}

# DBH is in centimeters! We do need to exclude trees with DBH below 7.5 though

trees <- trees[trees$DBH > 7.4, ] 

# Height is in meters

# Height to crown is in meters, has a lot of NAs. Lets get rid of em

trees$Height_to_crown[is.na(trees$Height_to_crown)] <- ""
trees$Height_to_crown[trees$Height_to_crown == 'NA'] <- ''

# Crown width is in meters

# Crown width has lots of NAs to get rid of too

trees$Crown_width_1[is.na(trees$Crown_width_1)] <- ""
trees$Crown_width_1[trees$Crown_width_1 == 'NA'] <- ''

trees$Crown_width_2[is.na(trees$Crown_width_2)] <- ""
trees$Crown_width_2[trees$Crown_width_2 == 'NA'] <- ''

```


Calculate distance to pith

```{r}

# distance measured is horizontal distance from subplot center to the face of the tree

# add a new column for horizontal distance to pith

trees <- trees %>%
  add_column(horiz_distance_pith = "")

# calculate

trees$horiz_distance_pith = trees$Distance + (0.5 * 0.01 * trees$DBH)
```

Health codes

```{r}

# get rid of NAs in health and remarks classes

trees$Health_1[is.na(trees$Health_1)] <- ""
trees$Health_1[trees$Health_1 == 'NA'] <- ''

trees$Health_2[is.na(trees$Health_2)] <- ""
trees$Health_2[trees$Health_2 == 'NA'] <- ''

trees$Health_3[is.na(trees$Health_3)] <- ""
trees$Health_3[trees$Health_3 == 'NA'] <- ''

trees$Remarks[is.na(trees$Remarks)] <- ""
trees$Remarks[trees$Remarks == 'NA'] <- ''

# Black CADEs don't have a damage code-- that's a non-pathogenic fungus called sooty mold (Arthrobotryum spongiosum)

trees$Health_1[trees$Health_1 == 'very black'] <- ''
trees$Health_1[trees$Health_1 == 'black'] <- ''
trees$Health_1[trees$Health_1 == 'slightly black'] <- ''

# get rid of "goods"

trees$Health_1[trees$Health_1 == 'good'] <- ''

# dieback (90010) for leaves sparse/very sparse and dead branches

trees$Health_1[trees$Health_1 == 'leaves sparse'] <- '90010'
trees$Health_1[trees$Health_1 == 'Leaves very sparse'] <- '90010'
trees$Health_1[trees$Health_1 == 'some branches dead'] <- '90010'
trees$Health_1[trees$Health_1 == 'only one live branch'] <- '90010'
trees$Health_1[trees$Health_1 == 'most branches dead'] <- '90010'
trees$Health_2[trees$Health_2 == 'leaves sparse'] <- '90010'
trees$Health_3[trees$Health_3 == 'leaves very sparse'] <- '90010'

# foliage discoloration (90008) for leaves browning/yellowing

trees$Health_1[trees$Health_1 == 'some red leaves'] <- '90008'
trees$Health_1[trees$Health_1 == 'many red leaves'] <- '90008'
trees$Health_1[trees$Health_1 == 'leaves yellowing'] <- '90008'
trees$Health_2[trees$Health_2 == 'some red leaves'] <- '90008'
trees$Health_2[trees$Health_2 == 'leaves yellowing slightly'] <- '90008'
trees$Health_2[trees$Health_2 == 'leaves yellowing'] <- '90008'
trees$Health_3[trees$Health_3 == 'leaves yellowing'] <- '90008'

# 90007 is bole cracks

trees$Health_1[trees$Health_1 == 'trunk cracked'] <- '90007'

# 90001 is broken top

trees$Health_1[trees$Health_1 == 'top fell off'] <- '90001'
trees$Health_2[trees$Health_2 == 'top fell off'] <- '90001'

# get rid of "sapping" --> no way of knowing if this is mechanical damage, insect damage, etc. same with "spindly" --> can't tell if they're overtopped or have another disease. same with "all lower bark gone" --> no idea if it's human girdling, animal girdling, or something else. will keep these as an unknown damage code 99000 . 

trees$Health_1[trees$Health_1 == 'sapping'] <- '99000'
trees$Health_1[trees$Health_1 == 'slightly sapping'] <- '99000'
trees$Health_1[trees$Health_1 == 'heavily sapping'] <- '99000'
trees$Health_1[trees$Health_1 == 'sapping slightly'] <- '99000'
trees$Health_1[trees$Health_1 == 'sapping heavily'] <- '99000'
trees$Health_1[trees$Health_1 == 'spindly'] <- '99000'
trees$Remarks[trees$Remarks == 'all lower bark gone'] <- '99000'

# leaning is 90006 (crook or sweep)

trees$Health_1[trees$Health_1 == 'leaning over'] <- '90006'
trees$Health_2[trees$Health_2 == 'leaning over'] <- '90006'
trees$Remarks[trees$Remarks == 'leaning over'] <- '90006'

# big gash is 90011 open wound

trees$Health_2[trees$Health_2 == 'big gash'] <- '90011'

```

Species codes

```{r}

trees$Species[trees$Species == 'ABCO'] <- '015'
trees$Species[trees$Species == 'CADE'] <- '081'
trees$Species[trees$Species == 'PIPO'] <- '122'
trees$Species[trees$Species == 'PIJE'] <- '116'
trees$Species[trees$Species == 'TSME'] <- '264'
trees$Species[trees$Species == 'SALSCO'] <- '924'
trees$Species[trees$Species == 'PILA'] <- '117'
trees$Species[trees$Species == 'PICO'] <- '108'
trees$Species[trees$Species == 'ABMA'] <-'020'
trees$Species[trees$Species == 'PI'] <-'100'
```


Get rid of NAs

```{r}

# decay class

trees$Decay_class[is.na(trees$Decay_class)] <- ""
trees$Decay_class[trees$Decay_class == 'NA'] <- ''

```


Exclude duplicate trees and columns that are just NAs

```{r}

# exclude rows full of NAs

trees <- trees %>% drop_na (Plot)

# exclude duplicates

trees <- trees[trees$duplicated == FALSE, ] 

```

Exclude trees outside plot boundary

```{r}

trees_inside <- st_intersection(trees, plotpolygon)

# also confirmed this result in qGIS!

```

Just a note that azimuth_corrected is the column we want to use here! Has declination applied :)

Export to csv

```{r}
#Export

write.csv(st_drop_geometry(trees_inside), "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Weeks.EmeraldPoint\\20231012_WeeksEmeraldPoint_treesprocessed.csv")
```


# CORRECTIONS MADE IN EXCEL

Manually dragged some damage codes that were hiding in the "remarks" section to the damage code columns 

Added growth_form data
