---
title: "FOCAL dispersal kernel project"
author: "Emily Marie Purvis"
date: "2023-10-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ALTERATIONS MADE IN R

Libraries 

```{r, echo=FALSE}

# Load libraries 

library(tidyverse)
library(sf)

```
Importing tree data

```{r, ECHO=false}

treedata = st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0004\\data\\updated\\stems_v4.gpkg")

# st_crs(treedata)
# in EPSG 4326

```

Adding OFO plot_id and subplot_id

```{r}
# add a new column with the right plot ID

treedata <- treedata %>%
  add_column(plot_id_ofo = "0")

treedata$plot_id_ofo[treedata$stem_map_name == 'Chips_1_ABCO'] <- '0057'
treedata$plot_id_ofo[treedata$stem_map_name == 'Chips_1'] <- '0058'
treedata$plot_id_ofo[treedata$stem_map_name == 'Valley_1'] <- '0059'
treedata$plot_id_ofo[treedata$stem_map_name == 'Lassic_1'] <- '0060'
treedata$plot_id_ofo[treedata$stem_map_name == 'Delta_1'] <- '0061'
treedata$plot_id_ofo[treedata$stem_map_name == 'Lassic_2'] <- '0062'
treedata$plot_id_ofo[treedata$stem_map_name == 'Delta_2'] <- '0063'
treedata$plot_id_ofo[treedata$stem_map_name == 'Chips_2'] <- '0064'
treedata$plot_id_ofo[treedata$stem_map_name == 'Delta_3'] <- '0065'
treedata$plot_id_ofo[treedata$stem_map_name == 'Creek_1'] <- '0066'
treedata$plot_id_ofo[treedata$stem_map_name == 'Creek_2'] <- '0067'

# add a new column with the right subplot ID

treedata <- treedata %>%
  add_column(subplot_id_ofo = "") 

treedata$subplot_id_ofo[treedata$center_id == 'cABCOa'] <- '0025'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (treedata$center_id == 'c24a')] <- '0026'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (treedata$center_id == 'c23a')] <- '0027'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (treedata$center_id == 'c22a')] <- '0028'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (treedata$center_id == 'c21a')] <- '0029'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (treedata$center_id == 'c17a')] <- '0030'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (treedata$center_id == 'c18a')] <- '0031'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (treedata$center_id == 'c19a')] <- '0032'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (treedata$center_id == 'c20a')] <- '0033'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_2') & (treedata$center_id == 'C16A')] <- '0034'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_2') & (treedata$center_id == 'C12A')] <- '0035'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (treedata$center_id == 'c24a')] <- '0036'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (treedata$center_id == 'c23a')] <- '0037'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (treedata$center_id == 'c22a')] <- '0038'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (treedata$center_id == 'c21a')] <- '0039'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (treedata$center_id == 'c17a')] <- '0040'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (treedata$center_id == 'c18a')] <- '0041'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (treedata$center_id == 'c19a')] <- '0042'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (treedata$center_id == 'c20a')] <- '0043'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c11')] <- '0044'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c10')] <- '0045'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c14')] <- '0046'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c15')] <- '0047'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c16')] <- '0048'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c12')] <- '0049'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c08')] <- '0050'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c07')] <- '0051'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c06')] <- '0052'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c13')] <- '0053'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c09')] <- '0054'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c01')] <- '0055'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c05')] <- '0056'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c02')] <- '0057'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c03')] <- '0058'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_1') & (treedata$center_id == 'c04')] <- '0059'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c06')] <- '0060'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c02')] <- '0061'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c01')] <- '0062'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c05')] <- '0063'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c13')] <- '0064'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c14')] <- '0065'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c10')] <- '0066'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c15')] <- '0067'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c11')] <- '0068'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c07')] <- '0069'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c03')] <- '0070'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c04')] <- '0071'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c08')] <- '0072'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c12')] <- '0073'
treedata$subplot_id_ofo[(treedata$stem_map_name == 'Creek_2') & (treedata$center_id == 'c16')] <- '0074'

# some trees were mapped with high precision GPS instead of with distance/azimuth. as such, they don't have subplot centers. going to create a "catchall" or "undefined" subplot for each plot for these trees.

# Chips_1, Valley_1, Lassic_1, Delta_1, Lassic_2, Delta_2, Chips_2, and Delta_3

treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_1') & (is.na(treedata$center_id))] <- '0333'

treedata$subplot_id_ofo[(treedata$stem_map_name == 'Valley_1') & (is.na(treedata$center_id))] <- '0334'

treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_1') & (is.na(treedata$center_id))] <- '0335'

treedata$subplot_id_ofo[(treedata$stem_map_name == 'Delta_1') & (is.na(treedata$center_id))] <- '0336'

treedata$subplot_id_ofo[(treedata$stem_map_name == 'Lassic_2') & (is.na(treedata$center_id))] <- '0337'

treedata$subplot_id_ofo[(treedata$stem_map_name == 'Delta_2') & (is.na(treedata$center_id))] <- '0338'

treedata$subplot_id_ofo[(treedata$stem_map_name == 'Chips_2') & (is.na(treedata$center_id))] <- '0339'

treedata$subplot_id_ofo[(treedata$stem_map_name == 'Delta_3') & (is.na(treedata$center_id))] <- '0340'

```

Live/dead

```{r}

treedata <- treedata %>%
  add_column(livedead = "")

# if the species column is or contains "snag", the tree is dead

treedata$livedead <- (grepl('snag', treedata$species)) | (grepl('SNAG', treedata$species)) | (grepl('Snag', treedata$species))

treedata$livedead[(treedata$livedead == 'TRUE')] <- 'D'
treedata$livedead[(treedata$livedead == 'FALSE')] <- 'L'

```

Species

```{r}

treedata$species[treedata$species == 'ABCO'] <- '015'
treedata$species[treedata$species == 'ABCO snag'] <- '015'
treedata$species[treedata$species == 'abco'] <- '015'
treedata$species[treedata$species == 'abco snag'] <- '015'
treedata$species[treedata$species == 'PSME'] <- '202'
treedata$species[treedata$species == 'psme'] <- '202'
treedata$species[treedata$species == 'PILA'] <- '117'
treedata$species[treedata$species == 'PILA snag'] <- '117'
treedata$species[treedata$species == 'pila'] <- '117'
treedata$species[treedata$species == 'CADE'] <- '081'
treedata$species[treedata$species == 'CADE Snag'] <- '081'
treedata$species[treedata$species == 'CADE snag'] <- '081'
treedata$species[treedata$species == 'cade'] <- '081'
treedata$species[treedata$species == 'cade snag'] <- '081'
treedata$species[treedata$species == 'PIPO'] <- '122'
treedata$species[treedata$species == 'PIPO snag'] <- '122'
treedata$species[treedata$species == 'pipo'] <- '122'
treedata$species[treedata$species == 'QUKE'] <- '818'
treedata$species[treedata$species == 'quke'] <- '818'
treedata$species[treedata$species == 'quke snag'] <- '818'
treedata$species[treedata$species == 'QUGA'] <- '815'
treedata$species[treedata$species == 'quga'] <- '815'
treedata$species[treedata$species == 'qu spp'] <- '800'
treedata$species[treedata$species == 'qu spp.'] <- '800'
treedata$species[treedata$species == 'qu spp. snag'] <- '800'
treedata$species[treedata$species == 'pixx snag'] <- '100'
treedata$species[treedata$species == 'pixx'] <- '100'
treedata$species[treedata$species == 'pipj'] <- '150'
treedata$species[treedata$species == 'PIPJ'] <- '150'
treedata$species[treedata$species == 'PIPJ snag'] <- '150'
treedata$species[treedata$species == 'PIJE'] <- '116'
treedata$species[treedata$species == 'pije'] <- '116'
treedata$species[treedata$species == 'abxxx snag'] <- '010'
treedata$species[treedata$species == 'abxx snag'] <- '010'
treedata$species[treedata$species == 'abxx'] <- '010'
treedata$species[treedata$species == 'abma snag'] <- '020'
treedata$species[treedata$species == 'abma'] <- '020'
# assuming QUCH
treedata$species[treedata$species == 'QUCC'] <- '805'
treedata$species[treedata$species == 'QUEV'] <- '850'

# labelling snags without other species indications as unknowns

treedata$species[treedata$species == 'SNAG'] <- '000'
treedata$species[treedata$species == 'snag'] <- '000'
treedata$species[treedata$species == 'MISSING'] <- '000'
```

Growth form

```{r}

# everything measured in these plots was a tree

treedata <- treedata %>%
  add_column(growth_form = "tree")

```


Height lowest needle-- change NAs to ""

```{r}

treedata$ht_lowest_ndl[treedata$ht_lowest_ndl == 'NA'] <- ''
treedata$ht_lowest_ndl[treedata$ht_lowest_ndl == 'N/A'] <- ''
treedata$ht_lowest_ndl[treedata$ht_lowest_ndl == 'n/a'] <- ''

```

Get rid of NAs in percent green, height to lowest needle, distance, azimuth, and notes columns

```{r}

treedata$pct_prefire_green[is.na(treedata$pct_prefire_green)] <- ""
treedata$pct_prefire_green[treedata$pct_prefire_green == 'N/A'] <- ''
treedata$pct_prefire_green[treedata$pct_prefire_green == 'n/A'] <- ''
treedata$pct_prefire_green[treedata$pct_prefire_green == 'n/a'] <- ''
treedata$pct_prefire_green[treedata$pct_prefire_green == 'NA'] <- ''

treedata$pct_current_green[is.na(treedata$pct_current_green)] <- ""
treedata$pct_current_green[treedata$pct_current_green == 'N/A'] <- ''
treedata$pct_current_green[treedata$pct_current_green == 'n/a'] <- ''
treedata$pct_current_green[treedata$pct_current_green == 'NA'] <- ''

treedata$ht_lowest_ndl[is.na(treedata$ht_lowest_ndl)] <- ""
treedata$distance[is.na(treedata$distance)] <- ""
treedata$azimuth[is.na(treedata$azimuth)] <- ""
treedata$notes[is.na(treedata$notes)] <- ""

```

Just a note that distances were measured from subplot centers to the face of the tree. Because no DBH measurements were taken in this dataset, we can't calculate distance to the pith.


Export CSV

```{r}

write.csv(treedata, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\FOCAL.dispersal.kernel.project\\20240216_FOCAL_dispersal_kernel_project.csv")

```


## ALTERATIONS MADE IN EXCEL

None

