---
title: "STEF MOC"
author: "Emily Marie Purvis"
date: "2023-10-20"
output: html_document
---

## DATA MANIPULATED IN R

```{r, echo=FALSE}

# Load libraries

library(tidyverse)
library(sf)
```

Load data

```{r, echo=FALSE}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load data
trees <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0006\\data\\RDS-2021-0061\\Data\\STEF_MOCPlots_8_9_10_11_Trees_2007_2016.csv")
  
spatialdata <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0006\\data\\RDS-2021-0061\\Data\\STEF_MOC_Plots_8_9_10_11_archive.gdb")

spatiallayers <- st_layers(dsn = "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0006\\data\\RDS-2021-0061\\Data\\STEF_MOC_Plots_8_9_10_11_archive.gdb")

spatiallayers

MOC8boundary <- st_read ("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0006\\data\\RDS-2021-0061\\Data\\STEF_MOC_Plots_8_9_10_11_archive.gdb", layer = "MOC8_Boundary")

MOC9boundary <- st_read ("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0006\\data\\RDS-2021-0061\\Data\\STEF_MOC_Plots_8_9_10_11_archive.gdb", layer = "MOC9_Boundary")

MOC10boundary <- st_read ("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0006\\data\\RDS-2021-0061\\Data\\STEF_MOC_Plots_8_9_10_11_archive.gdb", layer = "MOC10_Boundary")

MOC11boundary <- st_read ("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0006\\data\\RDS-2021-0061\\Data\\STEF_MOC_Plots_8_9_10_11_archive.gdb", layer = "MOC11_Boundary_2007")
```

PLOT LEVEL DATA

Calculate centroid of plot polygons

```{r}

# As a note-- the .gdb layers already have area of each polygon in square meters

# Right now the MOC polygons are in UTM 10N. Transform to EPSG 4326.

MOC8 = st_transform(MOC8boundary, crs = 4326)
MOC9 = st_transform(MOC9boundary, crs = 4326)
MOC10 = st_transform(MOC10boundary, crs = 4326)
MOC11 = st_transform(MOC11boundary, crs = 4326)

#### Calculate centroid of each polygon ####

MOC8$centroid <- 0
MOC9$centroid <- 0
MOC10$centroid <- 0
MOC11$centroid <- 0

MOC8$centroid <- st_centroid(MOC8)
MOC9$centroid <- st_centroid(MOC9)
MOC10$centroid <- st_centroid(MOC10)
MOC11$centroid <- st_centroid(MOC11)

write.csv(MOC8, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\STEF.MOC\\20240225_MOC8.csv")

write.csv(MOC9, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\STEF.MOC\\20240225_MOC9.csv")

write.csv(MOC10, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\STEF.MOC\\20240225_MOC10.csv")

write.csv(MOC11, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\STEF.MOC\\20240225_MOC11.csv")
```

TREE LEVEL DATA

Add OFO plot ID

```{r}

# plot ID

trees <- trees %>%
  add_column(plot_id_ofo = "0")

trees$plot_id_ofo[trees$MOC_PlotNo == '8'] <- '0069'
trees$plot_id_ofo[trees$MOC_PlotNo == '9'] <- '0070'
trees$plot_id_ofo[trees$MOC_PlotNo == '10'] <- '0071'
trees$plot_id_ofo[trees$MOC_PlotNo == '11'] <- '0072'

```

Convert tree lat and tree lon to the right crs

```{r}

# right now tree lat and tree lon are in crs: NAD 1983 Universal Transverse Mercator (UTM) Zone 10N, or epsg 26910

# first make the regular tree dataframe into a spatial dataframe

trees_sp <- st_as_sf(trees, coords = c("X_UTM", "Y_UTM"), crs = 26910, remove=F)

# change CRS to WGS84

trees_sp <- st_transform (trees_sp, 4326)

# Extract new coordinates

trees_sp_coords <- data.frame(trees_sp$OBJECTID, st_coordinates(trees_sp[,1], st_coordinates(trees_sp[,2]))) 

trees_sp_coords <- trees_sp_coords %>% rename (OBJECTID=trees_sp.OBJECTID, TreeEasting=X, TreeNorthing=Y)

# merge back into tree dataframe

trees <- full_join(trees, trees_sp_coords, by="OBJECTID")

```

Convert height from feet to meters and get rid of NAs

```{r}
# add a new column for height in meters

trees <- trees %>%
  add_column(height_m = "")

# calculate 

trees$height_m = trees$Height * 0.3048

# get rid of pesky NAs

trees$height_m[is.na(trees$height_m)] <- ""
```

Convert dbh from inches to centimeters and get rid of NAs

```{r}
# add a new column for DBH in centimeters

trees <- trees %>%
  add_column(DBH_cm = "")

# calculate 

trees$DBH_cm = trees$DBH * 2.54

# get rid of NAs

trees$DBH_cm[is.na(trees$DBH_cm)] <- ""
```

Tree species to FIA codes
• ABCO = Abies concolor 
• CADE = Calocedrus decurrens 
• PILA = Pinus lambertiana 
• PIPO = Pinus ponderosa 
• PIJE = Pinus jeffreyi 
• QUKE = Quercus kelloggii 
• SASC = Salix scouleriana 
• UNKN = Unknown species 

```{r}

trees$Species[trees$Species == 'ABCO'] <- '015'
trees$Species[trees$Species == 'CADE'] <- '081'
trees$Species[trees$Species == 'PILA'] <- '117'
trees$Species[trees$Species == 'PIPO'] <- '122'
trees$Species[trees$Species == 'PIJE'] <- '116'
trees$Species[trees$Species == 'QUKE'] <- '818'
trees$Species[trees$Species == 'SASC'] <- '924'
trees$Species[trees$Species == 'UNKN'] <- '000'

```

Status from Live/Dead to L/D

```{r}

trees$Status[trees$Status == 'Live'] <- 'L'

trees$Status[trees$Status == 'Dead'] <- 'D'
```

There are some secret damage codes hiding in the notes section!

```{r}

# add a new column for damage

trees <- trees %>%
  add_column(damage = "")

# dead top damage code
trees$damage[trees$Notes == 'dead top'] <- '90002'

# no specific damage code for something decayed below dbh-- I can't determine what caused the decay from this dataset alone-- so I'm marking it as the "unknown" damage code
trees$damage[trees$Notes == 'decayed below dbh'] <- '99000'
```


Remove NAs from snag decay class

```{r}

trees$SnagDecayClass[is.na(trees$SnagDecayClass)] <- ""
```

Export CSV

```{r}

write.csv(trees, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\STEF.MOC\\20231020_STEF_MOC.csv")

```


## DATA MANIPULATED IN EXCEL

Added growth_form data
