---
title: "Lamping"
author: "Emily Marie Purvis"
date: "2023-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## DATA MANIPULATED IN R

```{r, echo=FALSE}

# Load libraries

library(tidyverse)
library(sf)
library(stringr)
```

Load data

```{r}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load data. 

# First the stem map data. Each of these data sets is a different plot

BS2 <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\BS2.stemmap.shp")

# BS2 is in 26911 EPSG

MC1 <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\MC1.stemmap.shp")

# MC1 is in 26910 EPSG

OW2 <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\OW2.stemmap.shp")

# OW2 is in 26911 EPSG

TO1 <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\TO1.stemmap.shp")

# TO1 is in 26910 EPSG

TO2 <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\TO2.stemmap.shp")

# TO2 is in 26910 EPSG

UN3 <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\UN3.stemmap.shp")

# UN3 is in 26911 EPSG

# now the plot data. each of these data sets is a different plot. 

BS2plot<- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\BS2_plot.shp")

# BS2 is in 26911 EPSG

MC1plot <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\MC1_plot.shp")

# MC1 is in 26910 EPSG

OW2plot <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\OW2_plot.shp")

# OW2 is in 26911 EPSG

TO1plot <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\TO1_plot.shp")

# TO11 is in 26910 EPSG

TO2plot <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\TO2_plot.shp")

# TO2 is in 26910 EPSG

UN3plot <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Updated\\Stemmaps_HZald_JLamping_OFO\\Stemmaps_HZald_JLamping_OFO\\UN3_plot.shp")

# UN3 is in 26911 EPSG

```

PLOT LEVEL DATA

Calculate centroid of plot polygons

```{r}

#### Calculate centroid of each polygon ####

# BS2

BS2plotcentroid <- st_transform(BS2plot, 4326)

BS2plotcentroid$centroid <- st_centroid(BS2plotcentroid)

# MC1

MC1plotcentroid <- st_transform(MC1plot, 4326)

MC1plotcentroid$centroid <- st_centroid(MC1plotcentroid)

# OW2

OW2plotcentroid <- st_transform(OW2plot, 4326)

OW2plotcentroid$centroid <- st_centroid(OW2plotcentroid)

# TO1

TO1plotcentroid <- st_transform(TO1plot, 4326)

TO1plotcentroid$centroid <- st_centroid(TO1plotcentroid)

# TO2

TO2plotcentroid <- st_transform(TO2plot, 4326)

TO2plotcentroid$centroid <- st_centroid(TO2plotcentroid)

# UN3

UN3plotcentroid <- st_transform(UN3plot, 4326)

UN3plotcentroid$centroid <- st_centroid(UN3plotcentroid)

write.csv(BS2plotcentroid, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Lamping\\20240225_BS2.csv")

write.csv(MC1plotcentroid, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Lamping\\20240225_MC1.csv")

write.csv(OW2plotcentroid, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Lamping\\20240225_OW2.csv")

write.csv(TO1plotcentroid, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Lamping\\20240225_TO1.csv")

write.csv(TO2plotcentroid, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Lamping\\20240225_TO2.csv")

write.csv(UN3plotcentroid, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Lamping\\20240225_UN3.csv")

```

TREE LEVEL DATA

Add OFO plot ID

```{r}

# plot ID

BS2 <- BS2 %>%
  add_column(plot_id_ofo = "0082")

MC1 <- MC1 %>%
  add_column(plot_id_ofo = "0083")

OW2 <- OW2 %>%
  add_column(plot_id_ofo = "0084")

TO1 <- TO1 %>%
  add_column(plot_id_ofo = "0085")

TO2 <- TO2 %>%
  add_column(plot_id_ofo = "0086")

UN3 <- UN3 %>%
  add_column(plot_id_ofo = "0087")

```

Remove trees that fall outside of the plot polygons

```{r}

# load file with all plot boundaries

allLampingplots <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0009\\data\\Original\\Lamping_UAS_StemMap_boundarys\\doc.kml")

# st_crs(allLampingplots) is 4326

# filter stem maps one by one to remove trees that fall outside plot boundaries. first must switch them to the crs of the "allLampingplots" sf object, then switch them back.

# BS2

# st_crs(BS2) is 26911

BS2_inplot <- st_transform (BS2, 4326)

BS2_inplot <- st_filter(BS2_inplot, allLampingplots)

# dropped 9 trees!

BS2_inplot <- st_transform(BS2_inplot, 26911)

# MC1

# st_crs(MC1) is 26910

MC1_inplot <- st_transform (MC1, 4326)

MC1_inplot <- st_filter(MC1_inplot, allLampingplots)

# dropped 402 trees?!?

MC1_inplot <- st_transform(MC1_inplot, 26910)

# OW2

# st_crs(OW2) is 26911

OW2_inplot <- st_transform (OW2, 4326)

OW2_inplot <- st_filter(OW2_inplot, allLampingplots)

# dropped 80 trees!

OW2_inplot <- st_transform(OW2_inplot, 26911)

# T01

# st_crs(TO1) is 26910

TO1_inplot <- st_transform (TO1, 4326)

TO1_inplot <- st_filter(TO1_inplot, allLampingplots)

# dropped 416 trees!

TO1_inplot <- st_transform(TO1_inplot, 26910)

# TO2

# st_crs(TO2) is 26910

TO2_inplot <- st_transform (TO2, 4326)

TO2_inplot <- st_filter(TO2_inplot, allLampingplots)

# dropped 9 trees!

TO2_inplot <- st_transform(TO2_inplot, 26910)

# UN3

# st_crs(UN3) is 26911

UN3_inplot <- st_transform (UN3, 4326)

UN3_inplot <- st_filter(UN3_inplot, allLampingplots)

# dropped 26 trees!

UN3_inplot <- st_transform(UN3_inplot, 26911)

# write them out to visualize how these results look

# st_write(BS2_inplot, "C:\\Users\\emily\\Desktop\\BS2viz.gpkg")

# st_write(MC1_inplot, "C:\\Users\\emily\\Desktop\\MC1viz.gpkg")

# st_write(OW2_inplot, "C:\\Users\\emily\\Desktop\\OW2viz.gpkg")

# st_write(TO1_inplot, "C:\\Users\\emily\\Desktop\\TO1viz.gpkg")

# st_write(TO2_inplot, "C:\\Users\\emily\\Desktop\\TO2viz.gpkg")

# st_write(UN3_inplot, "C:\\Users\\emily\\Desktop\\UN3viz.gpkg")

# they look good!

```

Convert tree lat and tree lon to the right crs

```{r}

# right now tree lat and tree lon are in UTM! MC1, TO1, and TO2 are NAD83 UTM zone 10N (EPSG:26910); BS2, UN3, and OW2 are NAD83 UTM zone 11N (EPSG:26911)

# each spatial data frame has tree coordinates both as geometric points (in UTM 10N or 11N), AND tree coordinates in our desired EPSG (4326, WGS 84) in columns. we need to get rid of the "spatial" parts of these data frames to remove incompatibilities with combining them, then we can combine the data frames into one data.frame to work with. 

#### remove the spatial aspect of each stem map ####

BS2 <- st_drop_geometry(BS2_inplot)

MC1 <- st_drop_geometry(MC1_inplot)

OW2 <- st_drop_geometry(OW2_inplot)

TO1 <- st_drop_geometry(TO1_inplot)

TO2 <- st_drop_geometry(TO2_inplot)

UN3 <- st_drop_geometry(UN3_inplot)

#### let's combine the individual stem maps into one data frame ####

# first remove extraneous columns from the data.frames

joinby <- c("tree", "status", "species", "dbh", "height", "canpos", "comments", "latitude", "longitude", "plot_id_ofo")

BS2 = BS2[,(names(BS2) %in% joinby)]

MC1 = MC1[,(names(MC1) %in% joinby)]

OW2 = OW2[,(names(OW2) %in% joinby)]

TO1 = TO1[,(names(TO1) %in% joinby)]

TO2 = TO2[,(names(TO2) %in% joinby)]

UN3 = UN3[,(names(UN3) %in% joinby)]

# then make sure all the remaining columns are in the same format

BS2$tree <- as.character(BS2$tree)

MC1$tree <- as.character(MC1$tree)

TO1$tree <- as.character(TO1$tree)

TO2$tree <- as.character(TO2$tree)

UN3$tree <- as.character(UN3$tree)

# then join

by = join_by(tree, status, species, dbh, height, canpos, comments, latitude, longitude, plot_id_ofo)

alltrees <- full_join(BS2, MC1, by)

alltrees <- full_join (alltrees, OW2, by)

alltrees <- full_join (alltrees, TO1, by)

alltrees <- full_join (alltrees, TO2, by)

alltrees <- full_join (alltrees, UN3, by)

#### remove trees that don't have coordinates ####

alltrees <- alltrees %>% filter (!(latitude == 0))

```

DBH is in cm and height is in m :)

Remove heights and DBHs that are zero

```{r}

# there are multiple trees with heights and/or dbhs of 0. Harold Zald is unclear why but gave me a list of trees with height/DBH values that should be NA instead. I went through the list and confirmed that each height/DBH of 0 is actually NA. Getting rid of those 0s here. Phew!

alltrees$height[alltrees$height == 0] <- ""
alltrees$dbh[alltrees$dbh == 0] <- ""

```

Tree species to numeric codes

```{r}

alltrees$species[alltrees$species == 'PILA'] <- '117'
alltrees$species[alltrees$species == 'ABCO'] <- '015'
alltrees$species[alltrees$species == 'PIJE'] <- '116'
alltrees$species[alltrees$species == 'CADE'] <- '081'
alltrees$species[alltrees$species == 'CADE27'] <- '081'
alltrees$species[alltrees$species == 'ABMA'] <- '020'
alltrees$species[alltrees$species == 'QUKE'] <- '818'
alltrees$species[alltrees$species == 'ABGR'] <- '017'
alltrees$species[alltrees$species == 'PSME'] <- '202'
alltrees$species[alltrees$species == 'NODE'] <- '631'
alltrees$species[alltrees$species == 'NODE3'] <- '631'
alltrees$species[alltrees$species == 'LIDE'] <- '631'
alltrees$species[alltrees$species == 'UMCA'] <- '981'
alltrees$species[alltrees$species == 'ARME'] <- '361'
alltrees$species[alltrees$species == 'ASME'] <- '361'
alltrees$species[alltrees$species == 'ACMA'] <- '312'
alltrees$species[alltrees$species == 'FRLA'] <- '542'
alltrees$species[alltrees$species == 'SESE'] <- '211'
alltrees$species[alltrees$species == 'CONU'] <- '492'
alltrees$species[alltrees$species == 'CONU4'] <- '492'
alltrees$species[alltrees$species == 'COSP'] <- '492'
alltrees$species[alltrees$species == 'ALRU'] <- '351'
alltrees$species[alltrees$species == 'QUWI'] <- '839'
alltrees$species[alltrees$species == 'PISA'] <- '127'
alltrees$species[alltrees$species == 'AECA'] <- '333'
alltrees$species[alltrees$species == 'QUDU'] <- '807'
alltrees$species[alltrees$species == 'ACSP'] <- '310'
alltrees$species[alltrees$species == 'TSHE'] <- '263'
alltrees$species[alltrees$species == 'PISP'] <- '100'
alltrees$species[alltrees$species == 'SASP'] <- '920'
alltrees$species[alltrees$species == 'UNKN'] <- '000'
alltrees$species[alltrees$species == 'unkn'] <- '000'
alltrees$species[alltrees$species == 'ARBU'] <- '361'
alltrees$species[alltrees$species == 'SESE3'] <- '211'
alltrees$species[alltrees$species == 'SALIX'] <- '920'
alltrees$species[alltrees$species == 'QUWI2'] <- '839'
alltrees$species[alltrees$species == 'QUDO'] <- '807'
alltrees$species[alltrees$species == 'PINUS'] <- '100'
alltrees$species[alltrees$species == 'PISA2'] <- '127'
alltrees$species[alltrees$species == 'QUMO2'] <- '899' #Quercus x moreha, Oracle oak, hybrid of Q. kelloggii and Q. wislizeni. # NOTE: WE MADE UP THIS CODE, IT'S NOT AN OFFICIAL FIA CODE
alltrees$species[alltrees$species == 'ALRU2'] <- '351'
alltrees$species[alltrees$species == 'ACMA3'] <- '312'
alltrees$species[alltrees$species == 'ACACIA'] <- '1001' # unknown acacia species # NOTE: WE MADE UP THIS CODE, IT'S NOT AN OFFICIAL FIA CODE

# there are some NAs, replacing with blanks

alltrees$species[is.na(alltrees$species)] <- ''

```

Growth form (tree, shrub, intermediate)

```{r}

alltrees <- alltrees %>%
  add_column(growth_form = "")

alltrees$growth_form[alltrees$species == '117'] <- 'tree'
alltrees$growth_form[alltrees$species == '015'] <- 'tree'
alltrees$growth_form[alltrees$species == '116'] <- 'tree'
alltrees$growth_form[alltrees$species == '081'] <- 'tree'
alltrees$growth_form[alltrees$species == '020'] <- 'tree'
alltrees$growth_form[alltrees$species == '818'] <- 'tree'
alltrees$growth_form[alltrees$species == '017'] <- 'tree'
alltrees$growth_form[alltrees$species == '202'] <- 'tree'
alltrees$growth_form[alltrees$species == '631'] <- 'tree'
alltrees$growth_form[alltrees$species == '981'] <- 'tree'
alltrees$growth_form[alltrees$species == '361'] <- 'tree'
alltrees$growth_form[alltrees$species == '312'] <- 'tree'
alltrees$growth_form[alltrees$species == '542'] <- 'tree'
alltrees$growth_form[alltrees$species == '211'] <- 'tree'
alltrees$growth_form[alltrees$species == '492'] <- 'intermediate'
alltrees$growth_form[alltrees$species == '351'] <- 'tree'
alltrees$growth_form[alltrees$species == '839'] <- 'tree'
alltrees$growth_form[alltrees$species == '127'] <- 'tree'
alltrees$growth_form[alltrees$species == '333'] <- 'intermediate'
alltrees$growth_form[alltrees$species == '807'] <- 'tree'
alltrees$growth_form[alltrees$species == '310'] <- 'tree'
alltrees$growth_form[alltrees$species == '263'] <- 'tree'
alltrees$growth_form[alltrees$species == '100'] <- 'tree'
alltrees$growth_form[alltrees$species == '920'] <- 'intermediate'
alltrees$growth_form[alltrees$species == '000'] <- '' # unknowns could be either tree or shrub so we'll leave it blank
alltrees$growth_form[alltrees$species == '211'] <- 'tree'
alltrees$growth_form[alltrees$species == '920'] <- 'intermediate'
alltrees$growth_form[alltrees$species == '839'] <- 'tree'
alltrees$growth_form[alltrees$species == '807'] <- 'tree'
alltrees$growth_form[alltrees$species == '100'] <- 'tree'
alltrees$growth_form[alltrees$species == '127'] <- 'tree'
alltrees$growth_form[alltrees$species == '899'] <- 'tree'
alltrees$growth_form[alltrees$species == '351'] <- 'tree'
alltrees$growth_form[alltrees$species == '312'] <- 'tree'
alltrees$growth_form[alltrees$species == '1001'] <- 'intermediate'
alltrees$growth_form[alltrees$species == 'FRPU7'] <- 'intermediate'
alltrees$growth_form[alltrees$species == 'ARVI4'] <- 'shrub'
alltrees$growth_form[alltrees$species == 'RHCR'] <- 'shrub'
alltrees$growth_form[alltrees$species == 'FRCAC5'] <- 'shrub'

# trees with NA species will have a blank growth form
alltrees$growth_form[is.na(alltrees$species)] <- ''

```

Status live/dead to L/D

```{r}

alltrees$status[alltrees$status == 'live'] <- 'L'
alltrees$status[alltrees$status == 'dead'] <- 'D'

# there are some NAs, replacing with blanks

alltrees$status[is.na(alltrees$status)] <- ''

```

Canopy position
D = Dominant 	C = Co-dominant	I = Intermediate S = Subcanopy	O = Open Growth/Isolated

```{r}

# 1 (open grown or isolated), 2 (dominant), 3 (codominant), 4 (intermediate), 5 (overtopped)

alltrees$canpos[alltrees$canpos == 'dominant'] <- '2'
alltrees$canpos[alltrees$canpos == 'codominant'] <- '3'
alltrees$canpos[alltrees$canpos == 'intermediate'] <- '4'
alltrees$canpos[alltrees$canpos == 'suppressed'] <- '5'
alltrees$canpos[alltrees$canpos == 'open'] <- '1'

# there are some NAs, replacing with blanks

alltrees$canpos[is.na(alltrees$canpos)] <- ''

```

Calculate overhead visibility from canopy position

```{r}
# add a new column for ohvis

alltrees <- alltrees %>%
  add_column(ohvis = "")

alltrees$canpos <- as.numeric(alltrees$canpos)

# calculate ohvis from canopy position

alltrees$ohvis <- case_match(
  alltrees$canpos,
  c(1, 2, 3) ~ "TRUE",
  c(4, 5) ~ "FALSE",
  .default = ""
)

# introduced NAs into CANPOS column when I switched it to numeric, get rid of those now

alltrees$canpos[is.na(alltrees$canpos)] <- ""

```

Notes to damage codes

```{r}

alltrees$comments[is.na(alltrees$comments)] <- ""

alltrees$comments <- str_replace(alltrees$comments, "crushed tops from fallen tree", "50019")
alltrees$comments <- str_replace(alltrees$comments, "windsnap", "50013")
alltrees$comments <- str_replace(alltrees$comments, "topped", "60001")
alltrees$comments <- str_replace(alltrees$comments, "TOPPED", "60001")
alltrees$comments <- str_replace(alltrees$comments, "SWEEP", "90006")
alltrees$comments <- str_replace(alltrees$comments, "sweep", "90006")
alltrees$comments <- str_replace(alltrees$comments, "snapped", "90001")
alltrees$comments <- str_replace(alltrees$comments, "snap top", "90001")
alltrees$comments <- str_replace(alltrees$comments, "top kill", "90002")
alltrees$comments <- str_replace(alltrees$comments, "stump", "DELETE FROM DATA")
alltrees$comments <- str_replace(alltrees$comments, "heavy lean", "90006")
alltrees$comments <- str_replace(alltrees$comments, "lean", "90006")
alltrees$comments <- str_replace(alltrees$comments, "fallen", "DELETE FROM DATA")
alltrees$comments <- str_replace(alltrees$comments, "forks", "90004")
alltrees$comments <- str_replace(alltrees$comments, "FORKS", "90004")
alltrees$comments <- str_replace(alltrees$comments, "fork", "90004")
alltrees$comments <- str_replace(alltrees$comments, "FORK", "90004")
alltrees$comments <- str_replace(alltrees$comments, "SPLITS", "90004")
alltrees$comments <- str_replace(alltrees$comments, "DEAD TOP", "90002")
alltrees$comments <- str_replace(alltrees$comments, "dead top", "90002")
alltrees$comments <- str_replace(alltrees$comments, "bulldozed", "70008")
alltrees$comments <- str_replace(alltrees$comments, "broken top", "90001")
alltrees$comments <- str_replace(alltrees$comments, "arching", "90006")
alltrees$comments <- str_replace(alltrees$comments, "arches", "90006")
alltrees$comments <- str_replace(alltrees$comments, "arched", "90006")

```


Export CSV

```{r}

write.csv(alltrees, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\Lamping\\20240517_Lamping.csv")

```


## ALTERATIONS MADE IN EXCEL

Manually moved damage codes into new column

Manually deleted fallen trees
