---
title: "FOCAL_TNC_IRI"
author: "Emily Marie Purvis"
date: "2023-12-19"
updated: "2024-06-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA MANIPULATED IN R

```{r, echo=FALSE}

# Load libraries

library(tidyverse)
library(sf)
library(stringr)
library(readxl)
library(pracma)
```

Load data

```{r}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load tree data

Batch1trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_1\\updated (9.11.23) data and output files\\Stem_Batch_1\\Plot_Data\\StemData_Batch1_emp_edits_QC_EDITS.xlsx"),sheet=1,col_names = TRUE)

Batch2trees <- read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_2\\updated (9.11.23) data and output files\\Stem_Batch_2\\Plot_Data\\StemData_Batch2_QC_EDITS.xlsx"),sheet=1,col_names = TRUE)

Batch3trees <- read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_3\\StemData_Batch3_Master_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch4trees <- read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_4\\Stem_Batch_4\\Plot_Data\\StemData_Batch4_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch5trees <- read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_5\\Batch_5\\Plot_Data\\Stem_Batch5_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch6trees <- read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_6\\Batch_6\\Plot_Data\\StemData_Batch6_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch7trees <- read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_7\\Batch_7\\Plot_Data\\Stem_Batch7_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch8trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_8\\Batch_8\\Plot_Data\\Stem_Batch_8_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch9trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_9\\Batch_9\\Batch_9_Complete_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch10trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_10\\Batch_10_final\\Batch_10_Final_Data_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch11trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_11\\Batch_11_Final\\Batch_11_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch12trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_12\\Batch_12_updated\\Plot_Data\\Stem_Batch12_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch13trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_13\\Batch_13_Final_Geoff\\Batch13_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch14trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_14\\Batch14\\Plot_Data\\Stem_Batch14_emp_edits.xlsx"),sheet=1,col_names = TRUE)

Batch15trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_15\\Batch15\\Plot_Data\\Stem_Batch15_emp_edits.xlsx"),sheet=1,col_names = TRUE)

BatchL527trees <- read_excel (data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.Stem_Batch_L527\\L-527\\Data\\L-527.xlsx"),sheet=1,col_names = TRUE)

# Load plot center data

plotcenters <- read_sf("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.IRI.Compiled.Data\\allupdatedplotcentersWGS84.kml")

```

```{r}
# combine all batch data sheets into one data frame 

# Rename columns so they all match

Batch4trees <- Batch4trees %>% rename(`Slope distance`=`Slope Dist`, `H Distance`= `H Dist`)

Batch3trees <- Batch3trees %>% rename(`Tree#`=`Tree #`)

Batch5trees <- Batch5trees %>% rename(`Plot#`=`Plot #`)

Batch6trees <- Batch6trees %>% rename(`Plot#`=`Plot #`)

Batch7trees <- Batch7trees %>% rename(`Plot#`=`Plot #`)

Batch11trees <- Batch11trees %>% rename(`Slope distance`=`Slope Dist`, `H Distance`= `H Dist`)

Batch13trees <- Batch13trees %>% rename(`Slope distance`=`Slope Dist`, `H Distance`= `H Dist`)

Batch8trees <- Batch8trees %>% rename(`Plot#`=`Plot #`)

Batch9trees <- Batch9trees %>% rename(`Plot#`=`Plot #`)

Batch10trees <- Batch10trees %>% rename(`Plot#`=`Plot #`)

Batch12trees <- Batch12trees %>% rename(`Plot#`=`Plot #`)

Batch14trees <- Batch14trees %>% rename(`Plot#`=`Plot #`)

Batch15trees <- Batch15trees %>% rename(`Plot#`=`Plot #`)

BatchL527trees <- BatchL527trees %>% rename(`Plot#`=`Plot #`)

# combine 

combinedtrees <- rbind(Batch1trees, Batch2trees, Batch3trees, Batch4trees, Batch5trees, Batch6trees, Batch7trees, Batch8trees, Batch9trees, Batch10trees, Batch11trees, Batch12trees, Batch13trees, Batch14trees, Batch15trees, BatchL527trees)

# Rename columns to be more descriptive

combinedtrees <- combinedtrees %>% rename("PercentSlope" = `% slope`)

combinedtrees <- combinedtrees %>% rename(`Horizontal distance` = `H Distance`)

combinedtrees <- combinedtrees %>% rename(`Horizontal distance (feet)` = `Horizontal distance`)

combinedtrees <- combinedtrees %>% rename(`Slope distance (feet)` = `Slope distance`)

combinedtrees <- combinedtrees %>% rename(`Azimuth` = `AZM`)

combinedtrees <- combinedtrees %>% rename(`DBH (inches)` = `DBH`)

combinedtrees <- combinedtrees %>% rename(`Height (feet)` = `Actual HT`)

combinedtrees <- combinedtrees %>% rename(`CanopyPosition` = `Position`)

# need to delete random rows of all NAs in treedata

combinedtrees <- combinedtrees %>% drop_na(`Plot#`) # this gets rid of empty rows

combinedtrees <- combinedtrees %>% drop_na(`Azimuth`) # this gets rid of rows used for descriptions like "started new cutoff here" but otherwise don't have tree data-- manually checked each instance before running this code

```

Plot ID

```{r}

combinedtrees$`ofo_plot_id` <- "0"

combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L506'] <- '0088'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L511'] <- '0089'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L512'] <- '0090'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L515'] <- '0091'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L517'] <- '0092'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L525'] <- '0093'

combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L521'] <- '0099'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L505'] <- '0096'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L504'] <- '0095'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L536'] <- '0101'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L507'] <- '0097'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L502'] <- '0094'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L529'] <- '0100'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L508'] <- '0098'

combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L519'] <- '0102'

combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L522'] <- '0107'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L528'] <- '0104'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L524'] <- '0103'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L520'] <- '0106'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L535'] <- '0108'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L539'] <- '0109'
combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L534'] <- '0105'

combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L092'] <- '0110'

combinedtrees$`ofo_plot_id`[combinedtrees$`Plot#` == 'L527'] <- '0111'

```


Tree lat and long

```{r}

#### Add a complete horizontal distance column ####

# some of the trees were measured with horizontal distance, some with slope + slope distance, some with both. we want all trees to have horizontal distance. 

# create a blank column for horizontal distance

combinedtrees$`All Horizontal Distances (feet)` <- 0

# if horizontal distance is NA, then calculate it from percent slope and slope distance, then put that value in the horizontal distance column. if horizontal distance is not NA, use the value the crew measured

combinedtrees$`Slope distance (feet)` <- as.numeric(combinedtrees$`Slope distance (feet)`)

combinedtrees$`PercentSlope` <- as.numeric(combinedtrees$`PercentSlope`)

for(i in 1:nrow(combinedtrees)) {
  if(is.na(combinedtrees$`Horizontal distance (feet)`[i])) {
    combinedtrees[i,]$`All Horizontal Distances (feet)` = (combinedtrees[i,]$`Slope distance (feet)`)*cos(atan(combinedtrees[i,]$`PercentSlope`/100))
  }
  else {
    combinedtrees[i,]$`All Horizontal Distances (feet)` = combinedtrees[i,]$`Horizontal distance (feet)`
  }
}

# create a blank column for horizontal distance

combinedtrees$`All Horizontal Distances (m)` <- 0

combinedtrees$`All Horizontal Distances (m)` = combinedtrees$`All Horizontal Distances (feet)` * 0.3048

#### extract the geometry features of the plot center data into columns of lat and long ####

# extract coordinates

#plotcenters <- st_zm(combinedplotsWGS84)

plotcenters <- plotcenters %>% rename(`Plot#`=`Name`)

plotcentersUTM10N <- plotcenters %>% st_transform(32610)

plotcenters_coordinates <- data.frame(plotcenters$`Plot#`, st_coordinates(plotcenters[,1], st_coordinates(plotcenters[,2]))) 

plotcentersUTM10N_coordinates <- data.frame(plotcentersUTM10N$`Plot#`, st_coordinates(plotcentersUTM10N[,1], st_coordinates(plotcentersUTM10N[,2]))) 

# rename new columns

plotcenters_coordinates <- plotcenters_coordinates %>% rename (`Plot#`=plotcenters..Plot.., PlotLongitude=X, PlotLatitude=Y) 

plotcenters_coordinates <- plotcenters_coordinates %>% rename (PlotLongitudeWGS84=PlotLongitude, PlotLatitudeWGS84=PlotLatitude) 

plotcentersUTM10N_coordinates <- plotcentersUTM10N_coordinates %>% rename (`Plot#`=plotcentersUTM10N..Plot.., PlotLongitudeUTM10N=X, PlotLatitudeUTM10N=Y) 

write.csv(plotcenters, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.IRI.Compiled.Data\\20240225_IRItoOFO_plotcenters.csv")

# make sure all plot names are in the same format

# Batch 1

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-506'] <- 'L506'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-511'] <- 'L511'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-512'] <- 'L512'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-515'] <- 'L515'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-517'] <- 'L517'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-525'] <- 'L525'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-107'] <- 'S107'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-109'] <- 'S109'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-113'] <- 'S113'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-118'] <- 'S118'

# Batch 2

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-521'] <- 'L521'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-505'] <- 'L505'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-504'] <- 'L504'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-536'] <- 'L536'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-507'] <- 'L507'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-502'] <- 'L502'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-038'] <- 'S038'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-529'] <- 'L529'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-508'] <- 'L508'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-313'] <- 'S313'

# Batch 3

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-519'] <- 'L519'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-087'] <- 'S087'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-084'] <- 'S084'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-080'] <- 'S080'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-083'] <- 'S083'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-097'] <- 'S097'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-038'] <- 'S038'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-075'] <- 'S075'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-973'] <- 'S973'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-015'] <- 'S015'

# Batch 4

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-522'] <- 'L522'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-528'] <- 'L528'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-091'] <- 'S091'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-907'] <- 'S907'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-524'] <- 'L524'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-520'] <- 'L520'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-305'] <- 'S305'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-535'] <- 'L535'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-539'] <- 'L539'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-534'] <- 'L534'

# Batch 5

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-092'] <- 'L092'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-100'] <- 'S100'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-103'] <- 'S103'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-099'] <- 'S099'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-105'] <- 'S105'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-111'] <- 'S111'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-098'] <- 'S098'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-101'] <- 'S101'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-108'] <- 'S108'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-106'] <- 'S106'

# Batch 6

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-119'] <- 'S119'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-965'] <- 'S965'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-104'] <- 'S104'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-110'] <- 'S110'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-112'] <- 'S112'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-114'] <- 'S114'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-115'] <- 'S115'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-117'] <- 'S117'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-120'] <- 'S120'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-503'] <- 'S503'

# Batch 7

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-303'] <- 'S303'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-931'] <- 'S931'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-037'] <- 'S037'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-058'] <- 'S058'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-065'] <- 'S065'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-066'] <- 'S066'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-070'] <- 'S070'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-076'] <- 'S076'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-989'] <- 'S989'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-071'] <- 'S071'

# Batch 8

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-023'] <- 'S023'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-028'] <- 'S028'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-032'] <- 'S032'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-079'] <- 'S079'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-302'] <- 'S302'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-308'] <- 'S308'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-310'] <- 'S310'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-903'] <- 'S903'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-945'] <- 'S945'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-960'] <- 'S960'

# Batch 9

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-009'] <- 'S009'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-030'] <- 'S030'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-040'] <- 'S040'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-048'] <- 'S048'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-053'] <- 'S053'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-062'] <- 'S062'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-063'] <- 'S063'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-073'] <- 'S073'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-085'] <- 'S085'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-092'] <- 'S092'

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 's053'] <- 'S053'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 's063'] <- 'S063'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 's048'] <- 'S048'

# Batch 10

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-036'] <- 'S036'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-042'] <- 'S042'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-046'] <- 'S046'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-051'] <- 'S051'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-055'] <- 'S055'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-089'] <- 'S089'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-093'] <- 'S093'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-314'] <- 'S314'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-315'] <- 'S315'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-982'] <- 'S982'

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 's093'] <- 'S093'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 's982'] <- 'S982'

# Batch 11

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-001'] <- 'S001'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-007'] <- 'S007'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-008'] <- 'S008'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-307'] <- 'S307'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-505'] <- 'S505'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-940'] <- 'S940'

# Batch 12

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-021'] <- 'S021'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-022'] <- 'S022'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-026'] <- 'S026'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-034'] <- 'S034'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-045'] <- 'S045'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-061'] <- 'S061'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-069'] <- 'S069'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-096'] <- 'S096'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-507'] <- 'S507'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-508'] <- 'S508'

# Batch 13

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-003'] <- 'S003'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-006'] <- 'S006'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-014'] <- 'S014'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-044'] <- 'S044'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-047'] <- 'S047'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-054'] <- 'S054'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-312'] <- 'S312'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-316'] <- 'S316'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-502'] <- 'S502'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-935'] <- 'S935'

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 's054'] <- 'S054'

# Batch 14

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-049'] <- 'S049'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-004'] <- 'S004'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-010'] <- 'S010'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-011'] <- 'S011'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-013'] <- 'S013'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-017'] <- 'S017'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-043'] <- 'S043'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-086'] <- 'S086'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-311'] <- 'S311'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-506'] <- 'S506'

# Batch 15

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-005'] <- 'S005'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-018'] <- 'S018'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-020'] <- 'S020'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-024'] <- 'S024'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-041'] <- 'S041'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-309'] <- 'S309'
combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'S-501'] <- 'S501'

# Batch L527

combinedtrees$`Plot#`[combinedtrees$`Plot#` == 'L-527'] <- 'L527'

# remove duplicate plots from plot center data frames

plotcenters_coordinates <- plotcenters_coordinates[-c(104, 105, 106, 107),]

plotcentersUTM10N_coordinates <- plotcentersUTM10N_coordinates[-c(104, 105, 106, 107),]

# merge into new data frames

allplotcoordinates <- full_join(plotcenters_coordinates, plotcentersUTM10N_coordinates, by="Plot#") 

# merge into tree data frame

combinedtrees <- full_join(combinedtrees, allplotcoordinates, by="Plot#")

#### Tree coordinates ####

# Make new columns for tree coordinates

combinedtrees$`TreeLongitudeUTM10N` <- 0

combinedtrees$`TreeLatitudeUTM10N` <- 0

# Calculate tree coordinates in UTMs 

# Add declination to Azimuth!! I am using a declination from the middle of the survey area (Latitude:	39° 33' 37" N, Longitude:	120° 49' 43" W) from the middle of the survey dates (2023-10-22):	13° 10' E  ± 0° 22'  changing by  0° 5' W per year. 13° 10' is 13.166667°. We need to subtract this declination from the magnetic azimuths to get the true azimuths. Whew!

combinedtrees$`Declination` <- '13.166667'

combinedtrees$`TrueAzimuth` <- '0'

combinedtrees$Declination <- as.numeric(combinedtrees$Declination)

combinedtrees$TrueAzimuth <- as.numeric(combinedtrees$TrueAzimuth)

for(i in 1:nrow(combinedtrees)) {
    combinedtrees[i,]$`TrueAzimuth` = (combinedtrees[i,]$Azimuth) + (combinedtrees[i,]$Declination)
}

# Longitude UTM10N

for(i in 1:nrow(combinedtrees)) {
    combinedtrees[i,]$`TreeLongitudeUTM10N` = (combinedtrees[i,]$PlotLongitudeUTM10N) + ((sin(deg2rad(combinedtrees[i,]$TrueAzimuth))) * ((combinedtrees[i,]$`All Horizontal Distances (m)`)))
}

# Latitude UTM10N

for(i in 1:nrow(combinedtrees)) {
    combinedtrees[i,]$`TreeLatitudeUTM10N` = (combinedtrees[i,]$PlotLatitudeUTM10N) + ((cos(deg2rad(combinedtrees[i,]$TrueAzimuth))) * ((combinedtrees[i,]$`All Horizontal Distances (m)`)))
}

# Convert to WGS84 Coordinates

# give trees ID numbers

combinedtrees = combinedtrees %>% mutate(tree_id = 1:nrow(combinedtrees))

# remove trees without lat/lon --> this happens when there is neither 1. no horizontal distance nor 2. no percent slope/slope distance, so a horizontal distance cannot be calculated

# fortunately there are only six trees in this whole dataset that don't have a horizontal distance, five in batch 2 (an ABCO in plot L536, an ABCO in plot L502, two LIDE3 and one ACMA3 in L529) and one in batch 4 (a PILA in plot L535)

combinedtrees <- combinedtrees %>% drop_na(`TreeLongitudeUTM10N`)

# make a spatial data frame

treecoordinatesconversion <- data.frame(combinedtrees$tree_id, combinedtrees$TreeLongitudeUTM10N, combinedtrees$TreeLatitudeUTM10N) 

treecoordinatesconversion <- treecoordinatesconversion %>% rename (tree_id=combinedtrees.tree_id, TreeLongitudeUTM10N=combinedtrees.TreeLongitudeUTM10N, TreeLatitudeUTM10N=combinedtrees.TreeLatitudeUTM10N)

treecoordinatesconversion <- st_as_sf(treecoordinatesconversion, coords = c("TreeLongitudeUTM10N", "TreeLatitudeUTM10N"), crs = 32610, remove=F)

# change the CRS

treecoordinatesconversion <- treecoordinatesconversion %>% st_transform(4326)

# save this in case it's helpful later

st_write(treecoordinatesconversion, data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.IRI.Compiled.Data\\finaltreecoordinates.kml"),delete_dsn=TRUE)

# extract the geometry features into columns of lat and long WGS84, then merge back into the tree data frame

treecoordinatesconversionWGS84 <- data.frame(treecoordinatesconversion$tree_id, st_coordinates(treecoordinatesconversion[,1], st_coordinates(treecoordinatesconversion[,2]))) 

treecoordinatesconversionWGS84 <- treecoordinatesconversionWGS84 %>% rename (tree_id=treecoordinatesconversion.tree_id, TreeLongitudeWGS84=X, TreeLatitudeWGS84=Y)

combinedtrees <- full_join (combinedtrees, treecoordinatesconversionWGS84, by="tree_id")

```

Convert height from feet to meters

```{r}

# add new column

combinedtrees$`Height (m)` <- 0

# calculate 

combinedtrees$`Height (m)` = combinedtrees$`Height (feet)` * 0.3048

combinedtrees$`Height (m)`[is.na(combinedtrees$`Height (m)`)] <- ''

```


Convert DBH from inches to centimeters

```{r}

# add new column

combinedtrees$`DBH (cm)` <- 0

# calculate 

combinedtrees$`DBH (cm)` = combinedtrees$`DBH (inches)` * 2.54

```

Make sure all species codes are correct

```{r}

# correcting typos in species codes

combinedtrees$`Species`[combinedtrees$`Species` == '112'] <- '122' # 112 is pinus engelmannii, almost certainly a typo, PIPO is 122

combinedtrees$`Species`[combinedtrees$`Species` == '212'] <- '122' # 212 is giant sequoia, almost certainly a typo, PIPO is 122

combinedtrees$`Species`[combinedtrees$`Species` == '188'] <- '818'# another typo I'm guessing at the right answer for-- 188 isn't a code for anything, QUKE is 818

combinedtrees$`Species`[combinedtrees$`Species` == '816'] <- '818' # 816 is quercus ilicifolia which only grows in the eastern US so almost certainly a typo. this is a guess. QUKE is 818

combinedtrees$`Species`[combinedtrees$`Species` == '9'] <- '000' # unknown

combinedtrees$`Species`[combinedtrees$`Species` == '999'] <- '000' # unknown

combinedtrees$`Species`[is.na(combinedtrees$`Species`)] <- '000' # unknown

```

Growth form

```{r}

combinedtrees$`growth_form` <- ''

combinedtrees$`growth_form`[combinedtrees$`Species` == '122'] <- 'tree'  #'PIPO'
combinedtrees$`growth_form`[combinedtrees$`Species` == '15'] <- 'tree'  #'ABCO'
combinedtrees$`growth_form`[combinedtrees$`Species` == '20'] <- 'tree'  # 'ABMA'
combinedtrees$`growth_form`[combinedtrees$`Species` == '117'] <- 'tree' # 'PILA'
combinedtrees$`growth_form`[combinedtrees$`Species` == '101'] <- 'tree' # 'PIAL'
combinedtrees$`growth_form`[combinedtrees$`Species` == '119'] <- 'tree' # 'PIMO3'
combinedtrees$`growth_form`[combinedtrees$`Species` == '108'] <- 'tree' # 'PICOL'
combinedtrees$`growth_form`[combinedtrees$`Species` == '81'] <- 'tree' # 'CADE27'
combinedtrees$`growth_form`[combinedtrees$`Species` == '202'] <- 'tree' # 'PSME'
combinedtrees$`growth_form`[combinedtrees$`Species` == '127'] <- 'tree' # 'PISA2'
combinedtrees$`growth_form`[combinedtrees$`Species` == '116'] <- 'tree' # 'PIJE'
combinedtrees$`growth_form`[combinedtrees$`Species` == '103'] <- 'tree' # 'PIAT'
combinedtrees$`growth_form`[combinedtrees$`Species` == '361'] <- 'tree' # 'ARME'
combinedtrees$`growth_form`[combinedtrees$`Species` == '631'] <- 'intermediate' # 'LIDE3'
combinedtrees$`growth_form`[combinedtrees$`Species` == '312'] <- 'tree' # 'ACMA3' # Bigleaf Maple
combinedtrees$`growth_form`[combinedtrees$`Species` == '981'] <- 'intermediate' # 'UMCA'
combinedtrees$`growth_form`[combinedtrees$`Species` == '333'] <- 'tree' # 'AECA'
combinedtrees$`growth_form`[combinedtrees$`Species` == '805'] <- 'intermediate' # 'QUCH2'
combinedtrees$`growth_form`[combinedtrees$`Species` == '807'] <- 'tree' # 'QUDO'
combinedtrees$`growth_form`[combinedtrees$`Species` == '818'] <- 'tree' # 'QUKE'
combinedtrees$`growth_form`[combinedtrees$`Species` == '839'] <- 'intermediate' # 'QUWI2'
combinedtrees$`growth_form`[combinedtrees$`Species` == '64'] <- 'tree' # 'JUOC'
combinedtrees$`growth_form`[combinedtrees$`Species` == '768'] <- 'intermediate' # 'PREM'
combinedtrees$`growth_form`[combinedtrees$`Species` == '21'] <- 'tree' # 'ABMAS' # Shasta Red Fir
combinedtrees$`growth_form`[combinedtrees$`Species` == '313'] <- 'tree' # 'ACNE2' # Box Elder
combinedtrees$`growth_form`[combinedtrees$`Species` == '492'] <- 'intermediate' # 'CONU4'
combinedtrees$`growth_form`[combinedtrees$`Species` == '000'] <- 'intermediate' # 'Unknown'
combinedtrees$`growth_form`[combinedtrees$`Species` == '231'] <- 'intermediate' # 'TABR2' #Pacific yew
```

Live/dead

```{r}


combinedtrees$`Live/Dead`[combinedtrees$`Live/Dead` == '1'] <- 'L'

combinedtrees$`Live/Dead`[combinedtrees$`Live/Dead` == '11'] <- 'L' # an obvious typo

combinedtrees$`Live/Dead`[combinedtrees$`Live/Dead` == '2'] <- 'D'

combinedtrees$`Live/Dead`[combinedtrees$`Live/Dead` == '18.2'] <- ''

combinedtrees$`Live/Dead`[combinedtrees$`Live/Dead` == '3'] <- ''

```

Canopy position

```{r}

# getting rid of NAs and typos

combinedtrees$`CanopyPosition`[is.na(combinedtrees$`CanopyPosition`)] <- ''

combinedtrees$`CanopyPosition`[combinedtrees$`Species` == '35'] <- '' 

combinedtrees$`CanopyPosition`[combinedtrees$`Species` == '15'] <- '' 

combinedtrees$`CanopyPosition`[combinedtrees$`Species` == '12'] <- '' 

```

Overhead visibility

```{r}

# add a new column for ohvis

combinedtrees <- combinedtrees %>%
  add_column(ohvis = "")

combinedtrees$CanopyPosition <- as.numeric(combinedtrees$CanopyPosition)

# calculate ohvis from canopy position

combinedtrees$ohvis <- case_match(
  combinedtrees$CanopyPosition,
  c(1, 2, 3) ~ "TRUE",
  c(4, 5) ~ "FALSE",
  .default = ""
)

```

Percent crown is crown_ratio_compacted

```{r}

combinedtrees <- combinedtrees %>% rename (crown_ratio_compacted='%Crown')

combinedtrees$`crown_ratio_compacted`[is.na(combinedtrees$`crown_ratio_compacted`)] <- ''

```

Live tree defects/damage codes

```{r}

combinedtrees$`Live Tree Defects`[combinedtrees$`Live Tree Defects` == '0'] <- '' 

combinedtrees$`Live Tree Defects`[combinedtrees$`Live Tree Defects` == ']'] <- '' 

combinedtrees$`Live Tree Defects`[combinedtrees$`Live Tree Defects` == '3'] <- ''

combinedtrees$`Live Tree Defects`[is.na(combinedtrees$`Live Tree Defects`)] <- ''

```

Snag decay class

```{r}

combinedtrees$`Snag Decay Class`[is.na(combinedtrees$`Snag Decay Class`)] <- ''

combinedtrees$`Snag Decay Class`[combinedtrees$`Snag Decay Class` == '0'] <- ''

```

Distance to pith (remember, distance to face is the all horizontal distances (m) measurement!)

```{r}

combinedtrees <- combinedtrees %>%
  add_column(horiz_distance_pith = "")

# calculate

combinedtrees$horiz_distance_pith = combinedtrees$'All Horizontal Distances (m)' + (0.5 * 0.01 * combinedtrees$'DBH (cm)')
```

Removes NAs from notes

```{r}

combinedtrees$`Notes`[is.na(combinedtrees$`Notes`)] <- ''

```

Export CSV with all data

```{r}

write.csv(combinedtrees, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.IRI.Compiled.Data\\20240618_IRItoOFO_ALLplots.csv")

```

Drop rows without an OFO plot ID number-- this will get rid of all the trees in the small plots that aren't being incorporated in this dataset

```{r}

largeplots <- combinedtrees[combinedtrees$ofo_plot_id !=0,]

```

Export CSV with only large plot data

```{r}

write.csv(largeplots, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.IRI.Compiled.Data\\20240618_IRItoOFO_largeplots.csv")

```

Drop large plots to save CSV of just small plots

```{r}

smallplots <- combinedtrees[combinedtrees$ofo_plot_id == 0,]

write.csv(smallplots, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\TNC.IRI.Compiled.Data\\20240618_IRItoOFO_smallplots.csv")

```


## ALTERATIONS MADE IN EXCEL

Manually moved damage codes into new columns


