---
title: "20231004_OSU_stem_data"
author: "Emily Marie Purvis"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ALTERATIONS MADE IN R


```{r, echo=FALSE}

# Load libraries 

library(tidyverse)
library(readxl)
library(sf)
library(pracma)

```

Load data

```{r}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load data

Johnston14 = read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0002\\data\\stem_data.xls" ),sheet=1,col_names = TRUE,skip=2)

Johnston15 = read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0002\\data\\stem_data.xls" ),sheet=2,col_names = TRUE,skip=2)

Johnston11 = read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0002\\data\\stem_data.xls" ),sheet=3,col_names = TRUE,skip=2)

Johnston10 = read_excel(data("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0002\\data\\stem_data.xls" ),sheet=4,col_names = TRUE,skip=2)

# Remove trees with NA distance and/or azimuth

Johnston14 <- Johnston14 %>% drop_na (AZIMUTH) %>% drop_na (`DIST (m)`)

Johnston15 <- Johnston15 %>% drop_na (AZIMUTH) %>% drop_na (`DIST (m)`)

Johnston11 <- Johnston11 %>% drop_na (AZIMUTH) %>% drop_na (`DIST (m)`)

Johnston10 <- Johnston10 %>% drop_na (AZIMUTH) %>% drop_na (`DIST (m)`)

```


```{r}
#### create a dataframe of subplot centers and merge it to the plot dataframes ####

# Johnston14

# manually calculated the last 4 subplot coordinates. the first 2 were dd.dddd, the last four were dd mm.mmmm.

Johnston14_subplot_centers <- matrix(c('NW4', -122.447291, 43.597599, 'NE3', -122.446917, 43.597579, 'SW2', -122.444723, 43.597352, 'SE1', -122.446992, 43.597423, 'NW2', -122.447283, 43.59783, 'NE1', -122.44688, 43.597892), ncol=3, byrow=TRUE)

colnames(Johnston14_subplot_centers) <- c('SUBPLOT','PlotCenterEasting','PlotCenterNorthing')

Johnston14_subplot_centers <- as.data.frame.matrix(Johnston14_subplot_centers)

# make into spatial dataframe

Johnston14_subplot_centers_sp <- st_as_sf(Johnston14_subplot_centers, coords = c("PlotCenterEasting", "PlotCenterNorthing"), crs = 4326, remove=F)

# change CRS to UTM

Johnston14_subplot_centers_sp_UTM10N <- st_transform (Johnston14_subplot_centers_sp, 32610)

# Extract UTM coordinates

Johnston14_UTM10N_plotcoords <- data.frame(Johnston14_subplot_centers_sp$SUBPLOT, st_coordinates(Johnston14_subplot_centers_sp_UTM10N[,1], st_coordinates(Johnston14_subplot_centers_sp_UTM10N[,2]))) %>% rename (SUBPLOT=Johnston14_subplot_centers_sp.SUBPLOT, EastingUTM10N=X, NorthingUTM10N=Y)

Johnston14_subplot_centers_sp_UTM10N <- full_join(Johnston14_subplot_centers_sp_UTM10N, Johnston14_UTM10N_plotcoords, by="SUBPLOT")

# Merge UTM coordinates back into tree data dataframe

Johnston14 = full_join(Johnston14,st_drop_geometry(Johnston14_subplot_centers_sp_UTM10N),by="SUBPLOT")

# Johnston15

# doing this one a bit differently because all the subplot coordinates are in dd mm.mmmm

Johnston15_subplot_centers <- matrix(c('S2', -122, -24.0664, 43, 35.8758, 'E1', -122, -24.0618, 43, 35.8983, 'W4', -122, -24.0853, 43, 35.8865, 'N3', -122, -24.0726, 43, 35.8902, '6', -122, -24.0992, 43, 35.8932, '5', -122, -24.0931, 43, 35.8992), ncol=5, byrow=TRUE)

colnames(Johnston15_subplot_centers) <- c('SUBPLOT','PlotCenterEastingDegrees', 'PlotCenterEastingMinutes', 'PlotCenterNorthingDegrees', 'PlotCenterNorthingMinutes')

Johnston15_subplot_centers <- as.data.frame.matrix(Johnston15_subplot_centers)

# Calculate into dd.dddd format

Johnston15_subplot_centers$PlotCenterEasting = (as.numeric(Johnston15_subplot_centers$PlotCenterEastingDegrees)) + ((as.numeric(Johnston15_subplot_centers$PlotCenterEastingMinutes))/60)

Johnston15_subplot_centers$PlotCenterNorthing = (as.numeric(Johnston15_subplot_centers$PlotCenterNorthingDegrees)) + ((as.numeric(Johnston15_subplot_centers$PlotCenterNorthingMinutes))/60)

# make into spatial data frame

Johnston15_subplot_centers_sp <- st_as_sf(Johnston15_subplot_centers, coords = c("PlotCenterEasting", "PlotCenterNorthing"), crs = 4326, remove=F)

# change CRS to UTM

Johnston15_subplot_centers_sp_UTM10N <- st_transform (Johnston15_subplot_centers_sp, 32610)

# Extract UTM coordinates

Johnston15_UTM10N_plotcoords <- data.frame(Johnston15_subplot_centers_sp$SUBPLOT, st_coordinates(Johnston15_subplot_centers_sp_UTM10N[,1], st_coordinates(Johnston15_subplot_centers_sp_UTM10N[,2]))) %>% rename (SUBPLOT=Johnston15_subplot_centers_sp.SUBPLOT, EastingUTM10N=X, NorthingUTM10N=Y)

Johnston15_subplot_centers_sp_UTM10N <- full_join(Johnston15_subplot_centers_sp_UTM10N, Johnston15_UTM10N_plotcoords, by="SUBPLOT")

# Merge UTM coordinates back into tree data dataframe

Johnston15 = full_join(Johnston15,st_drop_geometry(Johnston15_subplot_centers_sp_UTM10N),by="SUBPLOT")

# Johnston11

# all the subplot coordinates are in dd mm.mmmm

Johnston11_subplot_centers <- matrix(c('SW2', -122, -23.5526, 43, 33.8795, 'SE1', -122, -23.5373, 43, 33.8782, 'NW4', -122, -23.5600, 43, 33.8777, 'NE3', -122, -23.5377, 43, 33.8937, '6', -122, -23.5545, 43, 33.9010, '5', -122, -23.5342, 43, 33.8985), ncol=5, byrow=TRUE)

colnames(Johnston11_subplot_centers) <- c('SUBPLOT','PlotCenterEastingDegrees', 'PlotCenterEastingMinutes', 'PlotCenterNorthingDegrees', 'PlotCenterNorthingMinutes')

Johnston11_subplot_centers <- as.data.frame.matrix(Johnston11_subplot_centers)

# Calculate into dd.dddd format

Johnston11_subplot_centers$PlotCenterEasting = (as.numeric(Johnston11_subplot_centers$PlotCenterEastingDegrees)) + ((as.numeric(Johnston11_subplot_centers$PlotCenterEastingMinutes))/60)

Johnston11_subplot_centers$PlotCenterNorthing = (as.numeric(Johnston11_subplot_centers$PlotCenterNorthingDegrees)) + ((as.numeric(Johnston11_subplot_centers$PlotCenterNorthingMinutes))/60)

# make into spatial data frame

Johnston11_subplot_centers_sp <- st_as_sf(Johnston11_subplot_centers, coords = c("PlotCenterEasting", "PlotCenterNorthing"), crs = 4326, remove=F)

# change CRS to UTM

Johnston11_subplot_centers_sp_UTM10N <- st_transform (Johnston11_subplot_centers_sp, 32610)

# Extract UTM coordinates

Johnston11_UTM10N_plotcoords <- data.frame(Johnston11_subplot_centers_sp$SUBPLOT, st_coordinates(Johnston11_subplot_centers_sp_UTM10N[,1], st_coordinates(Johnston11_subplot_centers_sp_UTM10N[,2]))) %>% rename (SUBPLOT=Johnston11_subplot_centers_sp.SUBPLOT, EastingUTM10N=X, NorthingUTM10N=Y)

Johnston11_subplot_centers_sp_UTM10N <- full_join(Johnston11_subplot_centers_sp_UTM10N, Johnston11_UTM10N_plotcoords, by="SUBPLOT")

# Merge UTM coordinates back into tree data dataframe

Johnston11 = full_join(Johnston11,st_drop_geometry(Johnston11_subplot_centers_sp_UTM10N),by="SUBPLOT")

# Johnston10

# all the subplot coordinates are in dd mm.mmmm

Johnston10_subplot_centers <- matrix(c('NW1', -122, -26.9530, 43, 33.6189, 'NE2', -122, -26.9322, 43, 33.6183, 'SE4', -122, -26.9307, 43, 33.6112, 'SW3', -122, -26.9527, 43, 33.6072, '5', -122, -26.9485, 43, 33.5955, '6', -122, -26.9319, 43, 33.5941), ncol=5, byrow=TRUE)

colnames(Johnston10_subplot_centers) <- c('SUBPLOT','PlotCenterEastingDegrees', 'PlotCenterEastingMinutes', 'PlotCenterNorthingDegrees', 'PlotCenterNorthingMinutes')

Johnston10_subplot_centers <- as.data.frame.matrix(Johnston10_subplot_centers)

# Calculate into dd.dddd format

Johnston10_subplot_centers$PlotCenterEasting = (as.numeric(Johnston10_subplot_centers$PlotCenterEastingDegrees)) + ((as.numeric(Johnston10_subplot_centers$PlotCenterEastingMinutes))/60)

Johnston10_subplot_centers$PlotCenterNorthing = (as.numeric(Johnston10_subplot_centers$PlotCenterNorthingDegrees)) + ((as.numeric(Johnston10_subplot_centers$PlotCenterNorthingMinutes))/60)

# make into spatial data frame

Johnston10_subplot_centers_sp <- st_as_sf(Johnston10_subplot_centers, coords = c("PlotCenterEasting", "PlotCenterNorthing"), crs = 4326, remove=F)

# change CRS to UTM

Johnston10_subplot_centers_sp_UTM10N <- st_transform (Johnston10_subplot_centers_sp, 32610)

# Extract UTM coordinates

Johnston10_UTM10N_plotcoords <- data.frame(Johnston10_subplot_centers_sp$SUBPLOT, st_coordinates(Johnston10_subplot_centers_sp_UTM10N[,1], st_coordinates(Johnston10_subplot_centers_sp_UTM10N[,2]))) %>% rename (SUBPLOT=Johnston10_subplot_centers_sp.SUBPLOT, EastingUTM10N=X, NorthingUTM10N=Y)

Johnston10_subplot_centers_sp_UTM10N <- full_join(Johnston10_subplot_centers_sp_UTM10N, Johnston10_UTM10N_plotcoords, by="SUBPLOT")

# Merge UTM coordinates back into tree data dataframe

Johnston10 = full_join(Johnston10,st_drop_geometry(Johnston10_subplot_centers_sp_UTM10N),by="SUBPLOT")

```

Finding tree lat and lon

```{r}

# calculate!

Johnston14 = Johnston14 %>%
  mutate(TreeEasting = as.numeric(EastingUTM10N) + sin(deg2rad(AZIMUTH)) * `DIST (m)`,
         TreeNorthing = as.numeric(NorthingUTM10N) + cos(deg2rad(AZIMUTH)) * `DIST (m)`)

Johnston15 = Johnston15 %>%
  mutate(TreeEasting = as.numeric(EastingUTM10N) + sin(deg2rad(AZIMUTH)) * `DIST (m)`,
         TreeNorthing = as.numeric(NorthingUTM10N) + cos(deg2rad(AZIMUTH)) * `DIST (m)`)

Johnston11 = Johnston11 %>%
  mutate(TreeEasting = as.numeric(EastingUTM10N) + sin(deg2rad(AZIMUTH)) * `DIST (m)`,
         TreeNorthing = as.numeric(NorthingUTM10N) + cos(deg2rad(AZIMUTH)) * `DIST (m)`)

Johnston10 = Johnston10 %>%
  mutate(TreeEasting = as.numeric(EastingUTM10N) + sin(deg2rad(AZIMUTH)) * `DIST (m)`,
         TreeNorthing = as.numeric(NorthingUTM10N) + cos(deg2rad(AZIMUTH)) * `DIST (m)`)

# Tree eastings and northings from UTM back to WGS84

Johnston14_sp <- st_as_sf(Johnston14, coords = c("TreeEasting", "TreeNorthing"), crs = 32610, remove=F) 

Johnston14_sp <- st_transform (Johnston14_sp, 4326)

Johnston15_sp <- st_as_sf(Johnston15, coords = c("TreeEasting", "TreeNorthing"), crs = 32610, remove=F) 

Johnston15_sp <- st_transform (Johnston15_sp, 4326)

Johnston11_sp <- st_as_sf(Johnston11, coords = c("TreeEasting", "TreeNorthing"), crs = 32610, remove=F) 

Johnston11_sp <- st_transform (Johnston11_sp, 4326)

Johnston10_sp <- st_as_sf(Johnston10, coords = c("TreeEasting", "TreeNorthing"), crs = 32610, remove=F) 

Johnston10_sp <- st_transform (Johnston10_sp, 4326)

# extract the geometry features into columns of lat and long, then merge back into the spatial data frame

Johnston14_sp <- data.frame(Johnston14_sp$'SERIES', st_coordinates(Johnston14_sp[,1], st_coordinates(Johnston14_sp[,2]))) %>% rename (tree_lon=X, tree_lat=Y, SERIES=Johnston14_sp.SERIES)

Johnston14 <- full_join(Johnston14, Johnston14_sp, by="SERIES")

Johnston15_sp <- data.frame(Johnston15_sp$'SERIES', st_coordinates(Johnston15_sp[,1], st_coordinates(Johnston15_sp[,2]))) %>% rename (tree_lon=X, tree_lat=Y, SERIES=Johnston15_sp.SERIES)

Johnston15 <- full_join(Johnston15, Johnston15_sp, by="SERIES")

Johnston11_sp <- data.frame(Johnston11_sp$'SERIES', st_coordinates(Johnston11_sp[,1], st_coordinates(Johnston11_sp[,2]))) %>% rename (tree_lon=X, tree_lat=Y, SERIES=Johnston11_sp.SERIES)

Johnston11 <- full_join(Johnston11, Johnston11_sp, by="SERIES")

Johnston10_sp <- data.frame(Johnston10_sp$'SERIES', st_coordinates(Johnston10_sp[,1], st_coordinates(Johnston10_sp[,2]))) %>% rename (tree_lon=X, tree_lat=Y, SERIES=Johnston10_sp.SERIES)

Johnston10 <- full_join(Johnston10, Johnston10_sp, by="SERIES")

```

Adding OFO plot_id

```{r}
# add a new column with the right plot ID

Johnston14 <- Johnston14 %>%
  add_column(plot_id_ofo = "0053")

Johnston15 <- Johnston15 %>%
  add_column(plot_id_ofo = "0054")

Johnston11 <- Johnston11 %>%
  add_column(plot_id_ofo = "0055")

Johnston10 <- Johnston10 %>%
  add_column(plot_id_ofo = "0056")

```


OFO subplot_id

```{r}
# add a new column with the right plot ID

Johnston14 <- Johnston14 %>%
  add_column(subplot_id_ofo = "0") 

Johnston14$subplot_id_ofo[Johnston14$SUBPLOT == 'NW4'] <- '0001'
Johnston14$subplot_id_ofo[Johnston14$SUBPLOT == 'NE3'] <- '0002'
Johnston14$subplot_id_ofo[Johnston14$SUBPLOT == 'SW2'] <- '0003'
Johnston14$subplot_id_ofo[Johnston14$SUBPLOT == 'SE1'] <- '0004'
Johnston14$subplot_id_ofo[Johnston14$SUBPLOT == 'NW2'] <- '0005'
Johnston14$subplot_id_ofo[Johnston14$SUBPLOT == 'NE1'] <- '0006'

Johnston15 <- Johnston15 %>%
  add_column(subplot_id_ofo = "0")

Johnston15$subplot_id_ofo[Johnston15$SUBPLOT == 'S2'] <- '0007'
Johnston15$subplot_id_ofo[Johnston15$SUBPLOT == 'E1'] <- '0008'
Johnston15$subplot_id_ofo[Johnston15$SUBPLOT == 'W4'] <- '0009'
Johnston15$subplot_id_ofo[Johnston15$SUBPLOT == 'N3'] <- '0010'
Johnston15$subplot_id_ofo[Johnston15$SUBPLOT == '6'] <- '0011'
Johnston15$subplot_id_ofo[Johnston15$SUBPLOT == '5'] <- '0012'

Johnston11 <- Johnston11 %>%
  add_column(subplot_id_ofo = "0")

Johnston11$subplot_id_ofo[Johnston11$SUBPLOT == 'SW2'] <- '0013'
Johnston11$subplot_id_ofo[Johnston11$SUBPLOT == 'SE1'] <- '0014'
Johnston11$subplot_id_ofo[Johnston11$SUBPLOT == 'NW4'] <- '0015'
Johnston11$subplot_id_ofo[Johnston11$SUBPLOT == 'NE3'] <- '0016'
Johnston11$subplot_id_ofo[Johnston11$SUBPLOT == '6'] <- '0017'
Johnston11$subplot_id_ofo[Johnston11$SUBPLOT == '5'] <- '0018'

Johnston10 <- Johnston10 %>%
  add_column(subplot_id_ofo = "0")

Johnston10$subplot_id_ofo[Johnston10$SUBPLOT == 'NW1'] <- '0019'
Johnston10$subplot_id_ofo[Johnston10$SUBPLOT == 'NE2'] <- '0020'
Johnston10$subplot_id_ofo[Johnston10$SUBPLOT == 'SE4'] <- '0021'
Johnston10$subplot_id_ofo[Johnston10$SUBPLOT == 'SW3'] <- '0022'
Johnston10$subplot_id_ofo[Johnston10$SUBPLOT == '5'] <- '0023'
Johnston10$subplot_id_ofo[Johnston10$SUBPLOT == '6'] <- '0024'

```

DBH is in cm, height is in m, everything in metric! :)

Make a csv that has subplot lat/lon for plots 15, 11, and 10

```{r}

subplotlatlon <- c ("SUBPLOT", "plot_id_ofo", "subplot_id_ofo", "PlotCenterEasting", "PlotCenterNorthing")

subplots15 <- Johnston15[subplotlatlon]
subplots11 <- Johnston11[subplotlatlon]
subplots10 <- Johnston10[subplotlatlon]

subplotlatlon <- bind_rows(subplots15, subplots11, subplots10)

write.csv(subplotlatlon, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\OSU.stem_data\\20231004_OSU_stem_data_subplotlatlon.csv")

```


Going to combine all 4 plot datasets to make the next manipulations easier

```{r}

# first make sure all dataframes have the same columns-- due to discrepancies in how subplot coordinates were recorded, some plots have dd / mm.mmmm columns while others have dd.dddd

treeskeep <- c("SUBPLOT", "SPECIES", "CONDITION        (LI, ST, SN, LO)", "DECAY CLASS (ST, SN only)", "DBH (cm)", "TREE HEIGHT (m)", "HEIGHT TO LIVE FOLIAGE (m)", "HEIGHT TO DEAD BRANCH (m)", "DIST (m)", "AZIMUTH", "CROWN TO BASE HEIGHT (m)", "NOTES", "PlotCenterEasting", "PlotCenterNorthing","tree_lon", "tree_lat", "plot_id_ofo", "subplot_id_ofo", "SERIES" )

# now make sure all data sets have the same column titles and same column formats, there are some discrepancies

Johnston14$PlotCenterEasting <- as.numeric(Johnston14$PlotCenterEasting)
Johnston14$PlotCenterNorthing <- as.numeric(Johnston14$PlotCenterNorthing)

Johnston15 <- Johnston15 %>% rename ("DBH (cm)" = "DBH", "TREE HEIGHT (m)" = "TREE HEIGHT", "HEIGHT TO LIVE FOLIAGE (m)" = "HEIGHT TO LIVE FOLIAGE", "HEIGHT TO DEAD BRANCH (m)" = "HEIGHT TO DEAD BRANCH", "CROWN TO BASE HEIGHT (m)" = "CROWN TO BASE HEIGHT")

Johnston11 <- Johnston11 %>% rename ("DBH (cm)" = "DBH", "TREE HEIGHT (m)" = "TREE HEIGHT", "HEIGHT TO LIVE FOLIAGE (m)" = "HEIGHT TO LIVE FOLIAGE", "HEIGHT TO DEAD BRANCH (m)" = "HEIGHT TO DEAD BRANCH", "CROWN TO BASE HEIGHT (m)" = "CROWN TO BASE HEIGHT")

Johnston10 <- Johnston10  %>% rename ("DBH (cm)" = "DBH", "TREE HEIGHT (m)" = "TREE HEIGHT", "HEIGHT TO LIVE FOLIAGE (m)" = "HEIGHT TO LIVE FOLIAGE", "HEIGHT TO DEAD BRANCH (m)" = "HEIGHT TO DEAD BRANCH", "CROWN TO BASE HEIGHT (m)" = "CROWN TO BASE HEIGHT")

# parse down plot data sheets so they have the same columns

Johnston14 <- Johnston14[treeskeep]
Johnston15 <- Johnston15[treeskeep]
Johnston11 <- Johnston11[treeskeep]
Johnston10 <- Johnston10[treeskeep]

# combine

alldata <-bind_rows (Johnston14, Johnston15, Johnston11, Johnston10)

```

Species

```{r}

alldata$SPECIES[alldata$SPECIES == 'PSME'] <- '202'

alldata$SPECIES[alldata$SPECIES == 'CADE'] <- '081'

alldata$SPECIES[alldata$SPECIES == 'PILA'] <- '117'

alldata$SPECIES[alldata$SPECIES == 'ABGR'] <- '017'

alldata$SPECIES[alldata$SPECIES == 'UNK'] <- '000'

alldata$SPECIES[alldata$SPECIES == 'CHCH'] <- '431'

alldata$SPECIES[alldata$SPECIES == 'TABR'] <- '231'

alldata$SPECIES[alldata$SPECIES == 'TSHE'] <- '263'

alldata$SPECIES[alldata$SPECIES == 'CONU'] <- '492'

alldata$SPECIES[alldata$SPECIES == 'ARME'] <- '361'

alldata$SPECIES[alldata$SPECIES == 'THPL'] <- '242'

alldata$SPECIES[alldata$SPECIES == 'ABAM'] <- '361'

alldata$SPECIES[alldata$SPECIES == 'TSME'] <- '264'

alldata$SPECIES[alldata$SPECIES == 'ABPR'] <- '022'

alldata$SPECIES[alldata$SPECIES == 'ABIES SP.'] <- '010'

alldata$SPECIES[alldata$SPECIES == 'PIPO'] <- '122'

```

Condition

```{r}

# LO is log, ST is stump. We want to delete these from the dataset. Let's rename them as NAs

alldata$`CONDITION        (LI, ST, SN, LO)`[alldata$`CONDITION        (LI, ST, SN, LO)` == 'LO'] <- 'NA'

alldata$`CONDITION        (LI, ST, SN, LO)`[alldata$`CONDITION        (LI, ST, SN, LO)` == 'ST'] <- 'NA'

alldata$`CONDITION        (LI, ST, SN, LO)`[alldata$`CONDITION        (LI, ST, SN, LO)` == "NA"] <- NA

# LI is live, SN is snag

alldata$`CONDITION        (LI, ST, SN, LO)`[alldata$`CONDITION        (LI, ST, SN, LO)` == 'LI'] <- 'L'

alldata$`CONDITION        (LI, ST, SN, LO)`[alldata$`CONDITION        (LI, ST, SN, LO)` == 'SN'] <- 'D'

# delete all NAs in this column

alldata <- alldata[complete.cases(alldata$`CONDITION        (LI, ST, SN, LO)`),]

```

Decay class

```{r}

# We don't want NAs in this column, let's get rid of them

alldata$`DECAY CLASS (ST, SN only)`[alldata$`DECAY CLASS (ST, SN only)` == 'NA'] <- ''

```

Notes

```{r}

# We don't want NAs in this column, let's get rid of them

alldata$NOTES[is.na(alldata$NOTES)] <- ""

```

Calculated distance to pith with distance + 1/2(DBH)

```{r}
# Horizontal distance from plot center to tree surface to horizontal distance from plot center to tree pith 

# add a new column for horizontal distance to pith 

alldata <- alldata %>%
  add_column(horiz_distance_pith = "")

# calculate 

alldata$horiz_distance_pith = alldata$`DIST (m)`+ (0.5 * 0.01 * alldata$`DBH (cm)`)

```


Export CSV

```{r}

write.csv(alldata, "C:\\Users\\emily\\OneDrive\\Desktop\\FOCAL\\field-data-standardization\\contributor_data_to_OFO_schema\\20231004_OSU_stem_data.csv")

```


## ALTERATIONS MADE IN EXCEL

- dd = d + m/60 + s/3600 --> manually calculated lat/lon with this formula in excel for the subplots in Johnston 14 that had dd mm.mmmmm. This is because some were in dd.dddd already.

- All other lat/lon calculations were made in R

- In an email to me (E.M. Purvis), dataset contributor Sven Rodne clarified that "As for the "duplicate trees", those notes are in reference to coring plots that were established at each site, so that is the proper tree inventory and not duplicates in the stem data." I manually deleted all notes about duplicate trees to avoid confusion. In previous stem map creation, I confirmed that all trees labelled "duplicate" in fact had unique spatial coordinates.

- Added growth_form data