---
title: "San Jacinto ForestGEO"
author: "Emily Marie Purvis"
date: "2023-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## DATA MANIPULATED IN R

```{r, echo=FALSE}

# Load libraries

library(tidyverse)
library(sf)
library(stringr)
library(pracma)
```

Load data

```{r}

# Set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

# Load data
trees <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0008\\data\\SJFDP_Census_4ha_full_202230427.csv")

subplots <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0008\\data\\SJFDP_Env_4ha_20230504.csv")

```

We want to delete any stems that aren't main stems-- spatial data was not collected on them. 

```{r}

# Only keep main stems

trees <- subset(trees, Main!="S")

```

Calculate plot centroid

```{r}

# x coordinate centroid is the 16s

# y coordinate centroid is the Fs

# therefore F16 is the centroid of the whole plot-- F16 coordinates taken from the subplot data.frame

```

Add OFO plot ID and subplot IDs


```{r}
# plot ID 

trees <- trees %>%
  add_column(plot_id_ofo = "0081")

# we are defining subplots based on where plants were measured FROM, not which quadrat they reside in. subplots in this plot are therefore overlapping.

# some plants were shot with a rangefinder from quadrat corners. these have an "RF_grid_corner" cell value-- this is the subplot.

# some plants were measured from along the grid line with a tape measure (e.g. not shot from a quadrat corner). the "quad" cell value for these plants is the subplot.

# going to add subplot ID to the subplot datasheet and then merge it into the tree datasheet

# add column

subplots <- subplots %>%
  add_column(subplot_id_ofo = "0")

# sort in descending order

 subplots <- arrange(subplots, Quad)
 
# add subplot ID numbers to each quadrat
 
for(i in 1:nrow(subplots)) {
subplots$subplot_id_ofo[i] <- i + 211
}
 
# change some column names and get rid of some columns in the subplot data sheet to prepare to merge it into the tree data sheet
 
subplots <- subset(subplots, select = c(Quad, Lat, Long, subplot_id_ofo))

subplots <- subplots %>% rename("SubplotLat" = "Lat") %>% rename("SubplotLon" = "Long")

subplots <- subplots %>% rename("quad_measured_from" = "Quad") 

# we need the coordinates in UTMs to do the tree coordinate calculations!

# first, make the subplot data into a spatial dataframe

subplots_sp <- st_as_sf(subplots, coords = c("SubplotLon","SubplotLat"), crs = 4326)

# save the subplot spatial data in case it's useful later

subplots_sp <- subplots_sp %>% st_transform(4326)

#st_write(subplots_sp, data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\San_Jacinto_ForestGeo\\quadrat_corners.gpkg"))

# convert to UTM11N

subplots_sp <- subplots_sp %>% st_transform(32611)

# extract the geometry features into columns of lat and long, then merge back into the spatial data frame

subplots_UTM <- data.frame(subplots_sp$'subplot_id_ofo', st_coordinates(subplots_sp[,1], st_coordinates(subplots_sp[,2]))) 

subplots_UTM <- subplots_UTM %>% rename ('subplot_id_ofo'=subplots_sp.subplot_id_ofo, SubplotLonUTM=X, SubplotLatUTM=Y)

subplots <- full_join(subplots, subplots_UTM, by="subplot_id_ofo")

# great, time to get a "real quad" column in the tree data set so we can match it up with the subplot data set

# add a new column

trees <- trees %>%
  add_column(quad_measured_from = "")

# get the right value in the column

for(i in 1:nrow(trees)) {
  if(is.na(trees$`RF_Grid_Corner`[i])) {
    trees[i,]$quad_measured_from = (trees[i,]$Quadrat)
  }
  else {
    trees[i,]$quad_measured_from = trees[i,]$`RF_Grid_Corner`
  }
}

# merge subplot id

trees <- full_join(trees, subplots, by="quad_measured_from")
```

Tree lat and tree lon

```{r}
# for trees that have distance and azimuth: calculate the coordinates of each tree using the subplot center, horizontal distance, and azimuth 

trees = trees %>%
  mutate(TreeLonUTM = as.numeric(SubplotLonUTM) + sin(deg2rad(AZ)) * HD,
         TreeLatUTM = as.numeric(SubplotLatUTM) + cos(deg2rad(AZ)) * HD)

# for trees that have x and y measurements from quadrat corner:

# used the measure bearing tool in Qgis to calculate an approximate rotation angle for the plot grid relative to N-S

# rotation angle measurements: 36.9, 37.7, 36.24. we'll say the rotation angle is 37 degrees.

# used the x and y to compute the distance (sqrt(x^2 + y^2)) and azimuth (atan(x/y))

# added the rotation to the azimuth

# computed the rotated x and y as though it was a normal distance-azimuth plot

trees = trees %>%
  mutate(HD_computed = (sqrt((trees$Xcoordinates^2) + (trees$Ycoordinates^2))))

trees = trees %>%
  mutate(atanxy = atan((trees$Xcoordinates)/(trees$Ycoordinates)))

for(i in 1:nrow(trees)) {
  if(is.na(trees$`AZ`[i])) {
    trees[i,]$TreeLonUTM = (
      (trees[i,]$SubplotLonUTM) + 
      ((trees[i,]$HD_computed) *
      (sin(trees[i,]$atanxy + deg2rad(37)))))
    trees[i,]$TreeLatUTM = (
      (trees[i,]$SubplotLatUTM) + 
      ((trees[i,]$HD_computed) *
      (cos(trees[i,]$atanxy + deg2rad(37)))))
  }
}

# need to delete random rows of all NAs 

trees <- trees %>% drop_na(`Quadrat`)

```

Convert tree data to spatial data and save

```{r}
# one plant is missing spatial data because it has a horizontal distance but no azimuth, we're deleting it

trees <- trees %>% drop_na(`TreeLonUTM`)

trees_sp <- st_as_sf(trees, coords = c("TreeLonUTM","TreeLatUTM"), crs = 32611)

trees_sp <- trees_sp %>% st_transform(4326)

# st_write(trees_sp, data("C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\San_Jacinto_ForestGeo\\trees.gpkg"))
```

Convert tree UTMs to WGS84 coordinates and merge back into the dataset

```{r}
# give trees ID numbers

trees_sp = trees_sp %>% mutate(tree_id = 1:nrow(trees_sp))

trees = trees %>% mutate(tree_id = 1:nrow(trees))

# extract the geometry features into columns of lat and long, then merge back into the spatial data frame

trees_WGS <- data.frame(trees_sp$'tree_id', st_coordinates(trees_sp[,1], st_coordinates(trees_sp[,2]))) 

trees_WGS <- trees_WGS %>% rename ('tree_id'=trees_sp.tree_id, TreeLonWGS=X, TreeLatWGS=Y)

trees <- full_join(trees, trees_WGS, by="tree_id")

```

Tree DBH is in cm :)

Tree species

```{r}

# putting FIA species codes in the pre-existing "Code" column, which I'm renaming to "Species Code"

trees$Code[trees$Code == 'ABICON'] <- '015'
trees$Code[trees$Code == 'CALDEC'] <- '081'
trees$Code[trees$Code == 'PINLAM'] <- '117'
trees$Code[trees$Code == 'PINPON'] <- '122'
trees$Code[trees$Code == 'PINJEF'] <- '116'
trees$Code[trees$Code == 'QUEKEL'] <- '818'
trees$Code[trees$Code == 'UNKN'] <- '000'
trees$Code[trees$Code == 'QUEWIS'] <- '839'
trees$Code[trees$Code == 'QUECHR'] <- '805'
trees$Code[trees$Code == 'PINCOU'] <- '109'
trees$Code[trees$Code == 'QUEXMO'] <- '899' # the Quercus x moreha (Q. kelloggii x wislizeni)

# shrubs without numeric codes left with USDA alphabetical codes; all were double checked against USDA site before continuing

# rename this column so it's more clear what it describes

trees <- trees %>% 
       rename("SpeciesCode" = "Code")
```

Tree, shrub, intermediate

```{r}
# add column

trees <- trees %>%
  add_column(growth_form = "")

trees$growth_form[trees$SpeciesCode == '015'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '081'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '117'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '122'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '116'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '818'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '000'] <- 'intermediate'
trees$growth_form[trees$SpeciesCode == '839'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '805'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '109'] <- 'tree'
trees$growth_form[trees$SpeciesCode == '899'] <- 'tree'

trees$growth_form[trees$SpeciesCode == 'FRACAL'] <- 'intermediate'

trees$growth_form[trees$SpeciesCode == 'ARCGLA'] <- 'shrub'
trees$growth_form[trees$SpeciesCode == 'ARCPRI'] <- 'shrub'
trees$growth_form[trees$SpeciesCode == 'ARCPUN'] <- 'shrub'
trees$growth_form[trees$SpeciesCode == 'CEAPAL'] <- 'shrub'
trees$growth_form[trees$SpeciesCode == 'LONSUB'] <- 'shrub'
trees$growth_form[trees$SpeciesCode == 'RHOOCC'] <- 'shrub'

```

Status Live/Dead

```{r}

# codes to L and D

trees$Status[trees$Status == 'DS'] <- 'D'

trees$Status[trees$Status == 'LI'] <- 'L'

# remove trees that are dead and down from the data set-- we only want standing individuals

trees <- subset(trees, Status!="Dt")

```

Distance to pith from wherever the tree was shot from

```{r}

# add a new column for horizontal distance to pith

trees <- trees %>%
  add_column(horiz_distance_pith = "")

# calculate

for(i in 1:nrow(trees)) {
  if(is.na(trees$`AZ`[i])) {
    trees[i,]$horiz_distance_pith = trees[i,]$HD_computed + (0.5 * 0.01 * trees[i,]$DBH)
  }
  else {
    trees[i,]$horiz_distance_pith = trees[i,]$HD + (0.5 * 0.01 * trees[i,]$DBH)
  }
}

```

Compute distance and azimuth for all plants

```{r}

# distance

trees <- trees %>%
  add_column(HD_combined = "")

for(i in 1:nrow(trees)) {
  if(is.na(trees$`AZ`[i])) {
    trees[i,]$HD_combined = trees[i,]$HD_computed
  }
  else {
    trees[i,]$HD_combined = trees[i,]$HD 
  }
}

# azimuth

trees <- trees %>%
  add_column(AZ_combined = "")

for(i in 1:nrow(trees)) {
  if(is.na(trees$`AZ`[i])) {
    trees[i,]$AZ_combined = (rad2deg(trees[i,]$atanxy) + 37)
  }
  else {
    trees[i,]$AZ_combined = trees[i,]$AZ 
  }
}

```

Delete any stems measured below 1.3m (regular DBH height)

```{r}

# remove stems measured below DBH

trees<- subset(trees, POM>=1.3)

```

Damage codes

BB	Bark beetle
CD	Crown >75% damaged
FS	Fire scar
L	Leaning > 30 degrees
MS	Mistletoe
PD	Pest damage
SBAD	Stem broken above DBH
SBBD	Stem broken below DBH
T	Topped

```{r}

trees$Codes <- str_replace(trees$Codes, "TG", "") #Tagged out of sequence
trees$Codes <- str_replace(trees$Codes, "RS", "") #Resprout
trees$Codes <- str_replace(trees$Codes, "I", "") #Stem irregular where measured
trees$Codes <- str_replace(trees$Codes, "QS", "") #ID uncertain
trees$Codes <- str_replace(trees$Codes, "SBBD", "90001")
trees$Codes <- str_replace(trees$Codes, "BB", "11000")
trees$Codes <- str_replace(trees$Codes, "CD", "90000") # unclear what kind of crown damage this is referring to (biotic or abiotic, mechanical damage, flagging...it's a mystery) so I'm labelling this as "other damages"
trees$Codes <- str_replace(trees$Codes, "FS", "30000")
trees$Codes <- str_replace(trees$Codes, "FL", "") #Flowering
trees$Codes <- str_replace(trees$Codes, "FR", "") #Fruiting
trees$Codes <- str_replace(trees$Codes, "FS", "") #Fire scar
trees$Codes <- str_replace(trees$Codes, "L", "90006")
trees$Codes <- str_replace(trees$Codes, "l", "90006")
trees$Codes <- str_replace(trees$Codes, "MS", "23000") # general code for parisitic/epiphytic plants, there are so many mistletoes 
trees$Codes <- str_replace(trees$Codes, "PD", "10000") # general code for insect damage
trees$Codes <- str_replace(trees$Codes, "SBAD", "90001")
trees$Codes <- str_replace(trees$Codes, "T", "60001")
trees$Codes <- str_replace(trees$Codes, "M", "") #Main stem
trees$Codes <- str_replace(trees$Codes, "A", "") #Alternate POM
trees$Codes <- str_replace(trees$Codes, "B", "") #Buttress
trees$Codes <- str_replace(trees$Codes, "S", "") #Secondary stem
trees$Codes <- str_replace(trees$Codes, "PRO", "") #Problem
trees$Codes <- str_replace(trees$Codes, "P", "") #Prostrate
trees$Codes <- str_replace(trees$Codes, "N", "") #??? not in metadata. There's not anything that contains N. 

```

Export

```{r}

write.csv(trees, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\San_Jacinto_ForestGeo\\20240423_SanJacintoForestGeo.csv")

```


## ALTERATIONS MADE IN EXCEL

Manually separated damage codes into damage code columns 

