---
title: "20240620_NEON_SJER_SOAP_TEAK"
author: "Emily Marie Purvis"
date: "2024-06-21"
output: html_document
---

## DEAR READER

You will need to insert your own API token in the for loops below, as mine is a secret. Check out these sites for info on using the NEON API (https://www.neonscience.org/resources/learning-hub/tutorials/neon-api-usage) and rate limits (https://data.neonscience.org/data-api/rate-limiting/)

## ALTERATIONS MADE IN R

Libraries 

```{r, echo=FALSE}

# Load libraries 

library(tidyverse)
library(sf)
library(neonUtilities)
library(geoNEON)
library(neonOS)
library(lubridate) # to sort data by collection date
library(httr) # to access the NEON API
library(jsonlite) # to process the NEON API data

```

Find intersection of OFO drone flight footprints and NEON plots from the SJER, SOAP, and TEAK sites

```{r}

#### load OFO data

OFO_flight_footprints <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\FOCAL_flight_polys.gpkg")
  
#### load NEON data and make plot polygons

# get the data
vst <- loadByProduct(dpID="DP1.10098.001", # the "data product ID" number for veg structure data from the NEON catalogue
                     site=c("SJER","SOAP","TEAK")) 

# put everything into the global environment
list2env(vst, .GlobalEnv)

# plot location data is in the vst_perplotperyear data.frame. the latitude and longitude represent the centroid of the sampling plot. tower plots are 40 m x 40 m, and distributed plots are 20 m x 20 m. 

# create spatial object of plot locations

NEON_plots <- st_as_sf(vst_perplotperyear, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)

# export for preliminary examination

# st_write(NEON_plots, "C:\\Users\\emily\\Desktop\\NEONplotcentercoords.gpkg")

# that sf object just has plot centroids only. let's make them into polygons. we can use st_buffer to buffer out a square of the appropriate size. since the plots are 40 m x 40 m or 20 m x 20 m, we need to change the crs of NEON_plots to UTMs so that the function will buffer out in meters. after we can change it back to WGS84 or whatever else we want.

# break data frame into the two different plot sizes

NEON_plots_tower <- NEON_plots[NEON_plots$plotType == "tower",]

NEON_plots_distributed <- NEON_plots[NEON_plots$plotType == "distributed",]

# change crs to UTM 11N

NEON_plots_tower <- st_transform(NEON_plots_tower, crs=32611)

NEON_plots_distributed <- st_transform(NEON_plots_distributed, crs=32611)

# add buffers

NEON_plots_tower <- st_buffer(NEON_plots_tower, dist = 40, endCapStyle = "SQUARE")

NEON_plots_distributed <- st_buffer(NEON_plots_distributed, dist = 20, endCapStyle = "SQUARE")

# join the two sf objects back together

NEON_plots_join <- rbind (NEON_plots_tower, NEON_plots_distributed)

# change crs back to WGS84

NEON_plots_join <- st_transform(NEON_plots_join, crs=4326)

# export for preliminary examination

# st_write(NEON_plots_join, "C:\\Users\\emily\\Desktop\\NEONplotpolygons.gpkg")

# nice we did it!!!

#### Find NEON plot polygons that intersect with drone footprints

# using st_intersection because it will create polygons of the overlap between NEON plots and OFO drone footprints. st_intersects will just return a T/F matrix. I think the polygons will be more helpful-- there may be scenarios where drone footprints only overlap partial NEON plots.

# first make sure the two sf objects have the same crs just in case

st_crs(NEON_plots_join) # 4326
st_crs(OFO_flight_footprints) #3310 AHA

OFO_flight_footprints <- st_transform(OFO_flight_footprints, crs = 4326)

intersection <- st_intersection (NEON_plots_join, OFO_flight_footprints)

# there are so many columns in here it's confusing...going to keep the NEON plot number, the OFO flight number, collection dates, and the geometry

OFONEONintersection <- intersection[,-c(1:2)]

OFONEONintersection <- OFONEONintersection %>% rename(NEONdate = date)

OFONEONintersection <- OFONEONintersection[,-c(2:3)]

OFONEONintersection <- OFONEONintersection[,-c(5:46)]

OFONEONintersection <- OFONEONintersection %>% rename(OFOdataset_id = dataset_id)

OFONEONintersection <- OFONEONintersection[,-c(6:15)]

OFONEONintersection <- OFONEONintersection %>% rename(NEONsiteID = siteID, NEONplotID = plotID, NEONplotType = plotType)

# save these polygons for later

# st_write(OFONEONintersection, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\OFO_NEON_intersection_polygons.gpkg")

# there are a LOT of plot duplicates (e.g. nine SJER_003 plots with identical geometries, three plots for each of three dates)....this is a good reminder to closely examine the tree coordinates for possible duplicates. if trees were all sampled on all of these dates, we just want the most recent dates. let's check it out below!

```

Part 2: finding overlap between WREF plots and OFO drone footprints

```{r}

#### Load FOCAL drone footprints ####

WREF_FOCAL_droneprints <- st_read("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\wref-flight-polys.gpkg")

#### load NEON data and make plot polygons

# get the data
vst <- loadByProduct(dpID="DP1.10098.001", # the "data product ID" number for veg structure data from the NEON catalogue
                     site="WREF") 

# put everything into the global environment
list2env(vst, .GlobalEnv)

# plot location data is in the vst_perplotperyear data.frame. the latitude and longitude represent the centroid of the sampling plot. tower plots are 40 m x 40 m, and distributed plots are 20 m x 20 m. 

# create spatial object of plot locations

WREF_NEON_plots <- st_as_sf(vst_perplotperyear, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)

# that sf object just has plot centroids only. let's make them into polygons. we can use st_buffer to buffer out a square of the appropriate size. since the plots are 40 m x 40 m or 20 m x 20 m, we need to change the crs of NEON_plots to UTMs so that the function will buffer out in meters. after we can change it back to WGS84 or whatever else we want.

# break data frame into the two different plot sizes

NEON_plots_tower <- WREF_NEON_plots[WREF_NEON_plots$plotType == "tower",]

NEON_plots_distributed <- WREF_NEON_plots[WREF_NEON_plots$plotType == "distributed",]

# change crs to UTM 10N

NEON_plots_tower <- st_transform(NEON_plots_tower, crs=32610)

NEON_plots_distributed <- st_transform(NEON_plots_distributed, crs=32610)

# add buffers

NEON_plots_tower <- st_buffer(NEON_plots_tower, dist = 20, endCapStyle = "SQUARE")

NEON_plots_distributed <- st_buffer(NEON_plots_distributed, dist = 10, endCapStyle = "SQUARE")

# join the two sf objects back together

NEON_plots_join <- rbind (NEON_plots_tower, NEON_plots_distributed)

# change crs back to WGS84

NEON_plots_join <- st_transform(NEON_plots_join, crs=4326)

# export for preliminary examination

# st_write(NEON_plots_join, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\WREFplotpolygons_fromcentroids.gpkg")

#### Find NEON plot polygons that intersect with drone footprints

# using st_intersection because it will create polygons of the overlap between NEON plots and OFO drone footprints. st_intersects will just return a T/F matrix. I think the polygons will be more helpful-- there may be scenarios where drone footprints only overlap partial NEON plots.

# first make sure the two sf objects have the same crs just in case

st_crs(NEON_plots_join) # 4326
st_crs(WREF_FOCAL_droneprints) #4326

intersection <- st_intersection (NEON_plots_join, WREF_FOCAL_droneprints)

# there are so many columns in here it's confusing...going to keep the NEON plot number, the OFO flight number, collection dates, and the geometry

OFONEONintersection <- intersection[,-c(1:2)]

OFONEONintersection <- OFONEONintersection %>% rename(NEONdate = date)

OFONEONintersection <- OFONEONintersection[,-c(2:3)]

OFONEONintersection <- OFONEONintersection[,-c(5:46)]

OFONEONintersection <- OFONEONintersection %>% rename(OFOdataset_id = dataset_id)

OFONEONintersection <- OFONEONintersection[,-c(6:15)]

OFONEONintersection <- OFONEONintersection %>% rename(NEONsiteID = siteID, NEONplotID = plotID, NEONplotType = plotType)

# save these polygons for later

# st_write(OFONEONintersection, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\WREF_OFO_NEON_intersection_polygons.gpkg")

# there are a LOT of plot duplicates 

```

Import veg structure data, get tree locations

```{r}

# first remove duplicates from veg structure data

mappingandtagging <- removeDups(data=vst_mappingandtagging,
                          variables=variables_10098)

vst_apparentindividual <- removeDups(data=vst_apparentindividual,
                          variables=variables_10098)

# no duplicates found doesn't mean there aren't duplicated trees across different dates, just ensures there aren't human error duplicates

# run the "getLocTOS" function to get locations of mapped woody individuals

mappingandtagging_locations <- getLocTOS(data = vst_mappingandtagging, dataProd = "vst_mappingandtagging")

# join that data.frame with another that contains more data

join <- joinTableNEON(mappingandtagging_locations,
                   vst_apparentindividual,
                     name1="vst_mappingandtagging",
                     name2="vst_apparentindividual")

# filter this massive dataset to 1. trees, that are 2. mapped, 3. have a stemDiameter ≥ 10 cm (smaller trees were not exhaustively sampled)

join = join |>
  filter(growthForm %in% c("multi-bole tree", "sapling", "single bole tree"), # exclude the "small tree" category because those are trees under 10cm DBH
         !is.na(adjDecimalLatitude),
         (stemDiameter >= 10 & !is.na(stemDiameter)))

# re-running code to include downed trees
#join = join |>
#  filter((growthForm %in% c("multi-bole tree", "sapling", "single bole tree") | #is.na(growthForm)), # exclude the "small tree" category because those are trees #under 10cm DBH
#         !is.na(adjDecimalLatitude),
#         (stemDiameter >= 10  | is.na(stemDiameter)))

# rename some columns, remove some extraneous columns to make this easier to examine

join <- join %>% rename(NEONindividualID = individualID, NEONsiteID = siteID, NEONplotID = plotID)

join <- join[,-2]

join <- join[,-c(4:5)]

join <- join %>% rename(NEONdate = date.x)

join <- join[,-c(5:6)]

join <- join[,-c(7:13)]

join <- join[,-c(9:13)]

join <- join[,-c(10:15)]

join <- join[,-c(16:19)]

join <- join[,-c(45:51)]

join <- join %>% rename(DBH_cm = stemDiameter, height_m = height)

# make into a spatial data frame

join_spatial = st_as_sf(join, coords = c("adjDecimalLongitude", "adjDecimalLatitude"), crs = 4326)

# export to examine

# st_write(join_spatial, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\NEONtrees.gpkg")

```

Remove data from plots outside OFO flight footprints

```{r}

# get a list of plots either partially or completely within OFO flight footprints

overlap_plots <- as.data.frame(unique(OFONEONintersection$NEONplotID)) # wow 50 unique overlap plots!

# remove trees from plots that are not these plots

treesinplots <- join_spatial %>% filter (NEONplotID == 'SOAP_054' | NEONplotID == 'SOAP_047' | NEONplotID == 'SOAP_031' | NEONplotID == 'SOAP_049' | NEONplotID == 'SOAP_056' | NEONplotID == 'SOAP_053' | NEONplotID == 'SOAP_045'
| NEONplotID == 'SJER_059' | NEONplotID == 'SJER_058' | NEONplotID == 'SJER_047' | NEONplotID == 'SJER_054' | NEONplotID == 'SJER_060' | NEONplotID == 'SJER_049' | NEONplotID == 'SJER_045' | NEONplotID == 'SJER_003' | NEONplotID == 'SJER_057' | NEONplotID == 'SJER_046' | NEONplotID == 'SJER_055' | NEONplotID == 'SJER_048' | NEONplotID == 'SJER_061' | NEONplotID == 'SJER_050' | NEONplotID == 'SJER_053' | NEONplotID == 'SJER_063' | NEONplotID == 'SJER_051' | NEONplotID == 'TEAK_062' | NEONplotID == 'TEAK_058' | NEONplotID == 'TEAK_043' | NEONplotID == 'TEAK_056' | NEONplotID == 'TEAK_044' | NEONplotID == 'TEAK_057' | NEONplotID == 'TEAK_047' | NEONplotID == 'TEAK_053' | NEONplotID == 'TEAK_051' | NEONplotID == 'TEAK_046' | NEONplotID == 'SOAP_057' | NEONplotID == 'SOAP_059' | NEONplotID == 'SOAP_051' | NEONplotID == 'SOAP_048' | NEONplotID == 'SOAP_058' | NEONplotID == 'SOAP_043' | NEONplotID == 'SOAP_050' | NEONplotID == 'SOAP_044' | NEONplotID == 'SOAP_052' | NEONplotID == 'SOAP_046' | NEONplotID == 'SOAP_060' | NEONplotID == 'SOAP_055' | NEONplotID == 'SOAP_061' | NEONplotID == 'SOAP_010')

# down to 547 rows

# export to examine

# st_write(treesinplots, "C:\\Users\\emily\\Desktop\\NEONtreesinOFO.gpkg")

```

Repeating this process with the WREF plots

```{r}

# get a list of plots either partially or completely within OFO flight footprints

overlap_plots <- as.data.frame(unique(OFONEONintersection$NEONplotID)) #14 overlap plots

# remove trees from plots that are not these plots

treesinplots <- join_spatial %>% filter (NEONplotID == 'WREF_082' | NEONplotID == 'WREF_074' | NEONplotID == 'WREF_085' | NEONplotID == 'WREF_078' | NEONplotID == 'WREF_084' | NEONplotID == 'WREF_088' | NEONplotID == 'WREF_071'
| NEONplotID == 'WREF_089' | NEONplotID == 'WREF_077' | NEONplotID == 'WREF_083' | NEONplotID == 'WREF_080' | NEONplotID == 'WREF_079' | NEONplotID == 'WREF_073' | NEONplotID == 'WREF_075')

```

Remove duplicated trees that were measured multiple times in multiple years

```{r}

# 547 trees in "treesinplots"

unique_trees <- dplyr::distinct(treesinplots, adjNorthing) # but only 329 unique northings

unique_trees <- dplyr::distinct(treesinplots, adjEasting) # AND 329 unique eastings

unique_trees <- dplyr::distinct(treesinplots, geometry) # And 329 sets of coordinates. strong pattern here.

unique_trees <- dplyr::distinct(treesinplots) # but this returns 547 unique rows 

# HMMMMMMM. So there are 329 unique sets of tree coordinates. But there are 547 unique rows in the original treesinplots sf object. 

# I did use the NEON "removeDups" function earlier. A reasonable question would be "why are there still so many duplicates???" and the answer is that each row in the original data.frames has a "uniqueID" which is a super long string of characters and numbers. So technically there are no duplicates in the original data.frames. But then I removed a bunch of columns that we didn't need, including these "uniqueID" columns. Removing the "uniqueID" columns resulted in a whole lot of identical rows. 

# My plan is to sort the data frame in by date in descending order (so that the most recent rows are on top). Then I'll use dplyr::distinct on the geometry column. When there are duplicated rows, dplyr::distinct keeps the first row. So theoretically we'll keep the most recent set of values for each tree.

# But of course it's more complicated than that. Turns out there are a lot of repeat measurements for single bole trees, from the same sampling date, but confusingly have different heights, dbhs, etc. So in situations where there are multiple unique most recent measurements I'll just choose which one to keep randomly (e.g. whichever one is arbitrarily first in the data.frame). 

# get the unique rows

unique_treesinplots <- dplyr::distinct(treesinplots)

# sort in descending order of dates

unique_treesinplots_bydate <- unique_treesinplots %>% mutate(NEONdate=ymd(NEONdate)) %>% arrange(desc(NEONdate))

# get the 329 trees with unique sets of coordinates

unique_treesinplots_final <- dplyr::distinct(unique_treesinplots_bydate, geometry, .keep_all = TRUE)

# well that was wild, gonna save that for later

# st_write(unique_treesinplots_final, "C:\\Users\\emily\\Desktop\\NEONtreesinOFO_unique.gpkg")

```

Update NEON plot polygons

```{r}

# in plot type = "distributed" (20 m x 20 m plots), the "total plot area sampled for single and multi-bole trees" is 400 m2. so, phew, in these smaller plots all the trees were sampled, and the plot polygons are fine as they are.

# in plot type = "tower" (40 m x 40 m plots) the the "total plot area sampled for single and multi-bole trees" is 800 m2. which is only half of the total plot. but which half????????? must find out and update plot polygons to reflect just this half. 

# first let's filter the appropriate data.frame with the plot info to just the plots that overlap with OFO drone footprints

vst_perplotperyear <- vst_perplotperyear %>% rename(NEONplotID = plotID)

plots <- vst_perplotperyear %>% filter (NEONplotID == 'SOAP_054' | NEONplotID == 'SOAP_047' | NEONplotID == 'SOAP_031' | NEONplotID == 'SOAP_049' | NEONplotID == 'SOAP_056' | NEONplotID == 'SOAP_053' | NEONplotID == 'SOAP_045'
| NEONplotID == 'SJER_059' | NEONplotID == 'SJER_058' | NEONplotID == 'SJER_047' | NEONplotID == 'SJER_054' | NEONplotID == 'SJER_060' | NEONplotID == 'SJER_049' | NEONplotID == 'SJER_045' | NEONplotID == 'SJER_003' | NEONplotID == 'SJER_057' | NEONplotID == 'SJER_046' | NEONplotID == 'SJER_055' | NEONplotID == 'SJER_048' | NEONplotID == 'SJER_061' | NEONplotID == 'SJER_050' | NEONplotID == 'SJER_053' | NEONplotID == 'SJER_063' | NEONplotID == 'SJER_051' | NEONplotID == 'TEAK_049' | NEONplotID == 'TEAK_062' | NEONplotID == 'TEAK_058' | NEONplotID == 'TEAK_043' | NEONplotID == 'TEAK_056' | NEONplotID == 'TEAK_044' | NEONplotID == 'TEAK_057' | NEONplotID == 'TEAK_047' | NEONplotID == 'TEAK_053' | NEONplotID == 'TEAK_051' | NEONplotID == 'TEAK_046' | NEONplotID == 'TEAK_052' | NEONplotID == 'SOAP_057' | NEONplotID == 'SOAP_059' | NEONplotID == 'SOAP_051' | NEONplotID == 'SOAP_048' | NEONplotID == 'SOAP_058' | NEONplotID == 'SOAP_043' | NEONplotID == 'SOAP_050' | NEONplotID == 'SOAP_044' | NEONplotID == 'SOAP_052' | NEONplotID == 'SOAP_046' | NEONplotID == 'SOAP_060' | NEONplotID == 'SOAP_055' | NEONplotID == 'SOAP_061' | NEONplotID == 'SOAP_010')

# oh no.....the subplots sampled are in different orientations in each plot sampled. sometimes they sampled two 400 m2 subplots, sometimes four 100m2 subplots. wtf.

# NEON helpfully sent us which subplots are sampled within each plot! NOTE THAT THESE ARE ONLY THE TOWER PLOTS!!!!!!!!!!!!

subplotssampled <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\D17_Tower_spatial_data.csv")

subplotssampled <- subplotssampled %>% rename(NEONplotID = plotid)

towersubplotssampled <- subplotssampled %>% filter (NEONplotID == 'SOAP_054' | NEONplotID == 'SOAP_047' | NEONplotID == 'SOAP_031' | NEONplotID == 'SOAP_049' | NEONplotID == 'SOAP_056' | NEONplotID == 'SOAP_053' | NEONplotID == 'SOAP_045'
| NEONplotID == 'SJER_059' | NEONplotID == 'SJER_058' | NEONplotID == 'SJER_047' | NEONplotID == 'SJER_054' | NEONplotID == 'SJER_060' | NEONplotID == 'SJER_049' | NEONplotID == 'SJER_045' | NEONplotID == 'SJER_003' | NEONplotID == 'SJER_057' | NEONplotID == 'SJER_046' | NEONplotID == 'SJER_055' | NEONplotID == 'SJER_048' | NEONplotID == 'SJER_061' | NEONplotID == 'SJER_050' | NEONplotID == 'SJER_053' | NEONplotID == 'SJER_063' | NEONplotID == 'SJER_051' | NEONplotID == 'TEAK_062' | NEONplotID == 'TEAK_058' | NEONplotID == 'TEAK_043' | NEONplotID == 'TEAK_056' | NEONplotID == 'TEAK_044' | NEONplotID == 'TEAK_057' | NEONplotID == 'TEAK_047' | NEONplotID == 'TEAK_053' | NEONplotID == 'TEAK_046' | NEONplotID == 'TEAK_052' | NEONplotID == 'SOAP_057' | NEONplotID == 'SOAP_059' | NEONplotID == 'SOAP_051' | NEONplotID == 'SOAP_048' | NEONplotID == 'SOAP_058' | NEONplotID == 'SOAP_043' | NEONplotID == 'SOAP_050' | NEONplotID == 'SOAP_044' | NEONplotID == 'SOAP_052' | NEONplotID == 'SOAP_046' | NEONplotID == 'SOAP_060' | NEONplotID == 'SOAP_055' | NEONplotID == 'SOAP_061' | NEONplotID == 'SOAP_010')

# make each plot polygon into four separate polygons (NE, NW, SE, SW)

# import plot polygons again

NEONplots <- st_read ("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\allD17NEONplotpolygons.gpkg")

# pare down to just the plots that intersect with OFO flight polys

NEONplots <- NEONplots %>% dplyr::rename(NEONplotID = plotID)

NEONplots <- NEONplots %>% filter (NEONplotID == 'SOAP_054' | NEONplotID == 'SOAP_047' | NEONplotID == 'SOAP_031' | NEONplotID == 'SOAP_049' | NEONplotID == 'SOAP_056' | NEONplotID == 'SOAP_053' | NEONplotID == 'SOAP_045'
| NEONplotID == 'SJER_059' | NEONplotID == 'SJER_058' | NEONplotID == 'SJER_047' | NEONplotID == 'SJER_054' | NEONplotID == 'SJER_060' | NEONplotID == 'SJER_049' | NEONplotID == 'SJER_045' | NEONplotID == 'SJER_003' | NEONplotID == 'SJER_057' | NEONplotID == 'SJER_046' | NEONplotID == 'SJER_055' | NEONplotID == 'SJER_048' | NEONplotID == 'SJER_061' | NEONplotID == 'SJER_050' | NEONplotID == 'SJER_053' | NEONplotID == 'SJER_063' | NEONplotID == 'SJER_051' | NEONplotID == 'TEAK_049' | NEONplotID == 'TEAK_062' | NEONplotID == 'TEAK_058' | NEONplotID == 'TEAK_043' | NEONplotID == 'TEAK_056' | NEONplotID == 'TEAK_044' | NEONplotID == 'TEAK_057' | NEONplotID == 'TEAK_047' | NEONplotID == 'TEAK_053' | NEONplotID == 'TEAK_051' | NEONplotID == 'TEAK_046' | NEONplotID == 'TEAK_052' | NEONplotID == 'SOAP_057' | NEONplotID == 'SOAP_059' | NEONplotID == 'SOAP_051' | NEONplotID == 'SOAP_048' | NEONplotID == 'SOAP_058' | NEONplotID == 'SOAP_043' | NEONplotID == 'SOAP_050' | NEONplotID == 'SOAP_044' | NEONplotID == 'SOAP_052' | NEONplotID == 'SOAP_046' | NEONplotID == 'SOAP_060' | NEONplotID == 'SOAP_055' | NEONplotID == 'SOAP_061' | NEONplotID == 'SOAP_010')

# remove extraneous columns and duplicate plot rows

NEONplots <- NEONplots[,-c(46:50)]

NEONplots <- NEONplots[,-c(41:43)]

NEONplots <- NEONplots[,-c(41)]

NEONplots <- NEONplots[,-c(23:25)]

NEONplots <- NEONplots %>% mutate(date=ymd_hms(date)) %>% arrange(desc(date))

NEONplots <- dplyr::distinct(NEONplots, geom, .keep_all = TRUE)

# separate into tower plots (need to further divide into subplots) and distributed plots (no need to further divide)

towerNEONplots <- NEONplots %>% filter(plotType == "tower")

distributedNEONplots <- NEONplots %>% filter(plotType == "distributed")

```

Repeating this process for the WREF plots

```{r}

# filter the appropriate data.frame with the plot info to just the plots that overlap with OFO drone footprints

vst_perplotperyear <- vst_perplotperyear %>% rename(NEONplotID = plotID)

plots <- vst_perplotperyear %>% filter (NEONplotID == 'WREF_082' | NEONplotID == 'WREF_074' | NEONplotID == 'WREF_085' | NEONplotID == 'WREF_078' | NEONplotID == 'WREF_084' | NEONplotID == 'WREF_088' | NEONplotID == 'WREF_071'
| NEONplotID == 'WREF_089' | NEONplotID == 'WREF_077' | NEONplotID == 'WREF_083' | NEONplotID == 'WREF_080' | NEONplotID == 'WREF_079' | NEONplotID == 'WREF_073' | NEONplotID == 'WREF_075')

# in helpful news, there are only tower plots and they all have listed subplots sampled! love this. to separate out the plot IDs in new columns, it's going to be fastest to export the plots data.frame to a csv, separate out the columms manually, and import it back in

# write.csv(plots, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\D16plots.csv")

# okay I did it, going to re-import the new csv

plots <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\D16plots.csv")

```

Make new spatial polys for all plots, plus bonus spatial polys for tower subplots

```{r}

#### Start with distributed plots because it's easier. We currently have the plot centroid for each distributed plot (point ID 41), but we can download the corner locations and connect those into a polygon (point IDs 31, 33, 49, and 51)

## SJER_003.basePlot.vst

# import coords

SJER003_31 <- GET("https://data.neonscience.org/api/v0/locations/SJER_003.basePlot.vst.31")

SJER003_31 <- jsonlite::fromJSON(content(SJER003_31, as="text"))

SJER003_33 <- GET("https://data.neonscience.org/api/v0/locations/SJER_003.basePlot.vst.33")

SJER003_33 <- jsonlite::fromJSON(content(SJER003_33, as="text"))

SJER003_49 <- GET("https://data.neonscience.org/api/v0/locations/SJER_003.basePlot.vst.49")

SJER003_49 <- jsonlite::fromJSON(content(SJER003_49, as="text"))

SJER003_51 <- GET("https://data.neonscience.org/api/v0/locations/SJER_003.basePlot.vst.51")

SJER003_51 <- jsonlite::fromJSON(content(SJER003_51, as="text"))

# make coords into a polygon

SJER003 <- data.frame(
  lon = c(SJER003_31$data$locationDecimalLongitude, SJER003_33$data$locationDecimalLongitude, SJER003_51$data$locationDecimalLongitude, SJER003_49$data$locationDecimalLongitude),
  lat = c(SJER003_31$data$locationDecimalLatitude, SJER003_33$data$locationDecimalLatitude, SJER003_51$data$locationDecimalLatitude, SJER003_49$data$locationDecimalLatitude)
)

SJER003_poly <- SJER003 %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")

SJER003_poly <- SJER003_poly %>% add_column(ofo_plot_id = 0119)

# st_write(SJER003_poly, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\NEON_plot_polys_from_field_coords\\SJER003.gpkg")

# st_write(SJER003_poly, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\2_standardized-data\\field-plot-boundaries\\0119.gpkg")

units::set_units(st_area(SJER003_poly), "m^2")

## SOAP_010.basePlot.vst

# import coords

SOAP010_31 <- GET("https://data.neonscience.org/api/v0/locations/SOAP_010.basePlot.vst.31")

SOAP010_31 <- jsonlite::fromJSON(content(SOAP010_31, as="text"))

SOAP010_33 <- GET("https://data.neonscience.org/api/v0/locations/SOAP_010.basePlot.vst.33")

SOAP010_33 <- jsonlite::fromJSON(content(SOAP010_33, as="text"))

SOAP010_49 <- GET("https://data.neonscience.org/api/v0/locations/SOAP_010.basePlot.vst.49")

SOAP010_49 <- jsonlite::fromJSON(content(SOAP010_49, as="text"))

SOAP010_51 <- GET("https://data.neonscience.org/api/v0/locations/SOAP_010.basePlot.vst.51")

SOAP010_51 <- jsonlite::fromJSON(content(SOAP010_51, as="text"))

# make coords into a polygon

SOAP010 <- data.frame(
  lon = c(SOAP010_31$data$locationDecimalLongitude, SOAP010_33$data$locationDecimalLongitude, SOAP010_51$data$locationDecimalLongitude, SOAP010_49$data$locationDecimalLongitude),
  lat = c(SOAP010_31$data$locationDecimalLatitude, SOAP010_33$data$locationDecimalLatitude, SOAP010_51$data$locationDecimalLatitude, SOAP010_49$data$locationDecimalLatitude)
)

SOAP010_poly <- SOAP010 %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")

SOAP010_poly <- SOAP010_poly %>% add_column(ofo_plot_id = 0120)

# st_write(SOAP010_poly, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\2_standardized-data\\field-plot-boundaries\\0120.gpkg")

# st_write(SOAP010_poly, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\NEON_plot_polys_from_field_coords\\SOAP010.gpkg")

units::set_units(st_area(SOAP010_poly), "m^2")

#### and now the tower plots. coordinates for the whole plots are 21, 25, 61, 57. coordinates for subplots 21_400 are 21, 23, 41, 39. coordinates for subplots 23_400 are 23, 25, 43, 41. coordinates for subplots 41_400 are 41, 43, 61, 59. coordinates for subplots 39_400 are 39, 41, 59, 57. 

## SJER_063.basePlot.vst; subplots sampled 23_400 and 41_400

# 23_400

# import coords

SJER063_23 <- GET("https://data.neonscience.org/api/v0/locations/SJER_063.basePlot.vst.23")

SJER063_23 <- jsonlite::fromJSON(content(SJER063_23, as="text"))

SJER063_25 <- GET("https://data.neonscience.org/api/v0/locations/SJER_063.basePlot.vst.25")

SJER063_25 <- jsonlite::fromJSON(content(SJER063_25, as="text"))

SJER063_43 <- GET("https://data.neonscience.org/api/v0/locations/SJER_063.basePlot.vst.43")

SJER063_43 <- jsonlite::fromJSON(content(SJER063_43, as="text"))

SJER063_41 <- GET("https://data.neonscience.org/api/v0/locations/SJER_063.basePlot.vst.41")

SJER063_41 <- jsonlite::fromJSON(content(SJER063_41, as="text"))

# make coords into a polygon

SJER063_a <- data.frame(
  lon = c(SJER063_23$data$locationDecimalLongitude, SJER063_25$data$locationDecimalLongitude, SJER063_43$data$locationDecimalLongitude, SJER063_41$data$locationDecimalLongitude),
  lat = c(SJER063_23$data$locationDecimalLatitude, SJER063_25$data$locationDecimalLatitude, SJER063_43$data$locationDecimalLatitude, SJER063_41$data$locationDecimalLatitude)
)

SJER063_a_poly <- SJER063_a %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")

# 41_400

# import coords

SJER063_41 <- GET("https://data.neonscience.org/api/v0/locations/SJER_063.basePlot.vst.41")

SJER063_41 <- jsonlite::fromJSON(content(SJER063_41, as="text"))

SJER063_43 <- GET("https://data.neonscience.org/api/v0/locations/SJER_063.basePlot.vst.43")

SJER063_43 <- jsonlite::fromJSON(content(SJER063_43, as="text"))

SJER063_61 <- GET("https://data.neonscience.org/api/v0/locations/SJER_063.basePlot.vst.61")

SJER063_61 <- jsonlite::fromJSON(content(SJER063_61, as="text"))

SJER063_59 <- GET("https://data.neonscience.org/api/v0/locations/SJER_063.basePlot.vst.59")

SJER063_59 <- jsonlite::fromJSON(content(SJER063_59, as="text"))

# make coords into a polygon

SJER063_b <- data.frame(
  lon = c(SJER063_41$data$locationDecimalLongitude, SJER063_43$data$locationDecimalLongitude, SJER063_61$data$locationDecimalLongitude, SJER063_59$data$locationDecimalLongitude),
  lat = c(SJER063_41$data$locationDecimalLatitude, SJER063_43$data$locationDecimalLatitude, SJER063_61$data$locationDecimalLatitude, SJER063_59$data$locationDecimalLatitude)
)

SJER063_b_poly <- SJER063_b %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")

# combine both subplots

SJER063_poly <- st_union (SJER063_a_poly, SJER063_b_poly)

units::set_units(st_area(SJER063_poly), "m^2")

st_centroid(SJER063_poly)

# st_write(SJER063_poly, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\NEON_plot_polys_from_field_coords\\SJER063.gpkg")

```

Automate this ridiculous process

```{r}
# automate

# going to make multiple for loops: one for creating the "subplot a" data/polygons, one for creating the "subplot b" polygons, and one for merging the polys together and calculating area/centroid and writing to gpkg

# subplot a calcs

for(i in 1:nrow(towersubplotssampled)) {
  
  plot_id_current = towersubplotssampled[i,]$NEONplotID
  
  if((towersubplotssampled[i,]$randomsubplota) == "21_400") {
    
    # import coords. coordinates for subplots 21_400 are points 21, 23, 41, 39.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.21?apiToken=")
    
assign(paste0(plot_id_current, "_21"), httr::GET(api_string))

assign(paste0(plot_id_current, "_21"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_21"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.23?apiToken=")

assign(paste0(plot_id_current, "_23"), httr::GET(api_string))

assign(paste0(plot_id_current, "_23"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_23"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=")

assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.39?apiToken=")

assign(paste0(plot_id_current, "_39"), httr::GET(api_string))

assign(paste0(plot_id_current, "_39"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_39"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_a"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_21"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_21"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_a_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_a"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_a_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_a_st_as_sf"))), "POLYGON")))
  }
  
  else if ((towersubplotssampled[i,]$randomsubplota) == "23_400") {
    
   # import coords. coordinates for subplots 23_400 are 23, 25, 43, 41.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.23?apiToken=.")
    
assign(paste0(plot_id_current, "_23"), httr::GET(api_string))

assign(paste0(plot_id_current, "_23"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_23"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.25?apiToken=.")
    
assign(paste0(plot_id_current, "_25"), httr::GET(api_string))

assign(paste0(plot_id_current, "_25"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_25"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.43?apiToken=.")
    
assign(paste0(plot_id_current, "_43"), httr::GET(api_string))

assign(paste0(plot_id_current, "_43"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_43"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_a"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_25"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_25"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_a_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_a"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_a_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_a_st_as_sf"))), "POLYGON")))
  }
  
  else if((towersubplotssampled[i,]$randomsubplota) == "41_400") {
    
    # import coords. coordinates for subplots 41_400 are 41, 43, 61, 59.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.43?apiToken=.")
    
assign(paste0(plot_id_current, "_43"), httr::GET(api_string))

assign(paste0(plot_id_current, "_43"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_43"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.61?apiToken=.")
    
assign(paste0(plot_id_current, "_61"), httr::GET(api_string))

assign(paste0(plot_id_current, "_61"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_61"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.59?apiToken=.")
    
assign(paste0(plot_id_current, "_59"), httr::GET(api_string))

assign(paste0(plot_id_current, "_59"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_59"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_a"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_61"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_61"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_a_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_a"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_a_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_a_st_as_sf"))), "POLYGON")))
  }
  
  else if ((towersubplotssampled[i,]$randomsubplota) == "39_400") {
    
    # import coords. coordinates for subplots 39_400 are 39, 41, 59, 57.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.39?apiToken=.")
    
assign(paste0(plot_id_current, "_39"), httr::GET(api_string))

assign(paste0(plot_id_current, "_39"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_39"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.59?apiToken=.")
    
assign(paste0(plot_id_current, "_59"), httr::GET(api_string))

assign(paste0(plot_id_current, "_59"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_59"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.57?apiToken=.")
    
assign(paste0(plot_id_current, "_57"), httr::GET(api_string))

assign(paste0(plot_id_current, "_57"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_57"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_a"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_57"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_57"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_a_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_a"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_a_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_a_st_as_sf"))), "POLYGON")))
  }
}

# subplot b calcs

for(i in 1:nrow(towersubplotssampled)) {
  
  plot_id_current = towersubplotssampled[i,]$NEONplotID
  
  if((towersubplotssampled[i,]$randomsubplotb) == "21_400") {
    
    # import coords. coordinates for subplots 21_400 are points 21, 23, 41, 39.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.21?apiToken=.")
    
assign(paste0(plot_id_current, "_21"), httr::GET(api_string))

assign(paste0(plot_id_current, "_21"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_21"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.23?apiToken=.")

assign(paste0(plot_id_current, "_23"), httr::GET(api_string))

assign(paste0(plot_id_current, "_23"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_23"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.")

assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.39?apiToken=.")

assign(paste0(plot_id_current, "_39"), httr::GET(api_string))

assign(paste0(plot_id_current, "_39"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_39"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_b"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_21"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_21"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_b_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_b"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_b_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_b_st_as_sf"))), "POLYGON")))
  }
  
  else if ((towersubplotssampled[i,]$randomsubplotb) == "23_400") {
    
   # import coords. coordinates for subplots 23_400 are 23, 25, 43, 41.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.23?apiToken=.")
    
assign(paste0(plot_id_current, "_23"), httr::GET(api_string))

assign(paste0(plot_id_current, "_23"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_23"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.25?apiToken=.")
    
assign(paste0(plot_id_current, "_25"), httr::GET(api_string))

assign(paste0(plot_id_current, "_25"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_25"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.43?apiToken=.")
    
assign(paste0(plot_id_current, "_43"), httr::GET(api_string))

assign(paste0(plot_id_current, "_43"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_43"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_b"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_25"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_25"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_b_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_b"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_b_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_b_st_as_sf"))), "POLYGON")))
  }
  
  else if((towersubplotssampled[i,]$randomsubplotb) == "41_400") {
    
    # import coords. coordinates for subplots 41_400 are 41, 43, 61, 59.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.43?apiToken=.")
    
assign(paste0(plot_id_current, "_43"), httr::GET(api_string))

assign(paste0(plot_id_current, "_43"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_43"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.61?apiToken=.")
    
assign(paste0(plot_id_current, "_61"), httr::GET(api_string))

assign(paste0(plot_id_current, "_61"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_61"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.59?apiToken=.")
    
assign(paste0(plot_id_current, "_59"), httr::GET(api_string))

assign(paste0(plot_id_current, "_59"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_59"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_b"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_61"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_61"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_b_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_b"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_b_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_b_st_as_sf"))), "POLYGON")))
  }
  
  else if ((towersubplotssampled[i,]$randomsubplotb) == "39_400") {
    
    # import coords. coordinates for subplots 39_400 are 39, 41, 59, 57.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.39?apiToken=.")
    
assign(paste0(plot_id_current, "_39"), httr::GET(api_string))

assign(paste0(plot_id_current, "_39"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_39"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.59?apiToken=.")
    
assign(paste0(plot_id_current, "_59"), httr::GET(api_string))

assign(paste0(plot_id_current, "_59"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_59"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.57?apiToken=.")
    
assign(paste0(plot_id_current, "_57"), httr::GET(api_string))

assign(paste0(plot_id_current, "_57"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_57"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_b"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_57"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_57"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_b_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_b"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_b_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_b_st_as_sf"))), "POLYGON")))
  }
}

# combine both subplots

SJER063_poly <- st_union (SJER063_a_poly, SJER063_b_poly)

units::set_units(st_area(SJER063_poly), "m^2")

st_centroid(SJER063_poly)

# st_write(SJER063_poly, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\NEON_plot_polys_from_field_coords\\SJER063.gpkg")

# towersubplotssampled <- towersubplotssampled %>%
#   add_column(poly = "")

# towersubplotssampled <- towersubplotssampled %>%
#   add_column(area_meters = "")

# towersubplotssampled <- towersubplotssampled %>%
#   add_column(centroid = "")

# towersubplotssampled <- towersubplotssampled %>%
#   add_column(ofo_plot_id = "")

# towersubplotssampled$index <- 1:nrow(towersubplotssampled)

for(i in 1:nrow(towersubplotssampled)) {
  
  plot_id_current = towersubplotssampled[i,]$NEONplotID
  
  # create spatial polygon and save list of coordinates to existing data.frame

assign(paste0(plot_id_current, "_poly"), st_union((get(paste0(plot_id_current, "_a_poly"))), (get(paste0(plot_id_current, "_b_poly")))))

towersubplotssampled[i,]$poly <-(get(paste0(plot_id_current, "_poly")))

  # calculate area of plot in square meters and save to existing data.frame

towersubplotssampled[i,]$area_meters <- units::set_units(st_area(get(paste0(plot_id_current, "_poly"))), "m^2")

  # calculate centroid of plot and save list of coordinates to existing data.frame

towersubplotssampled[i,]$centroid <-st_centroid(get(paste0(plot_id_current, "_poly")))

# add ofo plot IDs to spatial polygons and save them

assign(paste0(plot_id_current, "_poly"), ((get(paste0(plot_id_current, "_poly"))) %>% add_column(ofo_plot_id = ((towersubplotssampled[i,]$index) + 120))))

# save to primary data folder

st_write(get(paste0(plot_id_current, "_poly")), paste0("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\NEON_plot_polys_from_field_coords\\", plot_id_current, ".gpkg"), overwrite = TRUE)

# and field plot boundary folder

st_write(get(paste0(plot_id_current, "_poly")), paste0("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\2_standardized-data\\field-plot-boundaries\\0", (get(paste0(plot_id_current, "_poly"))$ofo_plot_id), ".gpkg"), overwrite = TRUE)

# add ofo plot IDs to existing data.frame of tower subplots sampled

towersubplotssampled[i,]$ofo_plot_id <- ((towersubplotssampled[i,]$index) + 120)

}

# export the towersubplotssampled data.frame to copy/paste into field-plots google sheet

towersubplotscsv <- towersubplotssampled

towersubplotscsv$centroid <- as.character(towersubplotscsv$centroid)

towersubplotscsv$poly <- as.character(towersubplotscsv$poly)

# write.csv(towersubplotscsv, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\towersubplotinfo.csv")

# OH NO two of the plots are completely outside the OFO flight footprints now that I have the correct coordinates and dimensions :( must remove TEAK052 and TEAK049, re-do OFO plot ID numbers, and then re-run the final for loop to generate correct plot boundaries with the right ofo plot id names. 

towersubplotssampled <- towersubplotssampled %>% filter (NEONplotID == 'SOAP_054' | NEONplotID == 'SOAP_047' | NEONplotID == 'SOAP_031' | NEONplotID == 'SOAP_049' | NEONplotID == 'SOAP_056' | NEONplotID == 'SOAP_053' | NEONplotID == 'SOAP_045' | NEONplotID == 'SJER_059' | NEONplotID == 'SJER_058' | NEONplotID == 'SJER_047' | NEONplotID == 'SJER_054' | NEONplotID == 'SJER_060' | NEONplotID == 'SJER_049' | NEONplotID == 'SJER_045' | NEONplotID == 'SJER_003' | NEONplotID == 'SJER_057' | NEONplotID == 'SJER_046' | NEONplotID == 'SJER_055' | NEONplotID == 'SJER_048' | NEONplotID == 'SJER_061' | NEONplotID == 'SJER_050' | NEONplotID == 'SJER_053' | NEONplotID == 'SJER_063' | NEONplotID == 'SJER_051' | NEONplotID == 'TEAK_062' | NEONplotID == 'TEAK_058' | NEONplotID == 'TEAK_043' | NEONplotID == 'TEAK_056' | NEONplotID == 'TEAK_044' | NEONplotID == 'TEAK_057' | NEONplotID == 'TEAK_047' | NEONplotID == 'TEAK_053' | NEONplotID == 'TEAK_051' | NEONplotID == 'TEAK_046' | NEONplotID == 'SOAP_057' | NEONplotID == 'SOAP_059' | NEONplotID == 'SOAP_051' | NEONplotID == 'SOAP_048' | NEONplotID == 'SOAP_058' | NEONplotID == 'SOAP_043' | NEONplotID == 'SOAP_050' | NEONplotID == 'SOAP_044' | NEONplotID == 'SOAP_052' | NEONplotID == 'SOAP_046' | NEONplotID == 'SOAP_060' | NEONplotID == 'SOAP_055' | NEONplotID == 'SOAP_061' | NEONplotID == 'SOAP_010')

towersubplotssampled <- towersubplotssampled[,-c(29:33)]

# re-ran final for loop code chunk! phew.

```

Automation loop for WREF plots

```{r}
# automate

plots <- dplyr::distinct(plots, NEONplotID, .keep_all = TRUE)

# going to make multiple for loops: one for creating the "subplot a" data/polygons, one for creating the "subplot b" polygons, and one for merging the polys together and calculating area/centroid and writing to gpkg

# subplot a calcs

for(i in 1:nrow(plots)) {
  
  plot_id_current = plots[i,]$NEONplotID
  
  if((plots[i,]$randomsubplota) == "21_400") {
    
    # import coords. coordinates for subplots 21_400 are points 21, 23, 41, 39.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.21?apiToken=eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_21"), httr::GET(api_string))

assign(paste0(plot_id_current, "_21"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_21"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.23?apiToken=eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")

assign(paste0(plot_id_current, "_23"), httr::GET(api_string))

assign(paste0(plot_id_current, "_23"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_23"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")

assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.39?apiToken=eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")

assign(paste0(plot_id_current, "_39"), httr::GET(api_string))

assign(paste0(plot_id_current, "_39"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_39"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_a"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_21"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_21"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_a_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_a"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_a_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_a_st_as_sf"))), "POLYGON")))
  }
  
  else if ((plots[i,]$randomsubplota) == "23_400") {
    
   # import coords. coordinates for subplots 23_400 are 23, 25, 43, 41.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.23?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_23"), httr::GET(api_string))

assign(paste0(plot_id_current, "_23"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_23"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.25?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_25"), httr::GET(api_string))

assign(paste0(plot_id_current, "_25"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_25"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.43?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_43"), httr::GET(api_string))

assign(paste0(plot_id_current, "_43"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_43"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_a"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_25"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_25"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_a_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_a"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_a_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_a_st_as_sf"))), "POLYGON")))
  }
  
  else if((plots[i,]$randomsubplota) == "41_400") {
    
    # import coords. coordinates for subplots 41_400 are 41, 43, 61, 59.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.43?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_43"), httr::GET(api_string))

assign(paste0(plot_id_current, "_43"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_43"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.61?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_61"), httr::GET(api_string))

assign(paste0(plot_id_current, "_61"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_61"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.59?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_59"), httr::GET(api_string))

assign(paste0(plot_id_current, "_59"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_59"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_a"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_61"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_61"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_a_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_a"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_a_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_a_st_as_sf"))), "POLYGON")))
  }
  
  else if ((plots[i,]$randomsubplota) == "39_400") {
    
    # import coords. coordinates for subplots 39_400 are 39, 41, 59, 57.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.39?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_39"), httr::GET(api_string))

assign(paste0(plot_id_current, "_39"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_39"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.59?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_59"), httr::GET(api_string))

assign(paste0(plot_id_current, "_59"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_59"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.57?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_57"), httr::GET(api_string))

assign(paste0(plot_id_current, "_57"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_57"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_a"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_57"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_57"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_a_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_a"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_a_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_a_st_as_sf"))), "POLYGON")))
  }
}

# subplot b calcs

for(i in 1:nrow(plots)) {
  
  plot_id_current = plots[i,]$NEONplotID
  
  if((plots[i,]$randomsubplotb) == "21_400") {
    
    # import coords. coordinates for subplots 21_400 are points 21, 23, 41, 39.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.21?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_21"), httr::GET(api_string))

assign(paste0(plot_id_current, "_21"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_21"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.23?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")

assign(paste0(plot_id_current, "_23"), httr::GET(api_string))

assign(paste0(plot_id_current, "_23"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_23"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")

assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.39?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")

assign(paste0(plot_id_current, "_39"), httr::GET(api_string))

assign(paste0(plot_id_current, "_39"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_39"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_b"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_21"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_21"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_b_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_b"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_b_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_b_st_as_sf"))), "POLYGON")))
  }
  
  else if ((plots[i,]$randomsubplotb) == "23_400") {
    
   # import coords. coordinates for subplots 23_400 are 23, 25, 43, 41.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.23?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_23"), httr::GET(api_string))

assign(paste0(plot_id_current, "_23"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_23"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.25?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_25"), httr::GET(api_string))

assign(paste0(plot_id_current, "_25"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_25"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.43?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_43"), httr::GET(api_string))

assign(paste0(plot_id_current, "_43"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_43"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_b"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_25"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_23"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_25"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_b_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_b"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_b_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_b_st_as_sf"))), "POLYGON")))
  }
  
  else if((plots[i,]$randomsubplotb) == "41_400") {
    
    # import coords. coordinates for subplots 41_400 are 41, 43, 61, 59.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.43?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_43"), httr::GET(api_string))

assign(paste0(plot_id_current, "_43"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_43"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.61?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_61"), httr::GET(api_string))

assign(paste0(plot_id_current, "_61"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_61"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.59?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_59"), httr::GET(api_string))

assign(paste0(plot_id_current, "_59"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_59"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_b"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_61"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_43"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_61"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_b_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_b"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_b_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_b_st_as_sf"))), "POLYGON")))
  }
  
  else if ((plots[i,]$randomsubplotb) == "39_400") {
    
    # import coords. coordinates for subplots 39_400 are 39, 41, 59, 57.
    
api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.39?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_39"), httr::GET(api_string))

assign(paste0(plot_id_current, "_39"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_39"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.41?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_41"), httr::GET(api_string))

assign(paste0(plot_id_current, "_41"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_41"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.59?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_59"), httr::GET(api_string))

assign(paste0(plot_id_current, "_59"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_59"), httr::GET(api_string))), as="text")))

api_string <- paste0("https://data.neonscience.org/api/v0/locations/", plot_id_current, ".basePlot.vst.57?apiToken=.eyJ0eXAiOiJKV1QiLCJhbGciOiJFUzI1NiJ9.eyJhdWQiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnL2FwaS92MC8iLCJzdWIiOiJlbXB1cnZpc0B1Y2RhdmlzLmVkdSIsInNjb3BlIjoicmF0ZTpwdWJsaWMiLCJpc3MiOiJodHRwczovL2RhdGEubmVvbnNjaWVuY2Uub3JnLyIsImV4cCI6MTg3NzkxODE4NCwiaWF0IjoxNzIwMjM4MTg0LCJlbWFpbCI6ImVtcHVydmlzQHVjZGF2aXMuZWR1In0.vU4811sSuSkebx4juHjXE-pVrX-lcactze1elv36s_TAuCQyDkeQuh1T8TW7TcvxBlq-XuhROtZGhmTp52t_JA")
    
assign(paste0(plot_id_current, "_57"), httr::GET(api_string))

assign(paste0(plot_id_current, "_57"), jsonlite::fromJSON(content((assign(paste0(plot_id_current, "_57"), httr::GET(api_string))), as="text")))

# make coords into a polygon

assign(paste0(plot_id_current, "_b"), data.frame(
  lon = c(
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLongitude),
    (get(paste0(plot_id_current, "_57"))$data$locationDecimalLongitude)),
  lat = c(
    (get(paste0(plot_id_current, "_39"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_41"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_59"))$data$locationDecimalLatitude),
    (get(paste0(plot_id_current, "_57"))$data$locationDecimalLatitude))
))

assign(paste0(plot_id_current, "_b_st_as_sf"), st_as_sf(as.data.frame(get(paste0(plot_id_current, "_b"))), coords = c("lon","lat"), crs = 4326))

assign(paste0(plot_id_current, "_b_poly"), st_as_sf(st_cast(st_combine(get(paste0(plot_id_current, "_b_st_as_sf"))), "POLYGON")))
  }
}

# combine both subplots

#plots <- plots %>%
#  add_column(poly = "")

#plots <- plots %>%
#  add_column(area_meters = "")

#plots <- plots %>%
#  add_column(centroid = "")

#plots <- plots %>%
#  add_column(ofo_plot_id = "")

#plots$index <- 1:nrow(plots)

for(i in 1:nrow(plots)) {
  
  plot_id_current = plots[i,]$NEONplotID
  
  # create spatial polygon and save list of coordinates to existing data.frame

assign(paste0(plot_id_current, "_poly"), st_union((get(paste0(plot_id_current, "_a_poly"))), (get(paste0(plot_id_current, "_b_poly")))))

plots[i,]$poly <-(get(paste0(plot_id_current, "_poly")))

  # calculate area of plot in square meters and save to existing data.frame

plots[i,]$area_meters <- units::set_units(st_area(get(paste0(plot_id_current, "_poly"))), "m^2")

  # calculate centroid of plot and save list of coordinates to existing data.frame

plots[i,]$centroid <-st_centroid(get(paste0(plot_id_current, "_poly")))

# add ofo plot IDs to spatial polygons and save them

assign(paste0(plot_id_current, "_poly"), ((get(paste0(plot_id_current, "_poly"))) %>% add_column(ofo_plot_id = ((plots[i,]$index) + 171))))

# save to primary data folder

st_write(get(paste0(plot_id_current, "_poly")), paste0("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\NEON_plot_polys_from_field_coords\\", plot_id_current, ".gpkg"), overwrite = TRUE)

# and field plot boundary folder

st_write(get(paste0(plot_id_current, "_poly")), paste0("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\2_standardized-data\\field-plot-boundaries\\0", (get(paste0(plot_id_current, "_poly"))$ofo_plot_id), ".gpkg"), overwrite = TRUE)

# add ofo plot IDs to existing data.frame of tower subplots sampled

plots[i,]$ofo_plot_id <- ((plots[i,]$index) + 171)

}

# export the towersubplotssampled data.frame to copy/paste into field-plots google sheet

subplotscsv <- plots

subplotscsv$centroid <- as.character(subplotscsv$centroid)

subplotscsv$poly <- as.character(subplotscsv$poly)

write.csv(subplotscsv, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\WREFsubplotinfo.csv")

```

Calculate centroids of subplots

```{r}

#towersubplotssampled <- towersubplotssampled %>%
#  add_column(centroid_a = "")

#towersubplotssampled <- towersubplotssampled %>%
#  add_column(centroid_b = "")

for(i in 1:nrow(towersubplotssampled)) {
  
  plot_id_current = towersubplotssampled[i,]$NEONplotID

towersubplotssampled[i,]$centroid_a <-st_centroid(get(paste0(plot_id_current, "_a_poly")))

towersubplotssampled[i,]$centroid_b <-st_centroid(get(paste0(plot_id_current, "_b_poly")))

}

towersubplotscsv <- towersubplotssampled

towersubplotscsv$centroid <- as.character(towersubplotscsv$centroid)

towersubplotscsv$poly <- as.character(towersubplotscsv$poly)

towersubplotscsv$centroid_a <- as.character(towersubplotscsv$centroid_a)

towersubplotscsv$centroid_b <- as.character(towersubplotscsv$centroid_b)

write.csv(towersubplotscsv, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\towersubplotinfo.csv")

```

Calculate subplot centroids for WREF plots

```{r}

plots <- plots %>%
  add_column(centroid_a = "")

plots <- plots %>%
  add_column(centroid_b = "")

for(i in 1:nrow(plots)) {
  
  plot_id_current = plots[i,]$NEONplotID

plots[i,]$centroid_a <-st_centroid(get(paste0(plot_id_current, "_a_poly")))

plots[i,]$centroid_b <-st_centroid(get(paste0(plot_id_current, "_b_poly")))

}

subplotscsv <- plots

subplotscsv$centroid <- as.character(subplotscsv$centroid)

subplotscsv$poly <- as.character(subplotscsv$poly)

subplotscsv$centroid_a <- as.character(subplotscsv$centroid_a)

subplotscsv$centroid_b <- as.character(subplotscsv$centroid_b)

write.csv(subplotscsv, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\WREFsubplotinfo_withsubplotcentroids.csv")

```

Put the tree coordinates back in the data frame

```{r}

unique_treesinplots_final <- unique_treesinplots_final %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])

```

Live/dead

```{r}

# rename a few variables while we're at it

unique_treesinplots_final <- unique_treesinplots_final %>% rename(stemDistance_m = stemDistance, stemAzimuth_degrees = stemAzimuth, notes_1 = remarks.x, notes_2 = remarks.y)

# live/dead

unique_treesinplots_final <- unique_treesinplots_final %>%
  add_column(livedead = "")

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == "Dead, broken bole")] <- 'D'

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == 'Live')] <- 'L'

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == "Live,  other damage")] <- "L"

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == 'Live, insect damaged')] <- 'L'

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == 'Live, physically damaged')] <- 'L'

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == 'Removed')] <- '' # no idea what this means, no indication in other columns or notes

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == 'Standing dead')] <- 'D'

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == 'Live, broken bole')] <- 'L'

unique_treesinplots_final$livedead[(unique_treesinplots_final$plantStatus == 'Live, disease damaged')] <- 'L'

```

Species codes

```{r}
# spp codes

unique_treesinplots_final <- unique_treesinplots_final %>%
  add_column(sppcode = "")

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == '2PLANT'] <- '000' # code for unknown

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ABCO'] <- '015' 

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ABLO'] <- '015' # I had no idea Abies concolor had a synonym, Abies lowiana

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ABMA'] <- '020'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ARVIM'] <- 'ARCVISM' # Arctostaphylos viscida ssp. mariposa

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'CADE27'] <- '081'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'CELE2'] <- 'CEALEU' # Ceanothus leucodermis

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'CEMOG'] <- 'CERMONG' # Cercocarpus montanus var. glaber

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PICO'] <- '108'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PIJE'] <- '116'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PILA'] <- '117'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PINACE'] <- '100' # Pinus spp

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PIPO'] <- '122'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PISA2'] <- '127'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'QUCH2'] <- '805'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'QUDO'] <- '807'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'QUKE'] <- '818'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'QUWI2'] <- '839'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'RHIL'] <- 'RHAILI' # Rhamnus ilicifolia

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ABAM'] <- '011'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ABGR'] <- '017'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ABIES'] <- '010'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ACCI'] <- 'ACECIR' # acer circinatum 

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'CONU4'] <- '492'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'ALRU2'] <- '351'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PIMO3'] <- '119'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PINACE'] <- '100'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'PSMEM'] <- '202'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'TABR2'] <- '231'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'THPL'] <- '242'

unique_treesinplots_final$sppcode[unique_treesinplots_final$taxonID == 'TSHE'] <- '263'

```

Canopy position and overhead visibility

```{r}

unique_treesinplots_final <- unique_treesinplots_final %>%
  add_column(canpos_code = "")

unique_treesinplots_final <- unique_treesinplots_final %>%
  add_column(ohvis = "")

unique_treesinplots_final$canpos_code[unique_treesinplots_final$canopyPosition == 'Full shade'] <- '5' 

unique_treesinplots_final$canpos_code[unique_treesinplots_final$canopyPosition == 'Full sun'] <- '2' 

unique_treesinplots_final$canpos_code[unique_treesinplots_final$canopyPosition == 'Mostly shaded'] <- '4' 

unique_treesinplots_final$canpos_code[unique_treesinplots_final$canopyPosition == 'Open grown'] <- '1' 

unique_treesinplots_final$canpos_code[unique_treesinplots_final$canopyPosition == 'Partially shaded'] <- '3' 

unique_treesinplots_final$canpos_code[is.na(unique_treesinplots_final$canopyPosition)] <- '' 

unique_treesinplots_final$ohvis[unique_treesinplots_final$canpos_code == '5'] <- 'FALSE' 

unique_treesinplots_final$ohvis[unique_treesinplots_final$canpos_code == '4'] <- 'FALSE' 

unique_treesinplots_final$ohvis[unique_treesinplots_final$canpos_code == '3'] <- 'TRUE' 

unique_treesinplots_final$ohvis[unique_treesinplots_final$canpos_code == '2'] <- 'TRUE' 

unique_treesinplots_final$ohvis[unique_treesinplots_final$canpos_code == '1'] <- 'TRUE' 

unique_treesinplots_final$ohvis[unique_treesinplots_final$canpos_code == ''] <- '' 

```

Growth form

```{r}

# we only took tree measurements from these plots

unique_treesinplots_final <- unique_treesinplots_final %>%
  add_column(ofo_growthform = "tree")

```

Crown width

```{r}

names(unique_treesinplots_final)[names(unique_treesinplots_final) == "maxCrownDiameter"] <- "crown_width_1_meters"


names(unique_treesinplots_final)[names(unique_treesinplots_final) == "ninetyCrownDiameter"] <- "crown_width_2_meters"

# get rid of NAs

unique_treesinplots_final$crown_width_1_meters[is.na(unique_treesinplots_final$crown_width_1_meters)] <- '' 

unique_treesinplots_final$crown_width_2_meters[is.na(unique_treesinplots_final$crown_width_2_meters)] <- '' 

```

Check the "remarks" for damage codes, downed trees we need to delete, etc.

```{r}

# nothing useful in column "notes1" but there is some good stuff in "notes2"

#unique_treesinplots_final <- unique_treesinplots_final %>%
#  add_column(damage_1 = "")

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Biotic; Mistletoe'] <- '23000' # code for parasitic/epiphytic plants 

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Biotic; Mistletoe. Browning needles'] <- '23000 90008' #code for parasitic/epiphytic plants and foliage discoloration

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'All needles are dead.'] <- '90008' #code for foliage discoloration

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Biotic; Mistletoe| Dead; Unknown cause of death'] <- '23000' 

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken Bole; Windfall'] <- '50019 90001' # mechanical damage and broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken Bole; Windfall; cannot discern top'] <- '50019 90001' # mechanical damage and broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken Bole; Windfall| Dead; Unknown cause of death'] <- '50019 90001' # mechanical damage and broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'dead and hollow charred shell from fire'] <- '30000' # fire

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Dead. Evidence of bark beetles'] <- '11000' # bark beetles

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Drought induced yellowing'] <- '90008 50003' # foliage discoloration and drought

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == '	
Drought induced yellowing | Apparent change in measurementHeight from 2015 due to duff layer changes'] <- '90008 50003' # foliage discoloration and drought

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Drought stressed | Physical; Other'] <- '50003 99000' # drought and other

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Drought stressed | Physical; Other | Apparent change in measurementHeight from 2015 due to duff layer changes'] <- '50003 99000' # drought and other

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'fire damaged'] <- '30000'

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'fire damaged bole partially hollow at 130cm'] <- '30000'

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'fire damaged; proofed and errors corrected'] <- '30000'

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'fire damaged; proofed and no errors'] <- '30000'

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Fire; Bole scorch'] <- '30000'

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Fire; Burned through at base'] <- '30000'

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Insect Damage; Bark| Dead; Unknown cause of death'] <- '10000' # general code for insect damage

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'live tree leaning more than 45 degrees'] <- '90006' # crook or lean

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Physical; Bole scar | Apparent change in measurementHeight from 2015 due to duff layer changes'] <- '10000' # general code for insects

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Physical; Crown damage'] <- '90013'

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'tree growing at an angle but still alive'] <- '90006' # crook or lean

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Leaning and has another tree leaning onto it'] <- '90006' # crook or lean

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Leaning at approximately 32 degrees'] <- '90006' # crook or lean

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Leaning dead'] <- '90006' # crook or lean

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Leaning due to another tree falling on it'] <- '90006' # crook or lean

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Leaning due to downed trees falling on it. Dead branches and browned foliage.'] <- '90006' # crook or lean

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken leader'] <- '90001' # broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken top'] <- '90001' # broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken top and branches'] <- '90001' # broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken'] <- '90001' # broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken Bole; Unknown cause of death'] <- '90001' # broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken top and heavily infested with dwarf mistletoe'] <- '90001 23000' # broken top and parasitic/epiphytic plants

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Broken top and leaning. Substantial bark missing up to 3m from base.'] <- '90001 90006' # broken top and crook or lean

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Butt rot'] <- '21000' # root/butt diseases

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'witches broom'] <- '19000' # many causes of witches' brooms so I'm labelling them as "general diseases"

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Black fungus'] <- '19000' # many fungi, noting it as "general diseases" 

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Black fungus on surface of branch scars'] <- '19000' # many fungi, noting it as "general diseases" 

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Black fungus.'] <- '19000' # many fungi, noting it as "general diseases" 

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Basal scarring and bark peeling'] <- '90000' # code for "other damages"-- I'm not sure if these things are due to insects, weather, parasites, fungi, mechanical damage....

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Bole damage'] <- '90000' # code for "other damages"-- I'm not sure if these things are due to insects, weather, parasites, fungi, mechanical damage....

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Bole damage - pitching out'] <- '90000' # code for "other damages"-- I'm not sure if these things are due to insects, weather, parasites, fungi, mechanical damage...

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Bole scar'] <- '90000' # code for "other damages"-- I'm not sure if the scar is due to mechanical damage, insects, fire, or what

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Bole scars'] <- '90000' # code for "other damages"-- I'm not sure if the scar is due to mechanical damage, insects, fire, or what

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Catface | Disease; Unknown'] <- '19000' # code for "general diseases"

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Catface. Pileated Woodpecker holes. Dband reset due to deep furrow causing band to bow'] <- '41000 90000' # code for general wild animal damage and "other"

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Cause of break unknown'] <- '90001' # broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Chlorosis'] <- '90008' # foliage discoloration

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Dead; Unknown cause of death. Broken bole'] <- '90001' # broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Disease; Canker'] <- '22000' # general code for cankers

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Disease; Other. Fungus'] <- '19000' # code for general diseases

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Disease; Unknown'] <- '19000' # code for general diseases

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Dwarf mistletoe'] <- '23001' # general code for mistletoe

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Dwarf mistletoe brooms in canopy'] <- '23001' # general code for mistletoe

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Inner needles turning brown and falling off'] <- '90008' # foliage discoloration

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Insect and woodpecker damage'] <- '10000 41000' # codes for "general insects" and "wild animals"

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Insect holes'] <- '10000' # general insects

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Insect Species; Other'] <- '10000' # general insects

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Mechanical damage lower bole'] <- '70008' # mechanical damage code

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Missing bark leaning and broken top'] <- '90000 90006 90001' # codes for "other," "crook or lean," and "broken top"

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Old damage to bole. New top grown'] <- '90000 90001' # code for other and broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Other Damage; Bole damage'] <- '90000' # other

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Other Damage; Crown damage'] <- '90000' # other

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Physical; Crown damage'] <- '90000' # other

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Physical; Other peeling bark| Broken Bole: break height 10.3m break diameter 15.0 cm'] <- '90000 90001' # other and broken top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Physically damaged from fallen trees'] <- '20000' # biotic damage

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Rust'] <- '19000' # general diseases-- not sure if it's a stem rust, broom rust, etc

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Significantly decayed at dbh'] <- '22500' # stem decay

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Small amount of dwarf mistletoe'] <- '23001' # mistletoe

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'Two leaders'] <- '90004' # forked top

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'White pine blister rust'] <- '26001' # WPBR

unique_treesinplots_final$damage_1[unique_treesinplots_final$notes_2 == 'White pine blister rust and basal resinosus'] <- '26001' # WPBR

unique_treesinplots_final$damage_1[is.na(unique_treesinplots_final$damage_1)] <- ''

# column "plantStatus" has some stuff about insect damage, physical damage, etc

unique_treesinplots_final <- unique_treesinplots_final %>%
  add_column(damage_2 = "")

unique_treesinplots_final$damage_2[unique_treesinplots_final$plantStatus == 'Live,  other damage'] <- '99000' # unknown damage code

unique_treesinplots_final$damage_2[unique_treesinplots_final$plantStatus == 'Live, insect damaged'] <- '10000' # general code for insect damage

unique_treesinplots_final$damage_2[unique_treesinplots_final$plantStatus == 'Live, physically damaged'] <- '50019' # general code for mechanical damage

unique_treesinplots_final$damage_2[unique_treesinplots_final$plantStatus == 'Live, broken bole'] <- '90001' # broken top

unique_treesinplots_final$damage_2[unique_treesinplots_final$plantStatus == 'Live, disease damaged'] <- '19000' # general diseases

```

OFO plot ID numbers

```{r}

treesfinal <- unique_treesinplots_final

treesfinal <- treesfinal[,-c(52:53)]

# treesfinal <- treesfinal |> left_join(towersubplotssampled)

treesfinal <- treesfinal |> left_join(plots)

treesfinal <- treesfinal[,-c(82:83)]

treesfinal <- st_drop_geometry(treesfinal)

treesfinal <- treesfinal[,-c(74:80)]

treesfinal <- treesfinal[,-c(65:72)]

treesfinal <- treesfinal[,-c(51:62)]

treesfinal$ofo_plot_id[treesfinal$NEONplotID == 'SJER_003'] <- '0119'

treesfinal$ofo_plot_id[treesfinal$NEONplotID == 'SOAP_010'] <- '0120'

```

Add subplot ID

```{r}

treesfinal <- treesfinal %>%
  add_column(ofo_subplot_id = "")

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_045' & treesfinal$subplotID == '23_400'] <- '0341'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_045' & treesfinal$subplotID == '39_400'] <- '0342'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_046' & treesfinal$subplotID == '39_400'] <- '0343'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_046' & treesfinal$subplotID == '41_400'] <- '0344'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_047' & treesfinal$subplotID == '21_400'] <- '0345'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_047' & treesfinal$subplotID == '39_400'] <- '0346'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_048' & treesfinal$subplotID == '39_400'] <- '0347'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_048' & treesfinal$subplotID == '41_400'] <- '0348'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_049' & treesfinal$subplotID == '21_400'] <- '0349'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_049' & treesfinal$subplotID == '23_400'] <- '0350'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_050' & treesfinal$subplotID == '21_400'] <- '0351'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_050' & treesfinal$subplotID == '39_400'] <- '0352'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_051' & treesfinal$subplotID == '39_400'] <- '0353'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_051' & treesfinal$subplotID == '41_400'] <- '0354'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_053' & treesfinal$subplotID == '21_400'] <- '0355'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_053' & treesfinal$subplotID == 'unknown_400'] <- '0355'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_053' & treesfinal$subplotID == '23_400'] <- '0356'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_054' & treesfinal$subplotID == '21_400'] <- '0357'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_054' & treesfinal$subplotID == '39_400'] <- '0358'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_055' & treesfinal$subplotID == '21_400'] <- '0359'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_055' & treesfinal$subplotID == '39_400'] <- '0360'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_057' & treesfinal$subplotID == '23_400'] <- '0361'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_057' & treesfinal$subplotID == '39_400'] <- '0362'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_058' & treesfinal$subplotID == '39_400'] <- '0363'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_058' & treesfinal$subplotID == '41_400'] <- '0364'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_059' & treesfinal$subplotID == '21_400'] <- '0365'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_059' & treesfinal$subplotID == '23_400'] <- '0366'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_060' & treesfinal$subplotID == '23_400'] <- '0367'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_060' & treesfinal$subplotID == '39_400'] <- '0368'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_061' & treesfinal$subplotID == '39_400'] <- '0369'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_061' & treesfinal$subplotID == '41_400'] <- '0370'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_063' & treesfinal$subplotID == '23_400'] <- '0371'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SJER_063' & treesfinal$subplotID == '41_400'] <- '0372'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_031' & treesfinal$subplotID == '21_400'] <- '0373'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_031' & treesfinal$subplotID == '39_400'] <- '0374'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_043' & treesfinal$subplotID == '23_400'] <- '0375'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_043' & treesfinal$subplotID == '39_400'] <- '0376'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_043' & treesfinal$subplotID == '39_100_unknown'] <- '0376'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_044' & treesfinal$subplotID == '21_400'] <- '0377'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_044' & treesfinal$subplotID == '41_400'] <- '0378'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_045' & treesfinal$subplotID == '21_400'] <- '0379'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_045' & treesfinal$subplotID == '23_400'] <- '0380'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_046' & treesfinal$subplotID == '23_400'] <- '0381'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_046' & treesfinal$subplotID == '41_400'] <- '0382'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_047' & treesfinal$subplotID == '23_400'] <- '0383'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_047' & treesfinal$subplotID == '41_400'] <- '0384'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_048' & treesfinal$subplotID == '21_400'] <- '0385'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_048' & treesfinal$subplotID == '39_400'] <- '0386'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_049' & treesfinal$subplotID == '23_400'] <- '0387'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_049' & treesfinal$subplotID == '39_400'] <- '0388'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_050' & treesfinal$subplotID == '23_400'] <- '0389'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_050' & treesfinal$subplotID == '41_400'] <- '0390'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_051' & treesfinal$subplotID == '21_400'] <- '0391'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_051' & treesfinal$subplotID == '39_400'] <- '0392'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_052' & treesfinal$subplotID == '23_400'] <- '0393'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_052' & treesfinal$subplotID == '39_400'] <- '0394'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_053' & treesfinal$subplotID == '23_400'] <- '0395'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_053' & treesfinal$subplotID == '39_400'] <- '0396'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_054' & treesfinal$subplotID == '23_400'] <- '0397'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_054' & treesfinal$subplotID == '23_25_unknown'] <- '0397'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_054' & treesfinal$subplotID == '39_400'] <- '0398'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_054' & treesfinal$subplotID == '39_25_unknown'] <- '0398'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_055' & treesfinal$subplotID == '23_400'] <- '0399'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_055' & treesfinal$subplotID == '39_400'] <- '0399'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_055' & treesfinal$subplotID == '41_400'] <- '0400'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_056' & treesfinal$subplotID == '39_400'] <- '0401'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_056' & treesfinal$subplotID == '41_400'] <- '0402'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_057' & treesfinal$subplotID == '23_400'] <- '0403'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_057' & treesfinal$subplotID == '21_400'] <- '0403'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_057' & treesfinal$subplotID == '39_400'] <- '0404'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_058' & treesfinal$subplotID == '39_400'] <- '0405'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_058' & treesfinal$subplotID == '41_400'] <- '0406'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_059' & treesfinal$subplotID == '39_400'] <- '0407'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_059' & treesfinal$subplotID == '41_400'] <- '0408'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_060' & treesfinal$subplotID == '23_400'] <- '0409'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_060' & treesfinal$subplotID == '39_400'] <- '0410'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_061' & treesfinal$subplotID == '39_400'] <- '0411'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'SOAP_061' & treesfinal$subplotID == '41_400'] <- '0412'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_043' & treesfinal$subplotID == '21_400'] <- '0413'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_043' & treesfinal$subplotID == '23_400'] <- '0414'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_044' & treesfinal$subplotID == '21_400'] <- '0415'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_044' & treesfinal$subplotID == '39_400'] <- '0416'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_046' & treesfinal$subplotID == '39_400'] <- '0417'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_046' & treesfinal$subplotID == '41_400'] <- '0418'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_047' & treesfinal$subplotID == '21_400'] <- '0419'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_047' & treesfinal$subplotID == '41_400'] <- '0420'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_051' & treesfinal$subplotID == '21_400'] <- '0421'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_051' & treesfinal$subplotID == '41_400'] <- '0422'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_053' & treesfinal$subplotID == '21_400'] <- '0423'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_053' & treesfinal$subplotID == '41_400'] <- '0424'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_056' & treesfinal$subplotID == '21_400'] <- '0425'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_056' & treesfinal$subplotID == '39_400'] <- '0426'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_057' & treesfinal$subplotID == '21_400'] <- '0427'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_057' & treesfinal$subplotID == '23_400'] <- '0428'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_058' & treesfinal$subplotID == '21_400'] <- '0429'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_058' & treesfinal$subplotID == '41_400'] <- '0430'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_062' & treesfinal$subplotID == '21_400'] <- '0431'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'TEAK_062' & treesfinal$subplotID == '41_400'] <- '0432'

treesfinal <- treesfinal[,-c(15:16)]

treesfinal <- treesfinal[,-c(26:30)]

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_089' & treesfinal$subplotID == '21_400'] <- '0461'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_089' & treesfinal$subplotID == '41_400'] <- '0462'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_082' & treesfinal$subplotID == '21_400'] <- '0463'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_082' & treesfinal$subplotID == '41_400'] <- '0464'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_077' & treesfinal$subplotID == '21_400'] <- '0465'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_077' & treesfinal$subplotID == '39_400'] <- '0466'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_083' & treesfinal$subplotID == '21_400'] <- '0467'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_083' & treesfinal$subplotID == '23_400'] <- '0468'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_074' & treesfinal$subplotID == '23_400'] <- '0469'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_074' & treesfinal$subplotID == '39_400'] <- '0470'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_080' & treesfinal$subplotID == '21_400'] <- '0471'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_080' & treesfinal$subplotID == '39_400'] <- '0472'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_079' & treesfinal$subplotID == '21_400'] <- '0473'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_079' & treesfinal$subplotID == '41_400'] <- '0474'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_073' & treesfinal$subplotID == '23_400'] <- '0475'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_073' & treesfinal$subplotID == '39_400'] <- '0476'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_085' & treesfinal$subplotID == '39_400'] <- '0477'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_085' & treesfinal$subplotID == '41_400'] <- '0478'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_078' & treesfinal$subplotID == '23_400'] <- '0479'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_078' & treesfinal$subplotID == '41_400'] <- '0480'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_084' & treesfinal$subplotID == '23_400'] <- '0481'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_084' & treesfinal$subplotID == '41_400'] <- '0482'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_088' & treesfinal$subplotID == '39_400'] <- '0483'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_088' & treesfinal$subplotID == '41_400'] <- '0484'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_075' & treesfinal$subplotID == '21_400'] <- '0485'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_075' & treesfinal$subplotID == '23_400'] <- '0486'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_071' & treesfinal$subplotID == '21_400'] <- '0487'

treesfinal$ofo_subplot_id[treesfinal$NEONplotID == 'WREF_071' & treesfinal$subplotID == '23_400'] <- '0487'

treesfinal <- treesfinal[,-c(76:77)]

treesfinal <- treesfinal[,-c(71:73)]

```

Distance to pith

```{r}

treesfinal <- treesfinal %>%
  add_column(distance_pith_m = "")

treesfinal$distance_pith_m <- treesfinal$stemDistance_m + ((treesfinal$DBH_cm/2)*0.01)

```


Export

```{r}

write.csv (treesfinal, "C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0013\\data\\trees.csv")

```

## ALTERATIONS MADE IN EXCEL

Manually separated damage_1 codes into separate columns

