---
title: "O'Harrell Canyon and Indiana Summit"
author: "Emily Marie Purvis"
date: "2024-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA MANIPULATED IN R

Load libraries

```{r, echo=FALSE}

library(readxl)
library(tidyverse)
library(sf)
library(elevatr)

```

Import data

```{r}

plots <- read_excel("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0017\\data\\East Side\\metadata_9-3_DB.xlsx")

trees <- read_excel("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0017\\data\\East Side\\Treedata_9-3_Em_DB.xlsx")

```

Find plot centers in WGS84

```{r}

# right now the plot center coordinates are in UTM11N, EPSG:32611

plots <- st_as_sf(plots, coords = c("Eutm", "Nutm"), crs = 32611)

plots <- st_transform(plots, crs = 4326)

plotcoordinates <- data.frame(plots$Plot, st_coordinates(plots[,1], st_coordinates(plots[,2])))

plotcoordinates <- plotcoordinates %>% rename (Plot=plots.Plot, PlotLongitudeWGS84=X, PlotLatitudeWGS84=Y)

plots <- full_join (plots, plotcoordinates, by="Plot")

plots <- st_drop_geometry(plots)

# export

write.csv(plots, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\plotdata.csv")

```

Add OFO plot IDs

```{r}

trees <- trees %>%
  add_column(ofo_plot_id = "")

for(i in 1:nrow(trees)) {
  if(trees[i,]$Plot == "IS1") {
    trees[i,]$ofo_plot_id = "0197"
  }
  else if (trees[i,]$Plot == "IS2") {
    trees[i,]$ofo_plot_id = "0198"
  }
  else if (trees[i,]$Plot == "IS3") {
    trees[i,]$ofo_plot_id = "0199"
  }
  else if (trees[i,]$Plot == "OH1") {
    trees[i,]$ofo_plot_id = "0200"
  }
  else if (trees[i,]$Plot == "OH2") {
    trees[i,]$ofo_plot_id = "0201"
  }
  else if (trees[i,]$Plot == "OH3") {
    trees[i,]$ofo_plot_id = "0202"
  }
}

```

Remove NAs from notes/comments column

```{r}

trees$Comments[is.na(trees$Comments)] <- ''

```

Rename columns to be more descriptive

```{r}
trees <- trees %>% rename (DBH_cm=DBH, Crown_height_meters=CrHt, Scorch_height_meters=SrchHt, tree_height_above_plot_center_m=Z, distance_x_m=X, distance_y_m=Y)

# turns out the Z column is NOT height above plot center, renaming again

trees <- trees %>% rename (tree_dbh_height_above_plot_center=tree_height_above_plot_center_m)
```

Snag decay class and live/dead

```{r}

# rename column

trees <- trees %>% rename (snag_decay_class=Dec)

# trees with a snag decay class value are dead, trees without a snag decay class value are alive. use this info to populate a live/dead colulmn

trees <- trees %>%
  add_column(live_dead = "")

for(i in 1:nrow(trees)) {
  if(is.na(trees[i,]$snag_decay_class)) {
    trees[i,]$live_dead = "L"
  }
  else if (!is.na(trees[i,]$snag_decay_class)){
    trees[i,]$live_dead = "D"
  }
}

# remove NAs from snag decay class column

trees$snag_decay_class[is.na(trees$snag_decay_class)] <- ''

```

Create species code column

```{r}

trees <- trees %>%
  add_column(spp_codes = "")

trees$spp_codes[trees$Spec == 'ABCO'] <- '15'

trees$spp_codes[trees$Spec == 'BEOC'] <- 'BETOCC'

trees$spp_codes[trees$Spec == 'CELE'] <- '475'

trees$spp_codes[trees$Spec == 'JUGR'] <- '64'

trees$spp_codes[trees$Spec == 'JUN'] <- '57'

trees$spp_codes[trees$Spec == 'PICO'] <- '108'

trees$spp_codes[trees$Spec == 'PIFL'] <- '113'

trees$spp_codes[trees$Spec == 'PIJE'] <- '116'

```

Growth form

```{r}

# only trees were surveyed in this project

trees <- trees %>%
  add_column(growth_form = "tree")

```

Tree lat/lon truthfinding-- this code chunk is not needed to replicate the code in the OFO database, it was just used to help find which permutation of tree coordinates was correct out of all the possibilities

```{r}

# we have the plot centers in UTMs, and the tree distances in x/y meter coordinates. join the plot center UTMs to the tree data frame, and then add the respective x/y coordinates to the UTMs to find the tree UTMs. 

# import plot UTMs

plots <- read_excel("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0017\\data\\East Side\\metadata_9-3_DB.xlsx")

# make a smaller plot data.frame ready to join with the tree data.frame

plots <- plots[,-1]

plots <- plots[,-2]

plots <- plots[,-c(4:7)]

plots <- plots %>% rename(Plot_UTM_N = Nutm, Plot_UTM_E = Eutm)

# merge plot coords into tree data.frame

trees <- dplyr::left_join(trees, plots, by = "Plot")

# add tree x and y distances to the plot UTMs to get tree coordinates

# surprise, this is harder than it might have been! I'm inclined to think the X distance is meters north (+) or south (-) from the plot center, and the Y distance is meters east (+) or west (-) from the plot center, but the data collectors don't remember. so I'll run with this assumption to calculate the tree coordinates, then export them to examine against satellite imagery. 

trees <- trees %>% add_column(Tree_UTM_N = "", Tree_UTM_E = "")

trees$Tree_UTM_N <- as.numeric(trees$Tree_UTM_N)
trees$Tree_UTM_E <- as.numeric(trees$Tree_UTM_E)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N = trees[i,]$Plot_UTM_N + trees[i,]$distance_x_m
  trees[i,]$Tree_UTM_E = trees[i,]$Plot_UTM_E + trees[i,]$distance_y_m
}

trees <- st_as_sf(trees, coords = c("Tree_UTM_E", "Tree_UTM_N"), crs = 32611)

trees_coordinate_export <- trees[,-c(3:4)]

trees_coordinate_export <- trees_coordinate_export[,-c(4)]

trees_coordinate_export <- trees_coordinate_export[,-c(1)]

trees_coordinate_export <- trees_coordinate_export[,-c(6:15)]

st_write(trees_coordinate_export, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\trees_test1.gpkg")

# damn that looks bad, I'll try again

trees <- trees %>% add_column(Tree_UTM_N_2 = "", Tree_UTM_E_2 = "")

trees$Tree_UTM_N_2 <- as.numeric(trees$Tree_UTM_N_2)
trees$Tree_UTM_E_2 <- as.numeric(trees$Tree_UTM_E_2)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N_2 = trees[i,]$Plot_UTM_N - trees[i,]$distance_x_m
  trees[i,]$Tree_UTM_E_2 = trees[i,]$Plot_UTM_E - trees[i,]$distance_y_m
}

trees <- st_drop_geometry(trees)

trees <- st_as_sf(trees, coords = c("Tree_UTM_E_2", "Tree_UTM_N_2"), crs = 32611)

trees_coordinate_export <- trees[,-c(3:4)]

trees_coordinate_export <- trees_coordinate_export[,-c(4)]

trees_coordinate_export <- trees_coordinate_export[,-c(1)]

trees_coordinate_export <- trees_coordinate_export[,-c(6:15)]

st_write(trees_coordinate_export, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\trees_test2.gpkg")

# still looks bad. 3rd try.

trees <- trees %>% add_column(Tree_UTM_N_3 = "", Tree_UTM_E_3 = "")

trees$Tree_UTM_N_3 <- as.numeric(trees$Tree_UTM_N_3)
trees$Tree_UTM_E_3 <- as.numeric(trees$Tree_UTM_E_3)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N_3 = trees[i,]$Plot_UTM_N + trees[i,]$distance_y_m
  trees[i,]$Tree_UTM_E_3 = trees[i,]$Plot_UTM_E + trees[i,]$distance_x_m
}

trees <- st_drop_geometry(trees)

trees <- st_as_sf(trees, coords = c("Tree_UTM_E_3", "Tree_UTM_N_3"), crs = 32611)

trees_coordinate_export <- trees[,-c(3:4)]

trees_coordinate_export <- trees_coordinate_export[,-c(4)]

trees_coordinate_export <- trees_coordinate_export[,-c(1)]

trees_coordinate_export <- trees_coordinate_export[,-c(6:15)]

st_write(trees_coordinate_export, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\trees_test3.gpkg")

# number three looks good I think! but I really doubt myself so I'm going to calculate a fourth possible solution. and possibly the other four possible solutions too idk. 

trees <- trees %>% add_column(Tree_UTM_N_4 = "", Tree_UTM_E_4 = "")

trees$Tree_UTM_N_4 <- as.numeric(trees$Tree_UTM_N_4)
trees$Tree_UTM_E_4 <- as.numeric(trees$Tree_UTM_E_4)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N_4 = trees[i,]$Plot_UTM_N - trees[i,]$distance_y_m
  trees[i,]$Tree_UTM_E_4 = trees[i,]$Plot_UTM_E - trees[i,]$distance_x_m
}

trees <- st_drop_geometry(trees)

trees <- st_as_sf(trees, coords = c("Tree_UTM_E_4", "Tree_UTM_N_4"), crs = 32611)

trees_coordinate_export <- trees[,-c(3:4)]

trees_coordinate_export <- trees_coordinate_export[,-c(4)]

trees_coordinate_export <- trees_coordinate_export[,-c(1)]

trees_coordinate_export <- trees_coordinate_export[,-c(6:15)]

st_write(trees_coordinate_export, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\trees_test4.gpkg")

# 5

trees <- trees %>% add_column(Tree_UTM_N_5 = "", Tree_UTM_E_5 = "")

trees$Tree_UTM_N_5 <- as.numeric(trees$Tree_UTM_N_5)
trees$Tree_UTM_E_5 <- as.numeric(trees$Tree_UTM_E_5)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N_5 = trees[i,]$Plot_UTM_N + trees[i,]$distance_y_m
  trees[i,]$Tree_UTM_E_5 = trees[i,]$Plot_UTM_E - trees[i,]$distance_x_m
}

trees <- st_drop_geometry(trees)

trees <- st_as_sf(trees, coords = c("Tree_UTM_E_5", "Tree_UTM_N_5"), crs = 32611)

trees_coordinate_export <- trees[,-c(3:4)]

trees_coordinate_export <- trees_coordinate_export[,-c(4)]

trees_coordinate_export <- trees_coordinate_export[,-c(1)]

trees_coordinate_export <- trees_coordinate_export[,-c(6:15)]

st_write(trees_coordinate_export, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\trees_test5.gpkg")

# 6

trees <- trees %>% add_column(Tree_UTM_N_6 = "", Tree_UTM_E_6 = "")

trees$Tree_UTM_N_6 <- as.numeric(trees$Tree_UTM_N_6)
trees$Tree_UTM_E_6 <- as.numeric(trees$Tree_UTM_E_6)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N_6 = trees[i,]$Plot_UTM_N - trees[i,]$distance_y_m
  trees[i,]$Tree_UTM_E_6 = trees[i,]$Plot_UTM_E + trees[i,]$distance_x_m
}

trees <- st_drop_geometry(trees)

trees <- st_as_sf(trees, coords = c("Tree_UTM_E_6", "Tree_UTM_N_6"), crs = 32611)

trees_coordinate_export <- trees[,-c(3:4)]

trees_coordinate_export <- trees_coordinate_export[,-c(4)]

trees_coordinate_export <- trees_coordinate_export[,-c(1)]

trees_coordinate_export <- trees_coordinate_export[,-c(6:15)]

st_write(trees_coordinate_export, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\trees_test6.gpkg")

# 7

trees <- trees %>% add_column(Tree_UTM_N_7 = "", Tree_UTM_E_7 = "")

trees$Tree_UTM_N_7 <- as.numeric(trees$Tree_UTM_N_7)
trees$Tree_UTM_E_7 <- as.numeric(trees$Tree_UTM_E_7)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N_7 = trees[i,]$Plot_UTM_N - trees[i,]$distance_x_m
  trees[i,]$Tree_UTM_E_7 = trees[i,]$Plot_UTM_E + trees[i,]$distance_y_m
}

trees <- st_drop_geometry(trees)

trees <- st_as_sf(trees, coords = c("Tree_UTM_E_7", "Tree_UTM_N_7"), crs = 32611)

trees_coordinate_export <- trees[,-c(3:4)]

trees_coordinate_export <- trees_coordinate_export[,-c(4)]

trees_coordinate_export <- trees_coordinate_export[,-c(1)]

trees_coordinate_export <- trees_coordinate_export[,-c(6:15)]

st_write(trees_coordinate_export, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\trees_test7.gpkg")

# 8

trees <- trees %>% add_column(Tree_UTM_N_8 = "", Tree_UTM_E_8 = "")

trees$Tree_UTM_N_8 <- as.numeric(trees$Tree_UTM_N_8)
trees$Tree_UTM_E_8 <- as.numeric(trees$Tree_UTM_E_8)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N_8 = trees[i,]$Plot_UTM_N + trees[i,]$distance_x_m
  trees[i,]$Tree_UTM_E_8 = trees[i,]$Plot_UTM_E - trees[i,]$distance_y_m
}

trees <- st_drop_geometry(trees)

trees <- st_as_sf(trees, coords = c("Tree_UTM_E_8", "Tree_UTM_N_8"), crs = 32611)

trees_coordinate_export <- trees[,-c(3:4)]

trees_coordinate_export <- trees_coordinate_export[,-c(4)]

trees_coordinate_export <- trees_coordinate_export[,-c(1)]

trees_coordinate_export <- trees_coordinate_export[,-c(6:15)]

st_write(trees_coordinate_export, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\trees_test8.gpkg")

```

I thought permutation #3 looked the most correct and Derek confirmed by generating a canopy height model. Going to add the final correct tree lat/lons.

```{r}

# import plot UTMs

plots <- read_excel("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0017\\data\\East Side\\metadata_9-3_DB.xlsx")

# make a smaller plot data.frame ready to join with the tree data.frame

plots <- plots[,-1]

plots <- plots[,-2]

plots <- plots[,-c(4:7)]

plots <- plots %>% rename(Plot_UTM_N = Nutm, Plot_UTM_E = Eutm)

# merge plot coords into tree data.frame

trees <- dplyr::left_join(trees, plots, by = "Plot")

# find tree coords

trees <- trees %>% add_column(Tree_UTM_N = "", Tree_UTM_E = "")

trees$Tree_UTM_N <- as.numeric(trees$Tree_UTM_N)
trees$Tree_UTM_E <- as.numeric(trees$Tree_UTM_E)

for(i in 1:nrow(trees)) {
  trees[i,]$Tree_UTM_N = trees[i,]$Plot_UTM_N + trees[i,]$distance_y_m
  trees[i,]$Tree_UTM_E = trees[i,]$Plot_UTM_E + trees[i,]$distance_x_m
}

trees <- st_as_sf(trees, coords = c("Tree_UTM_E", "Tree_UTM_N"), crs = 32611)

# transform into correct crs by creating a new dataframe of tree coords and merging into trees data.frame

trees <- st_transform(trees, crs = 4326)

trees$ID <- 1:nrow(trees)

treecoordinates <- data.frame(trees$ID, st_coordinates(trees[,1], st_coordinates(trees[,2])))

treecoordinates <- treecoordinates %>% rename (ID=trees.ID, TreeLongitudeWGS84=X, TreeLatitudeWGS84=Y)

trees <- full_join (trees, treecoordinates, by="ID")

trees <- st_drop_geometry(trees)

trees <- trees[,-c(21:23)]

```

Tree scorch heights and height to live crown are the actual heights, not the heights above plot center! Remove NAs

```{r}

trees$Crown_height_meters[is.na(trees$Crown_height_meters)] <- ''

trees$Scorch_height_meters[is.na(trees$Scorch_height_meters)] <- ''

```

Export

```{r}

write.csv(trees, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\O'Harrell_Canyon_Indiana_Summit\\OHarrellCanyonIndianaSummitTreesOFO.csv")

```

