---
title: "FOCAL_early_regen_dispersal"
author: "Emily Marie Purvis"
date: "2024-01-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA MANIPULATED IN R

```{r, echo=FALSE}

# Load libraries

library(tidyverse)
library(sf)
library(stringr)
library(pracma)
```

Load data

```{r}

# set working directory

setwd("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data")

#import data

treedata <- read.csv("C:\\Users\\emily\\Box\\FOCAL\\ofo-field-data\\1_received-data\\0011\\data\\dispersal-data-entry-2023-stemmap_tree.csv", header = TRUE)

```

Plot ID

```{r}

treedata$`ofo_plot_id` <- ""

for(i in 1:nrow(treedata)) {
  if(treedata$fire[i] == 'CA-SM05') {
    if(treedata$stem_map_id[i] == '4') {
      treedata[i,]$ofo_plot_id = '0352'
    }
    else (treedata[i,]$ofo_plot_id = '0353')
  }
else (treedata[i,]$ofo_plot_id = '0351')
}

```

Tree lat and tree long

```{r}

#### give trees ID numbers ####

treedata = treedata %>% mutate(tree_id = 1:nrow(treedata))

#### Convert plot center coordinates to UTMs ####

# convert to spatial dataframe with plot center lat and long

treedata_spatial <- st_as_sf(treedata, coords = c("centerpoint_longitude", "centerpoint_latitude"), crs = 4326, remove=F)

# convert to UTMs

treedata_spatial <- treedata_spatial %>% st_transform(32610)

# create new dataframe with UTM coordinates

treedata_spatial_coordinates <- data.frame(treedata_spatial$tree_id, st_coordinates(treedata_spatial[,1], st_coordinates(treedata_spatial[,2]))) 

# rename new columns

treedata_spatial_coordinates <- treedata_spatial_coordinates %>% rename (`tree_id`=treedata_spatial.tree_id, PlotLongitudeUTM=X, PlotLatitudeUTM=Y) 

# merge into tree data frame

treedata <- full_join(treedata, treedata_spatial_coordinates, by="tree_id")

#### calculate tree coordinates ####

# Make new columns for tree coordinates

treedata$`TreeLongitudeUTM10N` <- 0

treedata$`TreeLatitudeUTM10N` <- 0

# Calculate tree coordinates in UTMs 

# Longitude UTM10N

for(i in 1:nrow(treedata)) {
    treedata[i,]$`TreeLongitudeUTM10N` = (treedata[i,]$PlotLongitudeUTM) + ((sin(deg2rad(treedata[i,]$azimuth))) * (treedata[i,]$`distance`))
}

# Latitude UTM10N

for(i in 1:nrow(treedata)) {
    treedata[i,]$`TreeLatitudeUTM10N` = (treedata[i,]$PlotLatitudeUTM) + ((sin(deg2rad(treedata[i,]$azimuth))) * (treedata[i,]$`distance`))
}

# Convert tree coordinates to WGS84

# make a spatial data frame

treecoordinatesconversion <- data.frame(treedata$tree_id, treedata$TreeLongitudeUTM10N, treedata$TreeLatitudeUTM10N) 

treecoordinatesconversion <- treecoordinatesconversion %>% rename (tree_id=treedata.tree_id, TreeLongitudeUTM10N=treedata.TreeLongitudeUTM10N, TreeLatitudeUTM10N=treedata.TreeLatitudeUTM10N)

treecoordinatesconversion <- st_as_sf(treecoordinatesconversion, coords = c("TreeLongitudeUTM10N", "TreeLatitudeUTM10N"), crs = 32610, remove=F)

# change the CRS

treecoordinatesconversion <- treecoordinatesconversion %>% st_transform(4326)

# extract the geometry features into columns of lat and long WGS84, then merge back into the tree data frame

treecoordinatesconversionWGS84 <- data.frame(treecoordinatesconversion$tree_id, st_coordinates(treecoordinatesconversion[,1], st_coordinates(treecoordinatesconversion[,2]))) 

treecoordinatesconversionWGS84 <- treecoordinatesconversionWGS84 %>% rename (tree_id=treecoordinatesconversion.tree_id, TreeLongitudeWGS84=X, TreeLatitudeWGS84=Y)

treedata <- full_join (treedata, treecoordinatesconversionWGS84, by="tree_id")

```

Height is already in meters, DBH is already in centimeters

Species

```{r}

treedata$`species`[treedata$`species` == 'abco'] <- '15' 
treedata$`species`[treedata$`species` == 'cade'] <- '81' 
treedata$`species`[treedata$`species` == 'pipj'] <- '150' # a code we made up for "either PIPO or PIJE", not an official FIA species code number either

```

Growth form

```{r}

treedata$`growth_form` <- ''

treedata$`growth_form`[treedata$`species` == '15'] <- 'tree'  #'ABCO'
treedata$`growth_form`[treedata$`species` == '81'] <- 'tree' # 'CADE27'
treedata$`growth_form`[treedata$`species` == '150'] <- 'tree' # either PIPO or PIJE, "PIPJ" in terms of this dataset
```

Live/dead

```{r}

treedata$`live_dead` <- ''

treedata$`live_dead`[treedata$`defects` == 'M'] <- 'D'
treedata$`live_dead`[treedata$`defects` == 'M, F'] <- 'D'
treedata$`live_dead`[treedata$`defects` == 'M, B'] <- 'D'
treedata$`live_dead`[treedata$`live_dead` == ''] <- 'L'

```

Crown position

```{r}

treedata$`can_pos`[treedata$`can_pos` == 'I'] <- '4' 
treedata$`can_pos`[treedata$`can_pos` == 'D'] <- '2' 
treedata$`can_pos`[treedata$`can_pos` == 'C'] <- '3' 
treedata$`can_pos`[treedata$`can_pos` == 'S'] <- '5' 

```

Overhead visibility

```{r}

# add a new column for ohvis

treedata <- treedata %>%
  add_column(ohvis = "")

treedata$can_pos <- as.numeric(treedata$can_pos)

# calculate ohvis from canopy position

treedata$ohvis <- case_match(
  treedata$can_pos,
  c(1, 2, 3) ~ "TRUE",
  c(4, 5) ~ "FALSE",
  .default = ""
)

```

Damage codes

```{r}

treedata$defects <- str_replace(treedata$defects, "B", "90001") # broken top
treedata$defects <- str_replace(treedata$defects, "F", "90004") # forked top
treedata$defects <- str_replace(treedata$defects, "C", "90006") # crook/sweep
treedata$defects <- str_replace(treedata$defects, "D", "90002") # dead top
treedata$defects <- str_replace(treedata$defects, "L", "90006") # lean, there isn't a code for lean but it's pretty close to crook/sweep
treedata$defects <- str_replace(treedata$defects, "S", "90006") # crook/sweep

```

Distance to face is the same as the distance column that already exists in the dataset

Distance to pith

```{r}

treedata <- treedata %>%
  add_column(horiz_distance_pith = "")

# calculate

treedata$horiz_distance_pith = treedata$'distance' + (0.5 * 0.01 * treedata$'dbh')

```

Export CSV

```{r}

write.csv(treedata, "C:\\Users\\emily\\Box\\FOCAL\\field-data-standardization\\FOCAL_early_regen_dispersal\\20240110_FOCAL_early_regen_dispersal.csv")

```

## DATA MANIPULATED IN EXCEL

Damage codes manually seperated into new columns





